---
title: "Generating_Datasets"
author: "Tessa Rehill"
output: html_document
date: "2024-10-25"
goal: 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages and data
```{r}
#load packages 
library(dplyr)
library(tidyr)
library(lubridate)

#read in data
hakai<- read.csv("./raw_data/Hakai/quadrats_data.csv") #most up to date data (October 2024)

hakai_sites <- read.csv("./raw_data/Hakai/hakai_sites.csv") #metadata on site location

rls <- read.csv("./raw_data/RLS/IMOS_2.csv") #most up to date data (October 2024)

dfo_count <- read.csv("./raw_data/DFO_surveys/Urchin_Count.csv")

dfo_BHM <- read.csv("./raw_data/DFO_surveys/BHM_PA_algae_update.csv")
  
dfo_Msp <- read.csv("./raw_data/DFO_surveys/MSp_habitat_final.csv")

dfo_GSU <- read.csv("./raw_data/DFO_surveys/GSU_habitat.csv") 

dfo_transects <- read.csv("./raw_data/DFO_surveys/MERGED_TRANSECT_RA.csv") #transect data 

ROV_counts <- read.csv("./processed_data/Baum_ROV/Baum_ROV_Urchin_Counts.csv") #transect data 

#remove ALL BATI DATA 
ROV_counts <- ROV_counts[ROV_counts$site != "Broughtons", ]

```

#Formatting Hakai Data 
##Quadrat level summaries 
```{r}
#Species code for green urchin (gur), purple urchin (pur) and red urchin (rur) - Hakai
#Total length refers to test diameter 

#Filter the dataframe for habitat (rockyreef- this is where there is depth associated with quadrats)
hakai_depth  <- hakai %>%
  filter(survey_method %in% c('Rocky Reef', 'Both')) %>%
  filter(!is.na(depth)) #remove all rows where depth is NA (not useful)

#QUADRAT LEVEL SUMMARY 
#Create a new column for pur, gur, rur counts, summarize by quadrat 
#Create counts for each species by quadrat and site, including no observations
hakai_depth_counts <- hakai_depth %>%
  group_by(date, quadrat, site, depth) %>%
  summarise(
    gur_count = sum(species == "gur", na.rm = TRUE),
    pur_count = sum(species == "pur", na.rm = TRUE),
    rur_count = sum(species == "rur", na.rm = TRUE)
  ) %>%
  ungroup()

#Add a column that counts the number of quadrats per site (for area, and eventually density calculations)
hakai_depth_counts <- hakai_depth_counts %>%
  group_by(site, date) %>%
  mutate(quadrat_count = n()) %>%  # count number of quadrats per site and date
  ungroup()

#Add lat and long
hakai_depth_final <- hakai_depth_counts %>%
  left_join(hakai_sites, by = "site")

#Add presence/absence and add relative density and add absolute density 
#Summarize quadrat level total urchin counts
quadrat_level_summary <- hakai_depth_final %>%
  group_by(date, site, quadrat) %>%
  summarise(
    total_gur = sum(gur_count, na.rm = TRUE),
    total_pur = sum(pur_count, na.rm = TRUE),
    total_rur = sum(rur_count, na.rm = TRUE)
  ) %>%
  # Add relative density, absolute density, and presence/absence columns
  mutate(
    # Relative density categories
    rd_gur = case_when(
      total_gur == 0 ~ "None",
      total_gur > 0 & total_gur <= 10 ~ "Few",
      total_gur > 10 & total_gur <= 100 ~ "Many",
      total_gur > 100 ~ "Abundant"
    ),
    rd_pur = case_when(
      total_pur == 0 ~ "None",
      total_pur > 0 & total_pur <= 10 ~ "Few",
      total_pur > 10 & total_pur <= 100 ~ "Many",
      total_pur > 100 ~ "Abundant"
    ),
    rd_rur = case_when(
      total_rur == 0 ~ "None",
      total_rur > 0 & total_rur <= 10 ~ "Few",
      total_rur > 10 & total_rur <= 100 ~ "Many",
      total_rur > 100 ~ "Abundant"
    ),
    # Presence/absence indicators
    p_gur = ifelse(total_gur > 0, 1, 0),  # 1 if present, 0 if absent
    p_pur = ifelse(total_pur > 0, 1, 0),
    p_rur = ifelse(total_rur > 0, 1, 0),
    # Absolute density calculations
    ad_gur = total_gur / 1,  # Absolute density for green urchins
    ad_pur = total_pur / 1,   # Absolute density for purple urchins
    ad_rur = total_rur / 1    # Absolute density for red urchins
  )

#Join the quadrat-level relative density back to the main dataframe
quadrat_final <- hakai_depth_final %>%
  left_join(quadrat_level_summary %>%
              select(site, quadrat, rd_gur, rd_pur, rd_rur, date, ad_gur, ad_pur, ad_rur, p_gur, p_rur, p_pur),
            by = c("site","date", "quadrat")) 

#quadrat final, hakai_depth_final and quadrat_level_summary all have the same number of rows (180), no data being lost

#Add 'survey_type' column with all values as 'Hakai'
quadrat_final <- quadrat_final %>%
  mutate(survey_type = 'Hakai')

# Separate the date column into year, month, and day
quadrat_final <- quadrat_final %>%
  mutate(
    year = year(date),
    month = month(date),
    day = day(date)
  )

# Rename columns
quadrat_final <- quadrat_final %>%
  rename(start_lat = lat, start_lon = lon, survey_id = quadrat)

#update min and max depth based on pre-determined bins 
#filter out any record with no depth data 
quadrat_final <- quadrat_final %>% filter(!is.na(depth))
#check the maximum and minimum depth of any record
max(quadrat_final$depth) # it is 18 m
min(quadrat_final$depth) # it is -5 m

#Initialize the columns
quadrat_final$max_depth <- NA
quadrat_final$min_depth <- NA


#loop through each row of the dataset and set a max and min depth based on the ROV dataset's depth bins
for (i in 1:length(quadrat_final$site)){
  if (quadrat_final$depth[i] >= -6 & quadrat_final$depth[i] < -3){
    quadrat_final$max_depth[i] <- -3
    quadrat_finalp$min_depth[i] <- -6
  }
  if (quadrat_final$depth[i] >= -3 & quadrat_final$depth[i] < 0){
    quadrat_final$max_depth[i] <- 0
    quadrat_final$min_depth[i] <- -3
  }
  if (quadrat_final$depth[i] >= 0 & quadrat_final$depth[i] < 3){
    quadrat_final$max_depth[i] <- 3
    quadrat_final$min_depth[i] <- 0
  }
  if (quadrat_final$depth[i] >= 3 & quadrat_final$depth[i] < 6){
    quadrat_final$max_depth[i] <- 6
    quadrat_final$min_depth[i] <- 3
  }
  if (quadrat_final$depth[i] >= 6 & quadrat_final$depth[i] < 9){
    quadrat_final$max_depth[i] <- 9
    quadrat_final$min_depth[i] <- 6
  }
  if (quadrat_final$depth[i] >= 9 & quadrat_final$depth[i] < 12){
    quadrat_final$max_depth[i] <- 12
   quadrat_final$min_depth[i] <- 9
  }
  if (quadrat_final$depth[i] >= 12 & quadrat_final$depth[i] < 15){
    quadrat_final$max_depth[i] <- 15
    quadrat_final$min_depth[i] <- 12
  }
  if (quadrat_final$depth[i] >= 15 & quadrat_final$depth[i] < 18){
    quadrat_final$max_depth[i] <- 18
    quadrat_final$min_depth[i] <- 15
  }
  if (quadrat_final$depth[i] >= 18 & quadrat_final$depth[i] < 21){
    quadrat_final$max_depth[i] <- 21
    quadrat_final$min_depth[i] <- 18
  }
}

#Add columns to match other datasets but fill with NA 

quadrat_final$substrate <- NA
quadrat_final$end_lon <- NA
quadrat_final$end_lat <- NA 

#Reorder the columns to have survey_type, lat, and long at the beginning
quadrat_final <- quadrat_final %>%
  select(survey_type, survey_id, site, start_lat, start_lon, end_lat, end_lon, year, month, day, depth, min_depth, max_depth, substrate, everything(), -quadrat_count, -date)

#write dataframe as a new file
write.csv(quadrat_final, 
          "./processed_data/Hakai/hakai_quadrat_summary.csv", 
          row.names = FALSE)

```

#Formatting Hakai Data 
##Site level summaries 
```{r}
##summarize to the site level (group each site and make calculations at this scale)
##depth is an average

#SITE LEVEL SUMMARY 
#summarize site-level total urchin counts and depth statistics
site_level_summary <- hakai_depth_final %>%
  group_by(date, site) %>%
  summarise(
    total_gur = sum(gur_count, na.rm = TRUE),
    total_pur = sum(pur_count, na.rm = TRUE),
    total_rur = sum(rur_count, na.rm = TRUE),
    depth = mean(depth, na.rm = TRUE),  # Average depth for each site and date
    min_depth = min(depth, na.rm = TRUE),   # Minimum depth for each site and date
    max_depth = max(depth, na.rm = TRUE),   # Maximum depth for each site and date
    num_quadrats = n(),                     # Count the number of quadrats for each site and date
    .groups = "drop"                        # Drop grouping after summarizing
  ) %>%
  # Add relative density, presence/absence columns, and absolute density calculations
  mutate(
    rd_gur = case_when(
      total_gur == 0 ~ "None",
      total_gur > 0 & total_gur <= 10 ~ "Few",
      total_gur > 10 & total_gur <= 100 ~ "Many",
      total_gur > 100 ~ "Abundant"
    ),
    rd_pur = case_when(
      total_pur == 0 ~ "None",
      total_pur > 0 & total_pur <= 10 ~ "Few",
      total_pur > 10 & total_pur <= 100 ~ "Many",
      total_pur > 100 ~ "Abundant"
    ),
    rd_rur = case_when(
      total_rur == 0 ~ "None",
      total_rur > 0 & total_rur <= 10 ~ "Few",
      total_rur > 10 & total_rur <= 100 ~ "Many",
      total_rur > 100 ~ "Abundant"
    ),
    # Add presence/absence columns
    p_gur = ifelse(total_gur > 0, 1, 0),
    p_pur = ifelse(total_pur > 0, 1, 0),
    p_rur = ifelse(total_rur > 0, 1, 0),
    # Calculate absolute density for each species
    ad_gur = total_gur / num_quadrats,  # Absolute density for green urchins
    ad_pur = total_pur / num_quadrats,  # Absolute density for purple urchins
    ad_rur = total_rur / num_quadrats   # Absolute density for red urchins
  )

#Join the site-level relative density back to the main dataframe
site_final <- site_level_summary %>%
  left_join(
    hakai_depth_final %>%
      select(site, lat, lon) %>%     # Select only relevant columns
      distinct(),                    # Keep only unique rows to avoid duplication
    by = "site"                      # Join by 'site'
  )

#Add 'survey_type' column with all values as 'Hakai'
site_final <- site_final %>%
  mutate(survey_type = 'Hakai')

#Add 'survey_id' column 
site_final <- site_final %>%
  mutate(survey_id = site)

#Add 'survey_type' column with all values as 'Hakai'
site_final <- site_final %>%
  mutate(survey_type = 'Hakai')

# Separate the date column into year, month, and day
site_final <- site_final %>%
  mutate(
    year = year(date),
    month = month(date),
    day = day(date)
  )

#Add columns to match other datasets but fill with NA 
site_final$substrate <- NA
site_final$end_lon <- NA
site_final$end_lat <- NA 

#Reorder the columns to have survey_type, lat, and long at the beginning
site_final <- site_final %>%
  select(survey_type, survey_id, site, lat, lon, end_lat, end_lon, year, month, day, depth, min_depth, max_depth, substrate, everything(),-date, -num_quadrats)


#Rename columns
site_final <- site_final %>%
  rename(start_lat = lat, start_lon = lon)


#Write dataframe as a new file
write.csv(site_final, 
          "./processed_data/Hakai/hakai_site_summary.csv", 
          row.names = FALSE)


```

#Formatting RLS data - transect summaries
```{r}
#A block refers to survey unit within site. There are two blocks along the same transect
#A transect is 50m, with one block on the left 1m and another on the right 1m 
#Depth is a mean
#Survey ID = transect (includes both blocks)

#Remove all rls data that does NOT contain depth information 
rls <- rls[!is.na(rls$depth), ]

#Define species names mapping to codes
species_mapping <- c(
  "Strongylocentrotus purpuratus" = "pur",
  "Strongylocentrotus droebachiensis" = "gur",
  "Mesocentrotus franciscanus" = "rur"
)

#Map relevant species to to shortened names
#this new dataframe contains only data from rls WITH urchin species, but don't worry we include surveys later without! 
rls_filtered <- rls %>%
  filter(species_name %in% names(species_mapping)) %>%
  mutate(
    species_code = case_when(
      species_name == "Strongylocentrotus purpuratus" ~ "pur",
      species_name == "Strongylocentrotus droebachiensis" ~ "gur",
      species_name == "Mesocentrotus franciscanus" ~ "rur"
    )
  )

#TRANSECT LEVEL SUMMARY 
#summarize depth at transect level, find mean depth within a block (for all transects)
#survey ID is the same for transect, if min and max depth = depth, then fill with NA


mean_depth_df <- rls %>%
  group_by(survey_id) %>%
  summarise(
    depth = mean(depth, na.rm = TRUE))
#add to urchin only data
rls_transect <- rls %>%
  left_join(mean_depth_df, by = "survey_id")

#summarize counts of each urchin speceis by survey ID in rls filtered
urchin_counts <- rls_filtered %>%
  group_by(survey_id, species_code) %>%
  summarise(total_urchins = sum(total, na.rm = TRUE)) %>%
  ungroup()

#pivot the data so that each species (gur, rur, pur) becomes a column
urchin_counts_pivot <- urchin_counts %>%
  pivot_wider(names_from = species_code, values_from = total_urchins, values_fill = 0)

#Merge with the full rls dataframe to include surveys without urchins
rls_summary <- rls_transect %>%
  select(survey_id, site_name, latitude, longitude, survey_date, depth.y) %>%
  distinct() %>%         # Keep only unique survey_ids
  left_join(urchin_counts_pivot, by = "survey_id") %>%  # Merge with urchin counts
  replace(is.na(.), 0)   # Replace NAs with 0 for surveys with no urchin counts


#Rename columns 
rls_summary <- rls_summary %>%
  rename(
    gur_count = gur,
    pur_count = pur,
    rur_count = rur,
    site = site_name,
    date = survey_date,
    start_lat = latitude,
    start_lon = longitude,
    depth = depth.y
  )

#Create presence/absence columns for each species
transect_summary <- rls_summary %>%
  mutate(
    p_gur = ifelse(gur_count > 0, 1, 0),
    p_pur = ifelse(pur_count > 0, 1, 0),
    p_rur = ifelse(rur_count > 0, 1, 0)
  )

# Add relative and actual density columns for each species
transect_summary <- transect_summary %>%
  mutate(
    rd_gur = case_when(
      gur_count == 0 ~ "None",
      gur_count > 0 & gur_count <= 10 ~ "Few",
      gur_count > 10 & gur_count <= 100 ~ "Many",
      gur_count > 100 ~ "Abundant"
    ),
    rd_pur = case_when(
      pur_count == 0 ~ "None",
      pur_count > 0 & pur_count <= 10 ~ "Few",
      pur_count > 10 & pur_count <= 100 ~ "Many",
      pur_count > 100 ~ "Abundant"
    ),
    rd_rur = case_when(
      rur_count == 0 ~ "None",
      rur_count > 0 & rur_count <= 10 ~ "Few",
      rur_count > 10 & rur_count <= 100 ~ "Many",
      rur_count > 100 ~ "Abundant"
    ),
    # Add actual density columns for each species
    ad_gur = gur_count / 100, #100 because there are 2 x 1 x 50m areas being surveyed (2 blocks)
    ad_pur = pur_count / 100,
    ad_rur = rur_count / 100
  )


# Add 'survey_type' column with all values as 'Hakai'
transect_final <- transect_summary %>%
  mutate(survey_type = 'RL_Survey')


#Separate year month data values
transect_final <- transect_final %>%
  mutate(
    year = year(date),
    month = month(date),
    day = day(date)
  )


#Add columns to match other datasets but fill with NA 
transect_final$substrate <- NA
transect_final$end_lat <- NA
transect_final$end_lon <- NA


#Initialize the columns
transect_final$max_depth <- NA
transect_final$min_depth <- NA


#loop through each row of the dataset and set a max and min depth based on the ROV dataset's depth bins
for (i in 1:length(transect_final$survey_id)){
  if (transect_final$depth[i] >= -6 & transect_final$depth[i] < -3){
    transect_final$max_depth[i] <- -3
    transect_final$min_depth[i] <- -6
  }
  if (transect_final$depth[i] >= -3 & transect_final$depth[i] < 0){
    transect_final$max_depth[i] <- 0
    transect_final$min_depth[i] <- -3
  }
  if (transect_final$depth[i] >= 0 & transect_final$depth[i] < 3){
    transect_final$max_depth[i] <- 3
    transect_final$min_depth[i] <- 0
  }
  if (transect_final$depth[i] >= 3 & transect_final$depth[i] < 6){
    transect_final$max_depth[i] <- 6
    transect_final$min_depth[i] <- 3
  }
  if (transect_final$depth[i] >= 6 & transect_final$depth[i] < 9){
    transect_final$max_depth[i] <- 9
    transect_final$min_depth[i] <- 6
  }
  if (transect_final$depth[i] >= 9 & transect_final$depth[i] < 12){
    transect_final$max_depth[i] <- 12
   transect_final$min_depth[i] <- 9
  }
  if (transect_final$depth[i] >= 12 & transect_final$depth[i] < 15){
    transect_final$max_depth[i] <- 15
    transect_final$min_depth[i] <- 12
  }
  if (transect_final$depth[i] >= 15 & transect_final$depth[i] < 18){
    transect_final$max_depth[i] <- 18
    transect_final$min_depth[i] <- 15
  }
  if (transect_final$depth[i] >= 18 & transect_final$depth[i] < 21){
    transect_final$max_depth[i] <- 21
    transect_final$min_depth[i] <- 18
  }
}

#Reorder the columns to have survey_type, lat, and long at the beginning
transect_final <- transect_final %>%
  select(survey_type, survey_id, site, start_lat, start_lon, end_lat, end_lon, year, month, day,depth, min_depth, max_depth, substrate, everything(),-date)


#Write dataframe as a new file
write.csv(transect_final , 
          "./processed_data/RLS/RLS_transect_summary.csv", 
          row.names = FALSE)

```

#Formatting DFO count data (GSU, RSU, and multispecies surveys)
```{r}
#Remove all NA values in depth and urchin counts 
dfo_count <- dfo_count %>%
  filter(!is.na(ChartDepth) & !is.na(CountRSU) & !is.na(CountPSU) & !is.na(CountGSU))

#Make a new column that will serve as a unique identifier for each transect 
dfo_count <- dfo_count %>%
  group_by(Method, Year, Month, Day, LatitudeStart, LongitudeStart, TransectNum) %>%
  mutate(HKey = cur_group_id()) %>% #this HKey is one that we have made and does not directly match the other datasets 
  ungroup()  # Ungroup the data after creating HKey

#make a new column that will serve as a unique identifier for each quadrat
dfo_count <- dfo_count %>%
  group_by(Method, Year, Month, Day, LatitudeStart, LongitudeStart, TransectNum, QuadratNum) %>%
  mutate(QKey = cur_group_id()) %>%
  ungroup()  # Ungroup the data after creating HKey

#QUADRAT LEVEL SUMMARY ####
#Rename the columns in the dfo_count dataframe
dfo_summary <- dfo_count %>%
  rename(
    rur_count = CountRSU,
    pur_count = CountPSU,
    gur_count = CountGSU
  )

#Add a column that counts the number of quadrats per transect (for area, and eventually density calculations)
dfo_summary<- dfo_summary %>%
  group_by(HKey) %>%
  mutate(quadrat_count = n()) %>%  # count number of quadrats per transect
  ungroup()

#Add presence/absence and add relative density and add absolute density 
#Summarize quadrat level total urchin counts
quadrat_level_summary <- dfo_summary %>%
  group_by(QKey) %>%
  summarise(
    total_gur = sum(gur_count, na.rm = TRUE),
    total_pur = sum(pur_count, na.rm = TRUE),
    total_rur = sum(rur_count, na.rm = TRUE)
  ) %>%
  # Add relative density and absolute density columns
  mutate(
    rd_gur = case_when(
      total_gur == 0 ~ "None",
      total_gur > 0 & total_gur <= 10 ~ "Few",
      total_gur > 10 & total_gur <= 100 ~ "Many",
      total_gur > 100 ~ "Abundant"
    ),
    rd_pur = case_when(
      total_pur == 0 ~ "None",
      total_pur > 0 & total_pur <= 10 ~ "Few",
      total_pur > 10 & total_pur <= 100 ~ "Many",
      total_pur > 100 ~ "Abundant"
    ),
    rd_rur = case_when(
      total_rur == 0 ~ "None",
      total_rur > 0 & total_rur <= 10 ~ "Few",
      total_rur > 10 & total_rur <= 100 ~ "Many",
      total_rur > 100 ~ "Abundant"
    ),
    # Add presence/absence columns
    p_gur = ifelse(total_gur > 0, 1, 0),
    p_pur = ifelse(total_pur > 0, 1, 0),
    p_rur = ifelse(total_rur > 0, 1, 0),
    # Calculate absolute density for each species
    ad_gur = total_gur / 1,  # Absolute density for green urchins
    ad_pur = total_pur / 1,  # Absolute density for purple urchins
    ad_rur = total_rur / 1   # Absolute density for red urchins
  )

#Join the quadrat-level relative density back to the main dataframe
quadrat_final <- dfo_summary %>%
  left_join(quadrat_level_summary %>%
              select(QKey,rd_gur, rd_pur, rd_rur, ad_gur, ad_pur, ad_rur, p_gur, p_rur, p_pur),
            by = c("QKey")) 

#quadrat level summary has same number of quadrat (rows) as quadrat_final (no data lost)

#Rename columns 
quadrat_final <- quadrat_final %>%
  rename(
    start_lat = LatitudeStart,
    start_lon = LongitudeStart,
    depth = ChartDepth,
    survey_type = SourceDB,
    survey_id = QKey, 
    site = HKey,
    year = Year, 
    month = Month, 
    day = Day, 
  )


#Add columns from GSU + MSp (metadata files) that are relevant (we do not have this information for RSU)
#How can we link dfoGSU and quadrat final together? - survey_type, site, QuadratNum, TransectNum, date (first need to separate date in dfo_GSU)

#Separate date in GSU data
dfo_GSU <- dfo_GSU %>%
  mutate(
    year = as.integer(format(as.Date(HDate), "%Y")),
    month = as.integer(format(as.Date(HDate), "%m")),
    day = as.integer(format(as.Date(HDate), "%d"))
  )

#There are zero values in latend and lonend of GSU dataset, we need to remove these and turn them into NA values 
dfo_GSU <- dfo_GSU %>%
  mutate(
    LatEnd = ifelse(LatEnd == 0, NA, LatEnd),
    LonEnd = ifelse(LonEnd == 0, NA, LonEnd)
  )

#Separate date in Msp data 
dfo_Msp <- dfo_Msp %>%
  separate(Time, into = c("year", "month", "day"), sep = "-| ", remove = FALSE) %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day)
  )

#We're going to first separate each survey type, add meta data, and then combine them at the end 
#filter quadrat final for only methods GSU 
quadrat_final_GSU <- quadrat_final %>% 
  filter(survey_type == "GSU")

#now for only MSP
quadrat_final_MSP <- quadrat_final %>% 
  filter(survey_type == "Multispecies")

#now for only RSU 
quadrat_final_RSU <- quadrat_final %>% 
  filter(survey_type == "RSU")

#perform a left join to bring in dfo_GSU data based on specified columns
quadrat_final_GSU_matched <- quadrat_final_GSU %>%
  left_join(
    dfo_GSU %>%
      select(LatStart, LonStart, year, ChartDepth, LatEnd, LonEnd, SubCat, SubSubCat,month, day, Quadrat) %>%
      distinct(),
    by = c("start_lat" = "LatStart", 
           "start_lon" = "LonStart", 
           "year" = "year", 
           "month" = "month", 
           "day" = "day", 
           "depth" = "ChartDepth", 
           "QuadratNum" = "Quadrat")  # Ensure mapping is correct
  ) #same number of observations, no data lost 

#rename columns to match end and start latitude 
quadrat_final_GSU_matched <- quadrat_final_GSU_matched %>%
  rename(
    end_lat = LatEnd,
    end_lon = LonEnd
  )

#Replace substrate category values with actual names for categories
quadrat_final_GSU_matched <- quadrat_final_GSU_matched %>%
  #Create the 'substrate' column by combining 'SubCat' and 'SubSubCat'
  mutate(substrate = paste0(SubCat, SubSubCat)) %>%
  #Recode the combined values
  mutate(
    substrate = case_when( #based on key 
      substrate == "1a" ~ "Bedrock",
      substrate == "1b" ~ "Boulder",
      substrate == "2" ~ "Mixed",
      substrate == "3a" ~ "Sand",
      substrate == "3b" ~ "Mud",
      TRUE ~ NA_character_  # Keep NA for any unmatched cases
    )
  )%>%
  #Remove 'SubCat' and 'SubSubCat' columns
  select(-SubCat, -SubSubCat)


#perform a left join to bring dfo_Msp data based on specified columns 
quadrat_final_MSP_matched <- quadrat_final_MSP %>%
  left_join(
    dfo_Msp %>%
      select(latStart, lonStart, year, CorDepthM, latEnd, lonEnd, SubCat, SubSubCat, month, day, Quadrat) %>%
      distinct(),
    by = c("start_lat" = "latStart", 
           "start_lon" = "lonStart", 
           "year" = "year", 
           "month" = "month", 
           "day" = "day", 
           "depth" = "CorDepthM", 
           "QuadratNum" = "Quadrat")  # Ensure mapping is correct
) #same number, no data lost 

#rename columns to match end and start latitude 
quadrat_final_MSP_matched <- quadrat_final_MSP_matched %>%
  rename(
    end_lat = latEnd,
    end_lon = lonEnd
  )

#Replace substrate category values with actual names for categories
quadrat_final_MSP_matched <- quadrat_final_MSP_matched %>%
  #Create the 'substrate' column by combining 'SubCat' and 'SubSubCat'
  mutate(substrate = paste0(SubCat, SubSubCat)) %>%
  #Recode the combined values, based on key 
  mutate(
    substrate = case_when(
      substrate == "1a" ~ "Bedrock",
      substrate == "1b" ~ "Boulder",
      substrate == "2" ~ "Mixed",
      substrate == "3a" ~ "Sand",
      substrate == "3b" ~ "Mud",
      TRUE ~ NA_character_  #Keep NA for any unmatched cases
    )
  ) %>%
  # Remove 'SubCat' and 'SubSubCat' columns
  select(-SubCat, -SubSubCat)


#we dont have additional data for RSU do this remains all NA 
quadrat_final_RSU_matched <- quadrat_final_RSU %>%
  mutate(
    end_lat = NA,
    end_lon = NA,
    substrate = NA
  )

#quadrat final before combining data has 138312 rows 
#now rbind them all back together, should have same amount of rows as original quadrat_final (138312)
quadrat_final <- rbind(quadrat_final_RSU_matched, quadrat_final_GSU_matched, quadrat_final_MSP_matched) #same amount of rows, no data lost

#survey type should be a combination of the database that it was extracted from and the method
quadrat_final <- quadrat_final %>%
  mutate(survey_type = paste(survey_type, Method, sep = "_"))
  
#Initialize the columns
quadrat_final$max_depth <- NA
quadrat_final$min_depth <- NA


#loop through each row of the dataset and set a max and min depth based on the ROV dataset's depth bins
for (i in 1:length(quadrat_final$Key)){
  if (quadrat_final$depth[i] >= -6 & quadrat_final$depth[i] < -3){
    quadrat_final$max_depth[i] <- -3
    quadrat_final$min_depth[i] <- -6
  }
  if (quadrat_final$depth[i] >= -3 & quadrat_final$depth[i] < 0){
    quadrat_final$max_depth[i] <- 0
    quadrat_final$min_depth[i] <- -3
  }
  if (quadrat_final$depth[i] >= 0 & quadrat_final$depth[i] < 3){
    quadrat_final$max_depth[i] <- 3
    quadrat_final$min_depth[i] <- 0
  }
  if (quadrat_final$depth[i] >= 3 & quadrat_final$depth[i] < 6){
    quadrat_final$max_depth[i] <- 6
    quadrat_final$min_depth[i] <- 3
  }
  if (quadrat_final$depth[i] >= 6 & quadrat_final$depth[i] < 9){
    quadrat_final$max_depth[i] <- 9
    quadrat_final$min_depth[i] <- 6
  }
  if (quadrat_final$depth[i] >= 9 & quadrat_final$depth[i] < 12){
    quadrat_final$max_depth[i] <- 12
   quadrat_final$min_depth[i] <- 9
  }
  if (quadrat_final$depth[i] >= 12 & quadrat_final$depth[i] < 15){
    quadrat_final$max_depth[i] <- 15
    quadrat_final$min_depth[i] <- 12
  }
  if (quadrat_final$depth[i] >= 15 & quadrat_final$depth[i] < 18){
    quadrat_final$max_depth[i] <- 18
    quadrat_final$min_depth[i] <- 15
  }
  if (quadrat_final$depth[i] >= 18 & quadrat_final$depth[i] < 21){
    quadrat_final$max_depth[i] <- 21
    quadrat_final$min_depth[i] <- 18
  }
}

# Reorder the columns to have survey_type, lat, and long at the beginning
quadrat_final <- quadrat_final %>%
  select(survey_type, survey_id, site, start_lat, start_lon, end_lat, end_lon, year, month, day, depth, min_depth, max_depth, substrate, everything(), -quadrat_count, -Key, -Method, -StartTime, -TransectNum, -QuadratNum)
 
#write dataframe as a new file
write.csv(quadrat_final, 
          "./processed_data/DFO_surveys/dfo_quadrat_summary.csv", 
          row.names = FALSE)


#TRANSECT LEVEL SUMMARY ####
#Each HKey is a unique transects 
#Rename the columns in the dfo_count dataframe

dfo_summary <- dfo_count %>%
  rename(
    rur_count = CountRSU,
    pur_count = CountPSU,
    gur_count = CountGSU
  )

#Add a column that counts the number of quadrats per transect (for area, and eventually density calculations)
dfo_summary <- dfo_summary %>%
  group_by(HKey) %>%
  mutate(area = n()) %>%  # count number of quadrats per Key/Transect  
  ungroup()

# dd column that takes the mean, min and maximum depths for each Key/Transect 
mean_depth_df <- dfo_summary %>%
  group_by(HKey) %>%
  summarise(depth = mean(ChartDepth, na.rm = TRUE),
            min_depth = min(ChartDepth, na.rm = TRUE),
            max_depth = max(ChartDepth, na.rm = TRUE),
            .groups = 'drop')  # Drop grouping after summarizing

#Add depth information to dfo_summary
dfo_summary <- dfo_summary %>%
  left_join(mean_depth_df, by = "HKey") %>%
  select(-QuadratNum)

#Summarize transect level total urchin counts, ensuring unique Keys
transect_level_summary <- dfo_summary %>%
  group_by(HKey) %>%
  summarise(
    total_gur = sum(gur_count, na.rm = TRUE),
    total_pur = sum(pur_count, na.rm = TRUE),
    total_rur = sum(rur_count, na.rm = TRUE),
    area = first(area),  # Assuming quadrat_count is the same for each Key
    .groups = 'drop'                        # Drop grouping after summarizing
  ) %>%
  # Add relative density and absolute density columns
  mutate(
    rd_gur = case_when(
      total_gur == 0 ~ "None",
      total_gur > 0 & total_gur <= 10 ~ "Few",
      total_gur > 10 & total_gur <= 100 ~ "Many",
      total_gur > 100 ~ "Abundant"
    ),
    rd_pur = case_when(
      total_pur == 0 ~ "None",
      total_pur > 0 & total_pur <= 10 ~ "Few",
      total_pur > 10 & total_pur <= 100 ~ "Many",
      total_pur > 100 ~ "Abundant"
    ),
    rd_rur = case_when(
      total_rur == 0 ~ "None",
      total_rur > 0 & total_rur <= 10 ~ "Few",
      total_rur > 10 & total_rur <= 100 ~ "Many",
      total_rur > 100 ~ "Abundant"
    ),
    # Add presence/absence columns
    p_gur = ifelse(total_gur > 0, 1, 0),
    p_pur = ifelse(total_pur > 0, 1, 0),
    p_rur = ifelse(total_rur > 0, 1, 0),
    # Calculate absolute density for each species
    ad_gur = total_gur / area,  # Absolute density for green urchins
    ad_pur = total_pur / area,  # Absolute density for purple urchins
    ad_rur = total_rur / area   # Absolute density for red urchins
  )

#Join the transect-level relative density back to the main dataframe 
transect_final <- transect_level_summary %>%
  left_join(dfo_summary %>%
              select(HKey, Method, SourceDB, Year, Month, Day, LatitudeStart, LongitudeStart, depth, min_depth, max_depth), 
            by = "HKey")

transect_final <- distinct(transect_final)

# Survey type should be a combination of method and source database 
transect_final <- transect_final %>%
  mutate(survey_type = paste(SourceDB, Method, sep = "_"))


# Rename columns
transect_final <- transect_final %>%
  rename(
    start_lat = LatitudeStart,  # Assuming LatitudeStart should be in transect_final
    start_lon = LongitudeStart, # Assuming LongitudeStart should be in transect_final
    year = Year, 
    month = Month, 
    day = Day,
    survey_id = HKey
    
  )

#The metadata that nees to be added is substrate + end lat/lon 
#we're going to first separate each survey type, add meta data, and then combine them at the end 
#filter quadrat final for only methods GSU 
transect_final_GSU <- transect_final %>% 
  filter(SourceDB == "GSU")

#now for only MSP
transect_final_MSP <- transect_final %>% 
  filter(SourceDB == "Multispecies")

#now for only RSU 
transect_final_RSU <- transect_final %>% 
  filter(SourceDB == "RSU")

#need to summarize substrate across surveys to map 
#Summarize the most dominant SubSubCat for each transect (HKey) for GSU 
dfo_GSU_sub <- dfo_GSU %>%
  # Combine SubCat and SubSubCat into a single column 'substrate' for grouping
  mutate(substrate = paste0(SubCat, SubSubCat)) %>%
  group_by(HKey, year, month, day, LatStart, LonStart, LatEnd, LonEnd) %>%
  count(substrate, sort = TRUE) %>%            # Count occurrences of each substrate combination within each HKey
  slice_max(n, n = 1, with_ties = FALSE) %>%   # Keep the most frequent substrate
  ungroup() %>%
  # Rename substrate values as needed
  mutate(
    substrate = case_when(
      substrate == "1a" ~ "Bedrock",
      substrate == "1b" ~ "Boulder",
      substrate == "2"  ~ "Mixed",
      substrate == "3a" ~ "Sand",
      substrate == "3b" ~ "Mud",
      TRUE ~ NA_character_
    )
  ) %>%
  # Select relevant columns
  select(HKey, year, month, day, LatStart, LonStart, LatEnd, LonEnd, substrate)


#Summarize the most dominant SubSubCat for each transect (HKey) for Msp 
dfo_MSP_sub <- dfo_Msp %>%
  #Combine SubCat and SubSubCat into a single column 'substrate' for grouping
  mutate(substrate = paste0(SubCat, SubSubCat)) %>%
  group_by(Key, year, month, day, latStart, lonStart, latEnd, lonEnd) %>%
  count(substrate, sort = TRUE) %>%            # Count occurrences of each substrate combination within each HKey
  slice_max(n, n = 1, with_ties = FALSE) %>%   # Keep the most frequent substrate
  ungroup() %>%
  #Rename substrate values as needed
  mutate(
    substrate = case_when(
      substrate == "1a" ~ "Bedrock",
      substrate == "1b" ~ "Boulder",
      substrate == "2"  ~ "Mixed",
      substrate == "3a" ~ "Sand",
      substrate == "3b" ~ "Mud",
      TRUE ~ NA_character_
    )
  ) %>%
  #Select relevant columns
  select(Key, year, month, day, latStart, lonStart, latEnd, lonEnd, substrate)


#perform a left join to bring dfo_GSU data based on specified columns 
transect_final_GSU_matched <- transect_final_GSU %>%
  left_join(
    dfo_GSU_sub %>%
      select(LatStart, LonStart, year, LatEnd, LonEnd,  month, day, substrate, HKey) %>%
      distinct(),
    by = c("start_lat" = "LatStart", 
           "start_lon" = "LonStart", 
           "year" = "year", 
           "month" = "month", 
           "day" = "day", 
           "survey_id"= "HKey")  #Ensure mapping is correct
)

#perform a left join to bring dfo_MSP data based on specified columns 
transect_final_MSP_matched <- transect_final_MSP %>%
  left_join(
    dfo_MSP_sub %>%
      select(latStart, lonStart, year, latEnd, lonEnd,  month, day, substrate) %>%
      distinct(),
    by = c("start_lat" = "latStart", 
           "start_lon" = "lonStart", 
           "year" = "year", 
           "month" = "month", 
           "day" = "day")  #Ensure mapping is correct
)


#rename columns to match end and start latitude 
transect_final_GSU_matched <- transect_final_GSU_matched %>%
  rename(
    end_lat = LatEnd,
    end_lon = LonEnd
  )


#rename columns to match end and start latitude 
transect_final_MSP_matched <- transect_final_MSP_matched %>%
  rename(
    end_lat = latEnd,
    end_lon = lonEnd
  )


#Now for RSU, which again we don't have data for 
transect_final_RSU_matched <- transect_final_RSU %>%
  mutate(
    end_lat = NA,
    end_lon = NA,
    substrate = NA
  )

#transect final = 3935
#now rbind them all back together, 3937 (two additional transects?)
transect_final <- rbind(transect_final_RSU_matched, transect_final_GSU_matched, transect_final_MSP_matched)

transect_final$site <- transect_final$survey_id

# Reorder the columns to have survey_type, lat, and long at the beginning
transect_final <- transect_final %>%
  select(survey_type, survey_id, site, start_lat, start_lon, end_lat, end_lon, year, month, day, depth, min_depth, max_depth, substrate, everything(), -area, -Method, -SourceDB)


#write dataframe as a new file
write.csv(transect_final, 
          "./processed_data/DFO_surveys/dfo_transect_summary.csv", 
          row.names = FALSE)

```

#Relative density from DFO transects (includes MSp and BHM)
```{r}
#need to link metadata contained in dfo_BHM and dfo_Msp based on start latitude and start longitudes 

#this is contained in the MERGED_TRANSECT_FILE 
#information on start latitude and end longitude 
#for BHM - this information can be linked via HKey 
#for MSp - this information can be linked using Start/End (HKey is NOT a unique identifier)
#note - BHMs data does not have any information about purple urchins 

#Filter DFO transect data for only MSp and only BHM source 
mps_transects <- dfo_transects[dfo_transects$Source == "MSp", ]
bhm_transects <- dfo_transects[dfo_transects$Source == "BHM", ]


#Benthic Habitat Mapping (BHM) pre-processing 
#link bhm_transect to dfo_BHM by HKey
bhm_transect_cleaned <- bhm_transects %>% #select only relevant columns 
  select(-Transect, -Latitude, -Longitude, -RA_GDK, -RA_RSC, -RA_PYC, -RA_HRC, 
         -RA_OSC, -RA_BRF, -RA_BRE, -RA_SPF, -RA_SPE, -RA_OCC, -RA_GPA, -RA_TDA)

#Group and summarize dfo_BHM_cleaned by HKey
dfo_BHM_summarized <- dfo_BHM %>%
  group_by(HKey) %>%
  summarise(
    #Keep the first value for Lat/Lon since they are the same for each HKey
    LatDegStart = first(LatDegStart),
    LatMinStart = first(LatMinStart),
    LonDegStart = first(LonDegStart),
    LonMinStart = first(LonMinStart),
    LatDegEnd = first(LatDegEnd),
    LatMinEnd = first(LatMinEnd),
    LonDegEnd = first(LonDegEnd),
    LonMinEnd = first(LonMinEnd),
    
    #Find the most common category for SubCat and SubSubCat
    DomSubCat = names(which.max(table(SubCat))),
    DomSubSubCat = names(which.max(table(SubSubCat))),
    
    # Calculate depth summaries for quadrat
    Min_Quad_Depth = min(CorDepthM, na.rm = TRUE),
    Max_Quad_Depth = max(CorDepthM, na.rm = TRUE),
    MeanDepth = mean(CorDepthM, na.rm = TRUE)
  )

#Left join the cleaned bhm_transect with the summarized dfo_BHM, same number of rows (1177)
bhm_combined <- bhm_transect_cleaned %>%
  left_join(dfo_BHM_summarized, by = "HKey")

#Turn HKey into a character variable
bhm_combined$HKey <- as.character(bhm_combined$HKey)

#Calculate start and end latitudes into decimal degrees
bhm_combined$Start_Lat <- bhm_combined$LatDegStart + (bhm_combined$LatMinStart/60)
bhm_combined$Start_Long <- (bhm_combined$LonDegStart + (bhm_combined$LonMinStart/60))*(-1)
bhm_combined$Shore_Lat <- bhm_combined$LatDegEnd + (bhm_combined$LatMinEnd/60)
bhm_combined$Shore_Long <- (bhm_combined$LonDegEnd + (bhm_combined$LonMinEnd/60))*(-1)

#Replace substrate category values with actual names for categories
#filter out rows where there is no substrate data
#this might change based on new substrate ENV layer 
bhm_combined <- bhm_combined %>% filter(!is.na(DomSubCat)) #there are some NA but very few 
#loop through each row of data and create a SubstrateCat column based on the DFO substrate key
for (i in 1:length(bhm_combined$HKey)){
  if (bhm_combined$DomSubCat[i] == 1){
    if (bhm_combined$DomSubSubCat[i] == "a"){
      bhm_combined$DomSubstrate_Cat[i] <- "Bedrock"
    }
    if (bhm_combined$DomSubSubCat[i] == "b"){
      bhm_combined$DomSubstrate_Cat[i] <-"Boulder"
    }
  }
  if (bhm_combined$DomSubCat[i] == 2){
    bhm_combined$DomSubstrate_Cat[i] <-"Mixed"
  }
  if (bhm_combined$DomSubCat[i] == 3){
    if (bhm_combined$DomSubSubCat[i] == "a"){
      bhm_combined$DomSubstrate_Cat[i] <- "Sand"
    }
    if (bhm_combined$DomSubSubCat[i] == "b"){
      bhm_combined$DomSubstrate_Cat[i] <- "Mud"
    }
  }
}

#Add a column for min and max depth (this is putting the records into pre-determined depth bins from the ROV dataset)
#In the BHM dataset the depth for each quadrat represents the tide corrected depth at the end of the transect (shallowest point). 
#Therefore records will be treated as if they were in the lower of two depth bins if on the edge of two
#At the same time add BHM to the beginning of every HKey character
#check the maximum and minimum depth of any record
max(bhm_combined$MeanDepth) # it is 13.7 m
min(bhm_combined$MeanDepth) # it is -2.6 m

#filter out any record with no depth data 
bhm_combined <- bhm_combined %>% filter(!is.na(MeanDepth))

#loop through each row of the dataset and
for (i in 1:length(bhm_combined$HKey)){
  bhm_combined$HKey[i] <- paste0("BHM", bhm_combined$HKey[i])
  if (bhm_combined$MeanDepth[i] >= -6 & bhm_combined$MeanDepth[i] < -3){
    bhm_combined$Max_Depth[i] <- -3
    bhm_combined$Min_Depth[i] <- -6
  }
  if (bhm_combined$MeanDepth[i] >= -3 & bhm_combined$MeanDepth[i] < 0){
    bhm_combined$Max_Depth[i] <- 0
    bhm_combined$Min_Depth[i] <- -3
  }
  if (bhm_combined$MeanDepth[i] >= 0 & bhm_combined$MeanDepth[i] < 3){
    bhm_combined$Max_Depth[i] <- 3
    bhm_combined$Min_Depth[i] <- 0
  }
  if (bhm_combined$MeanDepth[i] >= 3 & bhm_combined$MeanDepth[i] < 6){
    bhm_combined$Max_Depth[i] <- 6
    bhm_combined$Min_Depth[i] <- 3
  }
  if (bhm_combined$MeanDepth[i] >= 6 & bhm_combined$MeanDepth[i] < 9){
    bhm_combined$Max_Depth[i] <- 9
    bhm_combined$Min_Depth[i] <- 6
  }
  if (bhm_combined$MeanDepth[i] >= 9 & bhm_combined$MeanDepth[i] < 12){
    bhm_combined$Max_Depth[i] <- 12
    bhm_combined$Min_Depth[i] <- 9
  }
  if (bhm_combined$MeanDepth[i] >= 12 & bhm_combined$MeanDepth[i] < 15){
    bhm_combined$Max_Depth[i] <- 15
    bhm_combined$Min_Depth[i] <- 12
  }
  if (bhm_combined$MeanDepth[i] >= 15 & bhm_combined$MeanDepth[i] < 18){
    bhm_combined$Max_Depth[i] <- 18
    bhm_combined$Min_Depth[i] <- 15
  }
  if (bhm_combined$MeanDepth[i] >= 18 & bhm_combined$MeanDepth[i] < 21){
    bhm_combined$Max_Depth[i] <- 21
    bhm_combined$Min_Depth[i] <- 18
  }
}
#Add a column for survey type
bhm_combined$Survey_Type <- "DFO_BHM"

#Select columns we want to include in the final dataset
bhm_combined <- bhm_combined %>% dplyr::select(Survey_Type, HKey, Start_Lat, Start_Long, Shore_Lat, Shore_Long, Year, Month, Day, Max_Quad_Depth, Min_Quad_Depth, DomSubstrate_Cat,MeanDepth, Min_Depth, Max_Depth, RA_RSU, RA_GSU, RA_PSU)

#Rename the columns in the bhm_combined dataset
bhm_combined <- bhm_combined %>%
  rename(
    rd_rur = RA_RSU,
    rd_pur = RA_PSU,
    rd_gur = RA_GSU
  )

#Rename the categories in relative density columns 
#Change the values in the columns rd_rur, rd_pur, and rd_gur
bhm_combined <- bhm_combined %>%
  mutate(
    rd_rur = recode(rd_rur, 'N' = 'None', 'F' = 'Few', 'A' = 'Abundant', 'M' = 'Many'),
    rd_pur = recode(rd_pur, 'N' = 'None', 'F' = 'Few', 'A' = 'Abundant', 'M' = 'Many'),
    rd_gur = recode(rd_gur, 'N' = 'None', 'F' = 'Few', 'A' = 'Abundant', 'M' = 'Many')
  )

#Add columns for presence/absence for each species 
bhm_combined <- bhm_combined %>%
  mutate(
    p_rur = ifelse(rd_rur == 'None', 0, 1),
    p_pur = ifelse(rd_pur == 'None', 0, 1),
    p_gur = ifelse(rd_gur == 'None', 0, 1)
  )

#Multi-species pre-processing 
#dfo_Msp - quadrat data 
#msp_transect - transect data 
#need to be linked via Start Lat and Start Lon 

mps_transect_cleaned <- mps_transects %>%
  select(-Transect, -RA_GDK, -RA_RSC, -RA_PYC, -RA_HRC, 
         -RA_OSC, -RA_BRF, -RA_BRE, -RA_SPF, -RA_SPE, -RA_OCC, -RA_GPA, -RA_TDA)

#Make Year, Month, Day colummn for quad data
#Ensure that Time is in Date-Time format
dfo_Msp$Time <- as.POSIXct(dfo_Msp$Time, format="%Y-%m-%d %H:%M")

# Create new Year, Month, and Day columns
dfo_Msp$Year <- format(dfo_Msp$Time, "%Y")
dfo_Msp$Month <- format(dfo_Msp$Time, "%m")
dfo_Msp$Day <- format(dfo_Msp$Time, "%d")


#Group and summarize dfo_MSP_cleaned by
dfo_MSP_summarized <- dfo_Msp %>%
  group_by(Key, Year, Month, Day) %>%
  summarise(
    # Keep the first value for Lat/Lon since they are the same for each Key
    latStart = first(latStart),
    lonStart = first(lonStart),
    latEnd = first(latEnd),
    lonEnd = first(lonEnd),

    
    #Find the most common category for SubCat and SubSubCat
    DomSubCat = names(which.max(table(SubCat))),
    DomSubSubCat = names(which.max(table(SubSubCat))),
    
    # Calculate depth summaries
    Min_Quad_Depth = min(CorDepthM, na.rm = TRUE),
    Max_Quad_Depth = max(CorDepthM, na.rm = TRUE),
    MeanDepth = mean(CorDepthM, na.rm = TRUE)
  )

#Rename latStart and lonStart to match the transect data 
dfo_MSP_summarized <- dfo_MSP_summarized %>%
  rename(
    Latitude = latStart,
    Longitude = lonStart
  )

dfo_MSP_summarized$Year <- as.numeric(dfo_MSP_summarized$Year )
dfo_MSP_summarized$Month <- as.numeric(dfo_MSP_summarized$Month )
dfo_MSP_summarized$Day <- as.numeric(dfo_MSP_summarized$Day )

#Left join the cleaned mps_transect with the summarized dfo_MSP, joined by year month day and start and end latitudes
mps_combined <- mps_transect_cleaned %>%
  left_join(dfo_MSP_summarized, by = c("Latitude", "Longitude", "Year", "Month", "Day")) 

#Rename latStart and lonStart to match the transect data 
mps_combined<- mps_combined %>%
  rename(
    latStart = Latitude ,
    lonStart = Longitude 
  )

#Turn HKey into a character variable
mps_combined$HKey <- as.character(mps_combined$HKey)

#This is where we need to turn HKey into unqiue value based on lat, lon, year, month, day
mps_combined <- mps_combined %>%
  group_by(latStart, lonStart, Year, Month, Day) %>%
  mutate(HKey = cur_group_id()) %>%  # Create a unique ID for each group
  ungroup()  

#Replace substrate category values with actual names for categories
#filter out rows where there is no substrate data
#this might change based on new substrate ENV layer ############################### edits needed 
mps_combined <- mps_combined %>% filter(!is.na(DomSubCat))
#loop through each row of data and create a Substrate)Cat column based on the DFO substrate key
for (i in 1:length(mps_combined$HKey)){
  if (mps_combined$DomSubCat[i] == 1){
    if (mps_combined$DomSubSubCat[i] == "a"){
      mps_combined$DomSubstrate_Cat[i] <- "Bedrock"
    }
    if (mps_combined$DomSubSubCat[i] == "b"){
      mps_combined$DomSubstrate_Cat[i] <-"Boulder"
    }
  }
  if (mps_combined$DomSubCat[i] == 2){
    mps_combined$DomSubstrate_Cat[i] <-"Mixed"
  }
  if (mps_combined$DomSubCat[i] == 3){
    if (mps_combined$DomSubSubCat[i] == "a"){
      mps_combined$DomSubstrate_Cat[i] <- "Sand"
    }
    if (mps_combined$DomSubSubCat[i] == "b"){
      mps_combined$DomSubstrate_Cat[i] <- "Mud"
    }
  }
}

#Add a column for min and max depth (this is putting the records into my pre-determined depth bins from the ROV dataset)
#In the BHM dataset the depth for each quadrat represents the tide corrected depth at the end of the transect (shallowest point). 
#Therefore records will be treated as if they were in the lower of two depth bins if on the edge of two
#At the same time add BHM to the beginning of every HKey character
#check the maximum and minimum depth of any record
max(mps_combined$MeanDepth) # it is 10 m
min(mps_combined$MeanDepth) # it is -1.68 m

#filter out any record with no depth data 
mps_combined <- mps_combined %>% filter(!is.na(MeanDepth))

#loop through each row of the dataset and
for (i in 1:length(mps_combined$HKey)){
  mps_combined$HKey[i] <- paste0("MPS", mps_combined$HKey[i])
  if (mps_combined$MeanDepth[i] >= -6 & mps_combined$MeanDepth[i] < -3){
    mps_combined$Max_Depth[i] <- -3
    mps_combined$Min_Depth[i] <- -6
  }
  if (mps_combined$MeanDepth[i] >= -3 & mps_combined$MeanDepth[i] < 0){
    mps_combined$Max_Depth[i] <- 0
    mps_combined$Min_Depth[i] <- -3
  }
  if (mps_combined$MeanDepth[i] >= 0 & mps_combined$MeanDepth[i] < 3){
    mps_combined$Max_Depth[i] <- 3
    mps_combined$Min_Depth[i] <- 0
  }
  if (mps_combined$MeanDepth[i] >= 3 & mps_combined$MeanDepth[i] < 6){
    mps_combined$Max_Depth[i] <- 6
    mps_combined$Min_Depth[i] <- 3
  }
  if (mps_combined$MeanDepth[i] >= 6 & mps_combined$MeanDepth[i] < 9){
    mps_combined$Max_Depth[i] <- 9
    mps_combined$Min_Depth[i] <- 6
  }
  if (mps_combined$MeanDepth[i] >= 9 & mps_combined$MeanDepth[i] < 12){
    mps_combined$Max_Depth[i] <- 12
    mps_combined$Min_Depth[i] <- 9
  }
  if (mps_combined$MeanDepth[i] >= 12 & mps_combined$MeanDepth[i] < 15){
    mps_combined$Max_Depth[i] <- 15
    mps_combined$Min_Depth[i] <- 12
  }
  if (mps_combined$MeanDepth[i] >= 15 & mps_combined$MeanDepth[i] < 18){
    mps_combined$Max_Depth[i] <- 18
    mps_combined$Min_Depth[i] <- 15
  }
  if (mps_combined$MeanDepth[i] >= 18 & mps_combined$MeanDepth[i] < 21){
    mps_combined$Max_Depth[i] <- 21
    mps_combined$Min_Depth[i] <- 18
  }
}
#Add a column for survey type
mps_combined$Survey_Type <- "DFO_MSP"

#Select columns we want to include in the final dataset
mps_combined <- mps_combined %>% dplyr::select(Survey_Type, HKey, latStart, lonStart, lonEnd, latEnd, Year, Month, Day, Max_Quad_Depth, Min_Quad_Depth, DomSubstrate_Cat,MeanDepth, Min_Depth, Max_Depth, RA_RSU, RA_GSU, RA_PSU)

#Rename the columns in the bhm_combined dataset
mps_combined <- mps_combined %>%
  rename(
    rd_rur = RA_RSU,
    rd_pur = RA_PSU,
    rd_gur = RA_GSU
  )

#Rename the categories in relative density columns 
#Change the values in the columns rd_rur, rd_pur, and rd_gur
mps_combined <- mps_combined %>%
  mutate(
    rd_rur = recode(rd_rur, 'N' = 'None', 'F' = 'Few', 'A' = 'Abundant', 'M' = 'Many'),
    rd_pur = recode(rd_pur, 'N' = 'None', 'F' = 'Few', 'A' = 'Abundant', 'M' = 'Many'),
    rd_gur = recode(rd_gur, 'N' = 'None', 'F' = 'Few', 'A' = 'Abundant', 'M' = 'Many')
  )

#Add columns for presence/absence for each species 
mps_combined <- mps_combined %>%
  mutate(
    p_rur = ifelse(rd_rur == 'None', 0, 1),
    p_pur = ifelse(rd_pur == 'None', 0, 1),
    p_gur = ifelse(rd_gur == 'None', 0, 1)
  )

#Add columns 'counts' with NA values 
#Add new columns with NA values to mps_combined and bhm_combined
#Add new columns with NA values to mps_combined and bhm_combined
mps_combined$gur_count <- NA
mps_combined$rur_count <- NA
mps_combined$pur_count <- NA
mps_combined$ad_pur <- NA
mps_combined$ad_gur <- NA
mps_combined$ad_rur <- NA
mps_combined$site <- mps_combined$HKey

bhm_combined$gur_count <- NA
bhm_combined$rur_count <- NA
bhm_combined$pur_count <- NA
bhm_combined$ad_pur <- NA
bhm_combined$ad_gur <- NA
bhm_combined$ad_rur <- NA
bhm_combined$site <- bhm_combined$HKey

#Rename columns in both dataframes
mps_combined <- mps_combined %>%
  rename(
    survey_type = Survey_Type,
    survey_id = HKey,
    start_lat = latStart,
    start_lon = lonStart,
    end_lon = lonEnd,
    end_lat = latEnd,
    month = Month,
    day = Day,
    year = Year,
    substrate = DomSubstrate_Cat,
    depth = MeanDepth,
    min_depth = Min_Depth,
    max_depth = Max_Depth
  )

bhm_combined <- bhm_combined %>%
  rename(
    survey_type = Survey_Type,
    survey_id = HKey,
    start_lat = Start_Lat,
    start_lon = Start_Long,
    end_lon = Shore_Long,
    end_lat = Shore_Lat,
    month = Month,
    day = Day,
    year = Year,
    substrate = DomSubstrate_Cat,
    depth = MeanDepth,
    min_depth = Min_Depth,
    max_depth = Max_Depth
  )

#Remove the specified columns from both dataframes
mps_combined <- mps_combined %>%
  select(-Max_Quad_Depth, -Min_Quad_Depth)

bhm_combined <- bhm_combined %>%
  select(-Max_Quad_Depth, -Min_Quad_Depth)

#COMBINE 

dfo_relative_combined <- rbind(mps_combined, bhm_combined)

#write dataframe as a new file
write.csv(dfo_relative_combined, 
          "./processed_data/DFO_surveys/dfo_relative_transect_summary.csv", 
          row.names = FALSE)


```

#Combining all our external DATA 
```{r}
hakai <- read.csv( "./processed_data/Hakai/hakai_quadrat_summary.csv")
RLS <- read.csv( "./processed_data/RLS/RLS_transect_summary.csv")
DFO_count <- read.csv("./processed_data/DFO_surveys/dfo_quadrat_summary.csv")
DFO_relative <- read.csv("./processed_data/DFO_surveys/dfo_relative_transect_summary.csv")


combined_data <- rbind(hakai, RLS, DFO_count, DFO_relative)
```

#Now add in Baum lab ROV data
```{r}
#need to reformat this data to fit our other data types 

rov_selected <- ROV_counts[, c("survey_type", "survey_id", "site", "start_lat", "start_lon", "end_lat", "end_lon", 
                          "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "occurrence", "Abundance")]

# Update the site column
rov_selected <- rov_selected %>%
  mutate(site = ifelse(site == "Leey’qsun_island", "Leey_island", site)) #special symbol causes issues later

#remove all total records 
rov_selected <- rov_selected[rov_selected$species != "TOT", ]

#change PU = PUR, GU = GUR, and RU = RUR
rov_selected <- rov_selected %>%
  mutate(species = case_when(
    species == "PU" ~ "PUR",
    species == "GU" ~ "GUR",
    species == "RU" ~ "RUR",
    TRUE ~ species  # Retain other values
  ))

```

#Creating present dataset for PUR, GUR, RUR
#make data long
```{r}
# Select the columns you want to keep
combined_presence <- combined_data %>%
  select(survey_type, survey_id, site, start_lat, start_lon, end_lat, end_lon, 
         year, month, day, depth, min_depth, max_depth, substrate, 
         p_gur, p_rur, p_pur)

# Pivot to long format
long_presence <- combined_presence %>%
  pivot_longer(cols = c(p_gur, p_rur, p_pur), 
               names_to = "species", 
               values_to = "occurrence", 
               names_prefix = "p_") %>%
  mutate(species = recode(species, gur = "GUR", rur = "RUR", pur = "PUR"))

#select relevant columns from rov data 
rov_selected1 <- rov_selected[, c("survey_type", "survey_id", "site", "start_lat", "start_lon", "end_lat", "end_lon", 
                          "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "occurrence")]
#bind 
presence <- rbind(rov_selected1, long_presence)

#Remove NA values in occurence column 
presence <- presence %>%
  filter(!is.na(occurrence))


#save file of ALL (both absence and presence)
write.csv(presence, 
          "./processed_data/Survey_Datasets/presence_absence.csv", 
          row.names = FALSE)


#save file that is ONLY presence 
presence_only <- presence[presence$occurrence == 1, ]

write.csv(presence_only, 
          "./processed_data/Survey_Datasets/presence_only.csv", 
          row.names = FALSE)



```

#Creating count actual density datase for PUR, GUR, PUR 

```{r}

#Something we REALLY need to think about here is how we standardize actual density across datasets
#Because the counts are divided by different areas. 
#As it stands right now, we divide count/area (so they should all be within m^2 values)

#Select the columns you want to keep
combined_actual <- combined_data %>%
  select(survey_type, survey_id, site, start_lat, start_lon, end_lat, end_lon, 
         year, month, day, depth, min_depth, max_depth, substrate, 
         ad_gur, ad_rur, ad_pur)

#remove NA values in actual density 
combined_actual <- combined_actual[complete.cases(combined_actual[, c("ad_gur", "ad_pur", "ad_rur")]), ]


#Pivot to long format
long_actual <- combined_actual %>%
  pivot_longer(cols = c(ad_gur, ad_rur, ad_pur), 
               names_to = "species", 
               values_to = "density_m2", 
               names_prefix = "ad_") %>%
  mutate(species = recode(species, gur = "GUR", rur = "RUR", pur = "PUR"))

#filter out NA values in occurence column 
long_actual <- long_actual %>%
  filter(!is.na(density_m2))

write.csv(long_actual, 
          "./processed_data/Survey_Datasets/actual_density.csv", 
          row.names = FALSE)


```

#Creating relative density data for PUR, GUR, PUR 
#this data only makes sense for ROV and DFO transects, the others are NOT comparable 

```{r}

#combine dfo_relative combine with baum ONLY 
# Select the columns you want to keep
dfo_relative_selected <- DFO_relative %>%
  select(survey_type, survey_id, site, start_lat, start_lon, end_lat, end_lon, 
         year, month, day, depth, min_depth, max_depth, substrate, 
         rd_gur, rd_rur, rd_pur)

#Pivot to long format
long_relative <- dfo_relative_selected %>%
  pivot_longer(cols = c(rd_gur, rd_rur, rd_pur), 
               names_to = "species", 
               values_to = "Abundance", 
               names_prefix = "rd_") %>%
  mutate(species = recode(species, gur = "GUR", rur = "RUR", pur = "PUR"))

#select relevant columns from rov data 
rov_selected2 <- rov_selected[, c("survey_type", "survey_id", "site", "start_lat", "start_lon", "end_lat", "end_lon", 
                          "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "Abundance")]


#bind 
relative <- rbind(rov_selected2, long_relative)

#save
write.csv(relative, 
          "./processed_data/Survey_Datasets/relative_density.csv", 
          row.names = FALSE)


```

