---
title: "Generating_Occurrence_Lines"
author: "Tessa Rehill"
date: "2025-01-23"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
```

#Load libraries
```{r}
library(dplyr)
library(tidyverse)
library(terra)
library(sf)
library(ggspatial)
library(cowplot)
library(patchwork)
```

#Load data
```{r}
Urchin_Presence_Final <- read.csv("./processed_data/Survey_Datasets/presence_only.csv")
```

#Change start/end coordinates 
```{r}
#Loop through the urchin presence dataframe. Wherever there is an NA for end coords, put in the start coords
#Also add a column for record number
for (i in 1:length(Urchin_Presence_Final$survey_type)){
  if (is.na(Urchin_Presence_Final$end_lat[i])){
    Urchin_Presence_Final$end_lat[i] <- Urchin_Presence_Final$start_lat[i]
  }
  if (is.na(Urchin_Presence_Final$end_lon[i])){
    Urchin_Presence_Final$end_lon[i] <- Urchin_Presence_Final$start_lon[i]
  }
  Urchin_Presence_Final$RecordNum[i] <- i 
}

#Loop back through and ensure no end and start lat and long are the exact same
for (i in 1:length(Urchin_Presence_Final$site)){
  if (Urchin_Presence_Final$start_lat[i] == Urchin_Presence_Final$end_lat[i]){
    Urchin_Presence_Final$end_lat[i] <- Urchin_Presence_Final$end_lat[i] + 0.000001
  }
  if (Urchin_Presence_Final$start_lon[i] == Urchin_Presence_Final$end_lon[i]){
    Urchin_Presence_Final$end_lon[i] <- Urchin_Presence_Final$end_lon[i] + 0.000001
  }
}

```

# Plot the depth distribution of urchin records from the datasets
This is to done to get an idea of generally how deep these species are occurring to validate the cutoffs that we selected when clamping our bathymetry during predictor layer creation.
```{r}
#Load in required data
BC_Mask <- terra::vect("./raw_data/BC_Buffer/BC_Buffer.shp") %>% terra::project("epsg:3005")

#Trim to the surveys used in the BC region
Urchin_Presence_Final_Points <- sf::st_as_sf(Urchin_Presence_Final, coords = c("start_lon", "start_lat")) %>% sf::st_set_crs(4326)
Urchin_Presence_Final_Vect <- terra::vect(Urchin_Presence_Final_Points) %>% terra::project(crs(BC_Mask)) %>% terra::mask(BC_Mask)
Urchin_Presence_Final_df <- terra::as.data.frame(Urchin_Presence_Final_Vect)

#How many records for each species?
Urchin_Record_Nums <- Urchin_Presence_Final_df %>% dplyr::group_by(species) %>% dplyr::summarise(n())

Depth_Plot_presence <- ggplot(Urchin_Presence_Final_df, aes(x = depth)) +
  theme_cowplot() +
  geom_boxplot() +
  xlab("Depth (m)") +
  ylab("Species Presence") +
  geom_vline(xintercept = 18, linetype = "dashed", color = "black", linewidth = 0.75)
  #dashed line at 18m 
Depth_Plot_presence
ggsave2(filename = "./Plots/Depth/urchin_presence_depth_plot.jpeg", plot = Depth_Plot_presence, scale = 1, height = 100, width = 175, units = "mm")
```

# For each species create a dataset that has one record per cell of the study area grid being used

## Create Urchin occurrence lines for each record, extract intersected cells and associate Urchin records with the intersected cell closest in depth to its in situ value.

```{r}
#Load the bathymetry data being used for SDMs
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Create a dataframe of all the Urchin start and end points in long form 
Start_Points <- Urchin_Presence_Final %>% dplyr::select(-end_lat, -end_lon) 
End_Points <- Urchin_Presence_Final %>% dplyr::select(-start_lat, -start_lon)
names(Start_Points) <- c("survey_type", "transect", "site", "lat", "lon", "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "occurrence", "RecordNum")
names(End_Points) <- c("survey_type", "transect", "site", "lat", "lon", "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "occurrence", "RecordNum")

Urchin_df <- rbind(Start_Points, End_Points) #depth only greater than 3
Urchin_df <- Urchin_df %>% arrange(RecordNum) #depth only greater than 3

#Create an sf dataframe of points using lat/long data
Urchin_Points <- Urchin_df %>% sf::st_as_sf(coords = c("lon", "lat")) %>% sf::st_set_crs(4326) 
#Cast this into linestrings grouped by record number
Urchin_Transects <- Urchin_Points %>% group_by(survey_type, site, transect, year, month, day, max_depth, min_depth, depth, substrate, species, RecordNum) %>% summarize(o = sum(occurrence))%>% st_cast("LINESTRING") #
#re-sort Urchin transects by recordnum
Urchin_Transects <- Urchin_Transects %>% arrange(RecordNum)

#write the occurrence lines to a shapefile
sf::st_write(Urchin_Transects, "./processed_data/Urchin_Presence/Urchin_Occurrence_Transects.shp", append = FALSE, overwrite = TRUE)

```

##Associate with single bathymetry cell 
```{r}

#Import the occurrence lines as a SpatVector
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Urchin_Transects <- terra::vect("./processed_data/Urchin_Presence/Urchin_Occurrence_Transects.shp") %>% 
#terra::project(crs(Bathymetry)) #%>% filter(depth >= 0) #Reproject the Lines to the crs of the bathymetry, remove all occurences above Canadian Chart Datum

###
Urchin_Transects <- terra::vect("./processed_data/Urchin_Presence/Urchin_Occurrence_Transects.shp") #read in file

Urchin_Transects <- terra::project(Urchin_Transects, crs(Bathymetry)) #match CRS of bathymetry 

#remove depth >=0  in urchin transects (removing all occurrences above Canadian Chart Datum)
Urchin_Transects <- Urchin_Transects[Urchin_Transects$depth >= 0, ]

###
#Extract the cell numbers for all intersected cells in the bathymetry layer
Extract_cells <- as.data.frame(cells(Bathymetry, Urchin_Transects, touches = TRUE))

#Initialize Columns in the Urchin Transect Data that will be used in the following loop
Urchin_Transects <- sf::st_as_sf(Urchin_Transects)
Urchin_Transects$CellNum <- 0 
Urchin_Transects$CellDepth <- 1000


#Loop through all records and compare to each cell depth they intersect to select a best cell to represent each presence point
for (i in 1:length(Urchin_Transects$srvy_ty)){
  Records <- Extract_cells %>% filter(ID == i)
  actual_depth <- as.numeric(Urchin_Transects$depth[i]) 
  for (j in 1:length(Records$ID)){
    cell_depth <- as.numeric(-Bathymetry$WEST_COAST_DEM2[Records[j,2]])
    if (!is.na(cell_depth)){
      if (j == 1){
        Urchin_Transects$CellNum[i] <- Records[j,2]
        Urchin_Transects$CellDepth[i] <- cell_depth
      }
      if (abs(actual_depth-cell_depth)<abs(actual_depth-as.numeric(Urchin_Transects$CellDepth[i]))){
        Urchin_Transects$CellNum[i] <- Records[j,2]
        Urchin_Transects$CellDepth[i] <- cell_depth
      }
    }
  }
}


#Filter out the occurrences that have zeros in the CellNum column still as this means they aren't within any cells
Urchin_Transects_Filtered <- sf::st_as_sf(Urchin_Transects) %>% filter(CellNum != 0)
Urchin_Transects_Filtered_Vect <- terra::vect(Urchin_Transects_Filtered)
terra::writeVector(Urchin_Transects_Filtered_Vect, "./processed_data/Urchin_Presence/Urchin_Occurrence_Transects_Filtered.shp", overwrite = TRUE)

summary(Urchin_Transects_Filtered)

```

#Look at dataset 
```{r}
# Read in the shapefile as a SpatVector
Urchin_Transects_Filtered_Vect <- vect("./processed_data/Urchin_Presence/Urchin_Occurrence_Transects_Filtered.shp")

# Convert to a dataframe
Urchin_Transects_df <- as.data.frame(Urchin_Transects_Filtered_Vect)


summary_table <- Urchin_Transects_df %>%
  group_by(species, srvy_ty) %>%
  summarise(n = n(), .groups = "drop")

print(summary_table)

#write.csv(summary_table, "./processed_data/Survey_Datasets/Presence_SurveyType_Summary.csv", row.names = FALSE)

summary_table <- Urchin_Transects_df %>%
  group_by(year, srvy_ty) %>%
  summarise(n = n(), .groups = "drop")


```


#Generating Urchin Counts ###
#Load data
```{r}
Urchin_Count_Final <- read.csv("./processed_data/Survey_Datasets/actual_density.csv")
```

#Change start/end coordinates 
```{r}
#Loop through the urchin Count dataframe. Wherever there is an NA for end coords, put in the start coords
#Also add a column for record number
for (i in 1:length(Urchin_Count_Final$survey_type)){
  if (is.na(Urchin_Count_Final$end_lat[i])){
    Urchin_Count_Final$end_lat[i] <- Urchin_Count_Final$start_lat[i]
  }
  if (is.na(Urchin_Count_Final$end_lon[i])){
    Urchin_Count_Final$end_lon[i] <- Urchin_Count_Final$start_lon[i]
  }
  Urchin_Count_Final$RecordNum[i] <- i 
}

#Loop back through and ensure no end and start lat and long are the exact same
for (i in 1:length(Urchin_Count_Final$site)){
  if (Urchin_Count_Final$start_lat[i] == Urchin_Count_Final$end_lat[i]){
    Urchin_Count_Final$end_lat[i] <- Urchin_Count_Final$end_lat[i] + 0.000001
  }
  if (Urchin_Count_Final$start_lon[i] == Urchin_Count_Final$end_lon[i]){
    Urchin_Count_Final$end_lon[i] <- Urchin_Count_Final$end_lon[i] + 0.000001
  }
}

```

# Plot the depth distribution of urchin records from the datasets
This is to done to get an idea of generally how deep these species are occurring to validate the cutoffs that we selected when clamping our bathymetry during predictor layer creation.
```{r}
#Load in required data
BC_Mask <- terra::vect("./raw_data/BC_Buffer/BC_Buffer.shp") %>% terra::project("epsg:3005")

#Trim to the surveys used in the BC region
Urchin_Count_Final_Points <- sf::st_as_sf(Urchin_Count_Final, coords = c("start_lon", "start_lat")) %>% sf::st_set_crs(4326)
Urchin_Count_Final_Vect <- terra::vect(Urchin_Count_Final_Points) %>% terra::project(crs(BC_Mask)) %>% terra::mask(BC_Mask)
Urchin_Count_Final_df <- terra::as.data.frame(Urchin_Count_Final_Vect)

#How many records for each species?
Urchin_Record_Nums <- Urchin_Count_Final_df %>% dplyr::group_by(species) %>% dplyr::summarise(n())

Depth_Plot_count <- ggplot(Urchin_Count_Final_df, aes(x = "", y = depth)) +
  theme_cowplot() +
  geom_boxplot(fill = "white") +
  scale_y_continuous(limits = c(0, 30), expand = expansion(mult = c(0, 0.05))) +
  xlab("Species Density") +
  ylab("Depth (m)") +
  geom_hline(yintercept = 18, linetype = "dashed", color = "black", linewidth = 0.75) +
  coord_flip()

Depth_Plot_count

ggsave2(filename = "./Plots/Depth/count_urchin_depth_lot.jpeg", plot = Depth_Plot_count, scale = 1, height = 100, width = 175, units = "mm")


```

# For each species create a dataset that has one record per cell of the study area grid being used

## Create Urchin occurrence lines for each record, extract intersected cells and associate Urchin records with the intersected cell closest in depth to its in situ value.

```{r}
#Load the bathymetry data being used for SDMs
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Create a dataframe of all the Urchin start and end points in long form 
Start_Points <- Urchin_Count_Final %>% dplyr::select(-end_lat, -end_lon) 
End_Points <- Urchin_Count_Final %>% dplyr::select(-start_lat, -start_lon)
names(Start_Points) <- c("survey_type", "transect", "site", "lat", "lon", "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "density_m2", "RecordNum")
names(End_Points) <- c("survey_type", "transect", "site", "lat", "lon", "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "density_m2", "RecordNum")

Urchin_df <- rbind(Start_Points, End_Points) #depth only greater than 3
Urchin_df <- Urchin_df %>% arrange(RecordNum) #depth only greater than 3

#Create an sf dataframe of points using lat/long data
Urchin_Points <- Urchin_df %>% sf::st_as_sf(coords = c("lon", "lat")) %>% sf::st_set_crs(4326) 

#Cast this into linestrings grouped by record number
Urchin_Transects <- Urchin_Points %>% group_by(survey_type, site, transect, year, month, day, max_depth, min_depth, depth, substrate, species, RecordNum) %>% summarize(density = sum(density_m2))%>% st_cast("LINESTRING") #
#re-sort Urchin transects by recordnum
Urchin_Transects <- Urchin_Transects %>% arrange(RecordNum)

#write the occurrence lines to a shapefile
sf::st_write(Urchin_Transects, "./processed_data/Urchin_Count/Urchin_Count_Transects.shp", append = FALSE, overwrite = TRUE)

```

##Associate with single bathymetry cell 
```{r}

#Import the occurrence lines as a SpatVector
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Urchin_Transects <- terra::vect("./processed_data/Urchin_Count/Urchin_Count_Transects.shp") %>% 
#terra::project(crs(Bathymetry)) %>% filter(depth >= 0) #Reproject the Lines to the crs of the bathymetry, remove all occurences above Canadian Chart Datum

###
Urchin_Transects <- terra::vect("./processed_data/Urchin_Count/Urchin_Count_Transects.shp") #read in file

Urchin_Transects <- terra::project(Urchin_Transects, crs(Bathymetry)) #match CRS of bathymetry 

#remove depth >=0  in urchin transects (removing all occurrences above Canadian Chart Datum)
Urchin_Transects <- Urchin_Transects[Urchin_Transects$depth >= 0, ]

###
#Extract the cell numbers for all intersected cells in the bathymetry layer
Extract_cells <- as.data.frame(cells(Bathymetry, Urchin_Transects, touches = TRUE))

#Initialize Columns in the Urchin Transect Data that will be used in the following loop
Urchin_Transects <- sf::st_as_sf(Urchin_Transects)
Urchin_Transects$CellNum <- 0 
Urchin_Transects$CellDepth <- 1000


#Loop through all records and compare to each cell depth they intersect to select a best cell to represent each Count point
for (i in 1:length(Urchin_Transects$srvy_ty)){
  Records <- Extract_cells %>% filter(ID == i)
  actual_depth <- as.numeric(Urchin_Transects$depth[i]) 
  for (j in 1:length(Records$ID)){
    cell_depth <- as.numeric(-Bathymetry$WEST_COAST_DEM2[Records[j,2]])
    if (!is.na(cell_depth)){
      if (j == 1){
        Urchin_Transects$CellNum[i] <- Records[j,2]
        Urchin_Transects$CellDepth[i] <- cell_depth
      }
      if (abs(actual_depth-cell_depth)<abs(actual_depth-as.numeric(Urchin_Transects$CellDepth[i]))){
        Urchin_Transects$CellNum[i] <- Records[j,2]
        Urchin_Transects$CellDepth[i] <- cell_depth
      }
    }
  }
}


#Filter out the occurrences that have zeros in the CellNum column still as this means they aren't within any cells
Urchin_Transects_Filtered <- sf::st_as_sf(Urchin_Transects) %>% filter(CellNum != 0)
Urchin_Transects_Filtered_Vect <- terra::vect(Urchin_Transects_Filtered)
terra::writeVector(Urchin_Transects_Filtered_Vect, "./processed_data/Urchin_Count/Urchin_Count_Transects_Filtered.shp", overwrite = TRUE)

summary(Urchin_Transects_Filtered_Vect)

```

#Look at dataset 
```{r}
# Read in the shapefile as a SpatVector
Urchin_Transects_Filtered_Vect <- vect("./processed_data/Urchin_Count/Urchin_Count_Transects_Filtered.shp")

# Convert to a dataframe
Urchin_Transects_df <- as.data.frame(Urchin_Transects_Filtered_Vect)


summary_table <- Urchin_Transects_df %>%
  group_by(species, srvy_ty) %>%
  summarise(n = n(), .groups = "drop")

print(summary_table)

#write.csv(summary_table, "./processed_data/Survey_Datasets/Count_SurveyType_Summary.csv", row.names = FALSE)

summary_table <- Urchin_Transects_df %>%
  group_by(year, srvy_ty) %>%
  summarise(n = n(), .groups = "drop")


```

#Generating Urchin Relative Densities ###
#Load data
```{r}
Urchin_RD_Final <- read.csv("./processed_data/Survey_Datasets/relative_density.csv")
```


#Change start/end coordinates 
```{r}
#Loop through the urchin RD dataframe. Wherever there is an NA for end coords, put in the start coords
#Also RDd a column for record number
for (i in 1:length(Urchin_RD_Final$survey_type)){
  if (is.na(Urchin_RD_Final$end_lat[i])){
    Urchin_RD_Final$end_lat[i] <- Urchin_RD_Final$start_lat[i]
  }
  if (is.na(Urchin_RD_Final$end_lon[i])){
    Urchin_RD_Final$end_lon[i] <- Urchin_RD_Final$start_lon[i]
  }
  Urchin_RD_Final$RecordNum[i] <- i 
}

#Loop back through and ensure no end and start lat and long are the exact same
for (i in 1:length(Urchin_RD_Final$site)){
  if (Urchin_RD_Final$start_lat[i] == Urchin_RD_Final$end_lat[i]){
    Urchin_RD_Final$end_lat[i] <- Urchin_RD_Final$end_lat[i] + 0.000001
  }
  if (Urchin_RD_Final$start_lon[i] == Urchin_RD_Final$end_lon[i]){
    Urchin_RD_Final$end_lon[i] <- Urchin_RD_Final$end_lon[i] + 0.000001
  }
}

```

# Plot the depth distribution of urchin records from the datasets
This is to done to get an idea of generally how deep these species are occurring to validate the cutoffs that we selected when clamping our bathymetry during predictor layer creation.
```{r}
#Load in required data
BC_Mask <- terra::vect("./raw_data/BC_Buffer/BC_Buffer.shp") %>% terra::project("epsg:3005")

#Trim to the surveys used in the BC region
Urchin_RD_Final_Points <- sf::st_as_sf(Urchin_RD_Final, coords = c("start_lon", "start_lat")) %>% sf::st_set_crs(4326)
Urchin_RD_Final_Vect <- terra::vect(Urchin_RD_Final_Points) %>% terra::project(crs(BC_Mask)) %>% terra::mask(BC_Mask)
Urchin_RD_Final_df <- terra::as.data.frame(Urchin_RD_Final_Vect)

#How many records for each species?
Urchin_Record_Nums <- Urchin_RD_Final_df %>% dplyr::group_by(species) %>% dplyr::summarise(n())

Depth_Plot_RD<- ggplot(Urchin_RD_Final_df, aes(x = "", y = depth)) +
  theme_cowplot() +
  geom_boxplot(fill = "white") +
  scale_y_continuous(limits = c(0, 30), expand = expansion(mult = c(0, 0.05))) +
  xlab("Species Relative Abundance") +
  ylab("Depth (m)") +
  geom_hline(yintercept = 18, linetype = "dashed", color = "black", linewidth = 0.75) +
  coord_flip()

Depth_Plot_RD

ggsave2(filename = "./Plots/Depth/rd_urchin_depth_lot.jpeg", plot = Depth_Plot_RD, scale = 1, height = 100, width = 175, units = "mm")
```

# For each species create a dataset that has one record per cell of the study area grid being used

## Create Urchin occurrence lines for each record, extract intersected cells and associate Urchin records with the intersected cell closest in depth to its in situ value.

```{r}
#Load the bathymetry data being used for SDMs
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Create a dataframe of all the Urchin start and end points in long form 
Start_Points <- Urchin_RD_Final %>% dplyr::select(-end_lat, -end_lon) 
End_Points <- Urchin_RD_Final %>% dplyr::select(-start_lat, -start_lon)
names(Start_Points) <- c("survey_type", "transect", "site", "lat", "lon", "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "Abundance", "RecordNum")
names(End_Points) <- c("survey_type", "transect", "site", "lat", "lon", "year", "month", "day", "depth", "min_depth", "max_depth", "substrate", "species", "Abundance", "RecordNum")

Urchin_df <- rbind(Start_Points, End_Points) #depth only greater than 3
Urchin_df <- Urchin_df %>% arrange(RecordNum) #depth only greater than 3

#Create an sf dataframe of points using lat/long data
Urchin_Points <- Urchin_df %>% sf::st_as_sf(coords = c("lon", "lat")) %>% sf::st_set_crs(4326) 

# Custom function to get the most common category
get_mode <- function(x) {
  x <- na.omit(x)  # Remove NA values
  if (length(x) == 0) return(NA)  # Return NA if no valid entries
  return(names(sort(table(x), decreasing = TRUE))[1])  # Most common category
}

Urchin_Transects <- Urchin_Points %>%
  group_by(survey_type, site, transect, year, month, day, max_depth, min_depth, depth, substrate, species, RecordNum) %>%
  summarize(density = get_mode(Abundance), .groups = "drop") %>%  # Use custom mode function
  st_cast("LINESTRING")

#re-sort Urchin transects by recordnum
Urchin_Transects <- Urchin_Transects %>% arrange(RecordNum)

#write the occurrence lines to a shapefile
sf::st_write(Urchin_Transects, "./processed_data/Urchin_RDensity/Urchin_RDensity_Transects.shp", append = FALSE, overwrite = TRUE)

```

##Associate with single bathymetry cell 
```{r}

#Import the occurrence lines as a SpatVector
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Urchin_Transects <- terra::vect("./processed_data/Urchin_RDensity/Urchin_RDensity_Transects.shp") %>% 
#terra::project(crs(Bathymetry)) %>% filter(depth >= 0) #Reproject the Lines to the crs of the bathymetry, remove all occurences above Canadian Chart Datum

###
Urchin_Transects <- terra::vect("./processed_data/Urchin_RDensity/Urchin_RDensity_Transects.shp") #read in file

Urchin_Transects <- terra::project(Urchin_Transects, crs(Bathymetry)) #match CRS of bathymetry 

#remove depth >=0  in urchin transects (removing all occurrences above Canadian Chart Datum)
Urchin_Transects <- Urchin_Transects[Urchin_Transects$depth >= 0, ]

###
#Extract the cell numbers for all intersected cells in the bathymetry layer
Extract_cells <- as.data.frame(cells(Bathymetry, Urchin_Transects, touches = TRUE))

#Initialize Columns in the Urchin Transect Data that will be used in the following loop
Urchin_Transects <- sf::st_as_sf(Urchin_Transects)
Urchin_Transects$CellNum <- 0 
Urchin_Transects$CellDepth <- 1000


#Loop through all records and compare to each cell depth they intersect to select a best cell to represent each RD point
for (i in 1:length(Urchin_Transects$srvy_ty)){
  Records <- Extract_cells %>% filter(ID == i)
  actual_depth <- as.numeric(Urchin_Transects$depth[i]) 
  for (j in 1:length(Records$ID)){
    cell_depth <- as.numeric(-Bathymetry$WEST_COAST_DEM2[Records[j,2]])
    if (!is.na(cell_depth)){
      if (j == 1){
        Urchin_Transects$CellNum[i] <- Records[j,2]
        Urchin_Transects$CellDepth[i] <- cell_depth
      }
      if (abs(actual_depth-cell_depth)<abs(actual_depth-as.numeric(Urchin_Transects$CellDepth[i]))){
        Urchin_Transects$CellNum[i] <- Records[j,2]
        Urchin_Transects$CellDepth[i] <- cell_depth
      }
    }
  }
}


#Filter out the occurrences that have zeros in the CellNum column still as this means they aren't within any cells
Urchin_Transects_Filtered <- sf::st_as_sf(Urchin_Transects) %>% filter(CellNum != 0)
Urchin_Transects_Filtered_Vect <- terra::vect(Urchin_Transects_Filtered)
terra::writeVector(Urchin_Transects_Filtered_Vect, "./processed_data/Urchin_RDensity/Urchin_RDensity_Transects_Filtered.shp", overwrite = TRUE)

summary(Urchin_Transects_Filtered_Vect)

```

#Look at dataset 
```{r}
# Read in the shapefile as a SpatVector
Urchin_Transects_Filtered_Vect <- vect("./processed_data/Urchin_RDensity/Urchin_RDensity_Transects_Filtered.shp")

# Convert to a dataframe
Urchin_Transects_df <- as.data.frame(Urchin_Transects_Filtered_Vect)


summary_table <- Urchin_Transects_df %>%
  group_by(species, srvy_ty) %>%
  summarise(n = n(), .groups = "drop")

print(summary_table)

#write.csv(summary_table, "./processed_data/Survey_Datasets/RD_SurveyType_Summary.csv", row.names = FALSE)

summary_table <- Urchin_Transects_df %>%
  group_by(year, srvy_ty) %>%
  summarise(n = n(), .groups = "drop")


```

#Combining depth plots 
```{r}

all_depth <- Depth_Plot_presence / Depth_Plot_count / Depth_Plot_RD 

ggsave2(filename = "./Plots/Depth/all_depth.jpeg", plot = all_depth, scale = 1, height = 200, width = 175, units = "mm")

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
