---
title: "Comparing_predictors"
author: "Tessa Rehill"
date: "2025-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load packages
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(cowplot)
library(terra)
library(sf)
library(ggspatial)
library(ggplot2)
library(cowplot)
library(dplyr)
library(forcats)
library(patchwork)

```

#Predictor Influence ####
##Green urchin 
```{r}
#Read in files ##################
occurrence_PI <- read.csv("./Generated_Data/Predictor_Importance_datasets/GUR_Predictor_Importance.csv")

density_PI <- read.csv("./Generated_Data/Evaluation_datasets/Count/GUR/GURcount_ensemble_variable_importance.csv")

ordinal_PI <- read.csv("./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_ensemble_variable_importance.csv")

#Formating occurrence data #############

Variable_Importance <- occurrence_PI

#Summarize a mean value across all threshold metric for each importance metric
  Threshold_Summary <- Variable_Importance %>% dplyr::group_by(predictors) %>% dplyr::summarise(across(TPR:IMAE, list(mean)))
  colSums(Threshold_Summary[,2:11])
  #Scale each column from TPR to IMAE so all of their values are rescaled between 0 and 100
  Threshold_Summary[,2:11] <- data.frame(lapply(Threshold_Summary[,2:11], function(x){scale(x, center = FALSE, scale = sum(x, na.rm = TRUE)/100)}))
  colSums(Threshold_Summary[,2:11])
  
  #Now pivot the dataframe longer and summarize a variable importance mean, stdev and stderr for each predictor
  Summarized_Importance <- Threshold_Summary %>% pivot_longer(cols = TPR_1:IMAE_1, names_to = "metric", values_to = "value") %>% dplyr::group_by(predictors) %>% dplyr::summarise(Mean_Importance = mean(value, na.rm = TRUE), StDev_Importance = sd(value, na.rm = TRUE), StErr_Importance = stats::sd(value, na.rm = TRUE)/sqrt(n()))
  #Calculate min and max values for error bars using stdev or sterr (I use stdev here)
  Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
  Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance
  
  occurrence_SI <- Summarized_Importance 

  occurrence_SI$type <- c('occurrence')
  #occurrence_SI <- occurrence_SI %>% select(c('Mean_Importance', 'StDev_Importance', 'predictors', 'type', 'min', 'max'))
  
  
#Formatting density data ################
  #Rename
  density_PI <- density_PI %>%
  rename(predictors = expl.var)

Variable_Importance <- density_PI

# Pivot and summarize variable importance by predictor
Summarized_Importance <- Variable_Importance %>%
  pivot_longer(cols = var.imp, names_to = "metric", values_to = "value") %>%
  dplyr::group_by(predictors) %>%
  dplyr::summarise(
    Mean_Importance = mean(value, na.rm = TRUE),
    StDev_Importance = sd(value, na.rm = TRUE),
    StErr_Importance = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Calculate the scaling factor based on the unscaled means
scaling_factor <- sum(Summarized_Importance$Mean_Importance, na.rm = TRUE) / 100

# Apply the same scaling factor to all importance metrics
Summarized_Importance$Mean_Importance <- Summarized_Importance$Mean_Importance / scaling_factor
Summarized_Importance$StDev_Importance <- Summarized_Importance$StDev_Importance / scaling_factor
Summarized_Importance$StErr_Importance <- Summarized_Importance$StErr_Importance / scaling_factor

# Calculate min and max for error bars using scaled standard deviation
Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance

# Check that the new scaled values add to 100
colSums(Summarized_Importance[, "Mean_Importance"])
  

  density_SI <- Summarized_Importance 
    
  density_SI$type <- c('density')

  
#Format ordinal data 
  #Rename
  ordinal_PI <- ordinal_PI %>%
  rename(predictors = expl.var)

Variable_Importance <- ordinal_PI

# Pivot and summarize variable importance by predictor
Summarized_Importance <- Variable_Importance %>%
  pivot_longer(cols = var.imp, names_to = "metric", values_to = "value") %>%
  dplyr::group_by(predictors) %>%
  dplyr::summarise(
    Mean_Importance = mean(value, na.rm = TRUE),
    StDev_Importance = sd(value, na.rm = TRUE),
    StErr_Importance = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Calculate the scaling factor based on the unscaled means
scaling_factor <- sum(Summarized_Importance$Mean_Importance, na.rm = TRUE) / 100

# Apply the same scaling factor to all importance metrics
Summarized_Importance$Mean_Importance <- Summarized_Importance$Mean_Importance / scaling_factor
Summarized_Importance$StDev_Importance <- Summarized_Importance$StDev_Importance / scaling_factor
Summarized_Importance$StErr_Importance <- Summarized_Importance$StErr_Importance / scaling_factor

# Calculate min and max for error bars using scaled standard deviation
Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance

# Check that the new scaled values add to 100
colSums(Summarized_Importance[, "Mean_Importance"])
  

  ordinal_SI <- Summarized_Importance 
    
  ordinal_SI$type <- c('relative abundance')
  

#Bind data together 
  data <- rbind(density_SI, occurrence_SI, ordinal_SI)

  
  
  green_data <- data
  green_data$species <- c('green')
  
  
```


```{r}

#Fix labels
predictor_labels <- c(
  do_10m_min      = "Minimum DO (bottom)",
  do_10m_mean     = "Mean DO (bottom)",
  temp_10m_mean   = "Mean Temperature (bottom)",
  temp_10m_max    = "Maximum Temperature (bottom)",
  temp_s_mean     = "Mean Temperature (surface)",
  temperature_s_max = "Max Temperature (surface)",
  REI             = "Relative Exposure Index (REI)",
  temp_s_min      = "Minimum Temperature (surface)",
  temp_s_max = "Maximum Temperature (surface)",
  salt_10m_mean   = "Mean Salinity (bottom)",
  temp_10m_min    = "Minimum Temperature (bottom)",
  TAlk_10m_mean   = "Mean Alkalinity (bottom)",
  salt_10m_min    = "Minimum Salinity (bottom)",
  Rock            = "Rockiness", 
  Slope = "Slope", 
  Depth = "Depth", 
  Current = "Current"
)

#Apply relabeling before combining with `type`
data <- data %>%
  mutate(type = factor(type, levels = c("occurrence", "density", "relative abundance")),
         predictors_clean = predictor_labels[predictors],
         y_lab = interaction(predictors_clean, type, sep = " - ")) %>%
  arrange(type, desc(Mean_Importance)) %>%
  mutate(y_lab = fct_rev(factor(y_lab, levels = unique(y_lab))))

# Step 3: Plot
Predictor_Plot <- ggplot(data) +
  geom_col(
    aes(x = Mean_Importance, y = y_lab, fill = type),
    width = 0.7
  ) +
  geom_errorbar(
    aes(xmin = min, xmax = max, y = y_lab),
    width = 0.3,
    size = 0.5
  ) +
  scale_fill_manual(
    name = "",
    values = c("occurrence" = "#1b5817", "density" = "#55C667FF", "relative abundance" = "#cfeec9"),
    breaks = c("occurrence", "density", "relative abundance")
  ) +
  scale_y_discrete(labels = function(x) gsub(" - (occurrence|density|relative abundance)", "", x)) + 
  theme_cowplot() +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 26),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 22), 
    legend.position = "bottom",
    legend.text = element_text(size = 20),       
    legend.title = element_text(size = 22)) +
    #legend.justification = c(1, 1) +
  xlab("Predictor Importance (%)") +
  ylab("Predictor Variable")

Predictor_Plot

Predictor_Plot_Green <- Predictor_Plot

#save
ggsave("./Plots/PI_Plots/GUR_Predictor_Plot.png", plot = Predictor_Plot, width = 10, height = 15, dpi = 300)

```

#Class plot
```{r}

data$Predictor_Class <- data$Predictor_Class
data <- data %>%
  mutate(
    Predictor_Class = case_when(
      grepl("temp", predictors) ~ "Temperature",
      grepl("salt", predictors) ~ "Salinity",
      #grepl("PAR", predictors) ~ "Light",
      grepl("Rock", predictors) ~ "Substrate",
      grepl("NH4|NO3", predictors) ~ "Nutrient",
      grepl("Current|REI", predictors) ~ "Current/Exposure",
      grepl("^do", predictors) ~ "Oxygen",
      grepl("TAlk", predictors, ignore.case = TRUE) ~ "Alkalinity",
      grepl("Slope|Depth|Aspect", predictors) ~ "Bathymetric",
      TRUE ~ "None"  # Default case
    )
  )


# Reverse factor levels to get occurrence on top
data$type <- factor(data$type, levels = c("relative abundance", "density", "occurrence"))

type_colors <- c(
  occurrence = "#1b5817",
  density = "#55C667FF",
  "relative abundance" = "#cfeec9"
)

class_type_summary <- data %>%
  group_by(Predictor_Class, type) %>%
  summarise(Predictor_Imp = sum(Mean_Importance), .groups = "drop")


class_type_plot <- ggplot(class_type_summary, 
                          aes(x = Predictor_Imp, 
                              y = reorder(Predictor_Class, Predictor_Imp), 
                              fill = type)) +
  geom_col(position = position_dodge2(preserve = "single", padding = 0), width = 0.7) +
  scale_fill_manual(values = type_colors, 
    breaks = c("occurrence", "density", "relative abundance")) +
  theme_cowplot() +
  guides(x.sec = "axis", y.sec = "axis") +
  theme(axis.line.y.left = element_line(size = 0.75, color = "black"),
        axis.line.y.right = element_line(size = 0.75, color = "black"),
        axis.line.x.top = element_line(size = 0.75, color = "black"),
        axis.line.x.bottom = element_line(size = 0.75, color = "black"),
        axis.ticks.x.top = element_blank(), 
        axis.ticks.y.right = element_blank(),
        axis.text.x.top = element_blank(), 
        axis.text.y.right = element_blank(),
        axis.title = element_text(size = 26), 
        axis.text.x = element_text(size = 22), 
        axis.text.y = element_text(size = 22), 
        legend.position = "top",
    legend.text = element_text(size = 20),       # <-- Increase legend text size
    legend.title = element_text(size = 22), 
    plot.title = element_text(size = 28, face = "bold", hjust = 0.5)
) +
  xlab("Predictor Importance (%)") + 
  ylab(" ") + 
  labs(fill = "  ") + labs(title = "Green urchin")

class_type_plot


PredictorClass_Plot_Green <- class_type_plot

ggsave("./Plots/PI_Plots/GUR_PredictorClass_Plot.png", plot = class_type_plot,  width = 20,        
  height = 8,   dpi = 300)

```

##Red Urchin 
```{r}
#Read in files ##################
occurrence_PI <- read.csv("./Generated_Data/Predictor_Importance_datasets/RUR_Predictor_Importance.csv")

density_PI <- read.csv("./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_ensemble_variable_importance.csv")

ordinal_PI <- read.csv("./Generated_Data/Evaluation_datasets/Ordinal/RUR/RURord_ensemble_variable_importance.csv")

#Formating occurrence data #############

Variable_Importance <- occurrence_PI

#Summarize a mean value across all threshold metric for each importance metric
  Threshold_Summary <- Variable_Importance %>% dplyr::group_by(predictors) %>% dplyr::summarise(across(TPR:IMAE, list(mean)))
  colSums(Threshold_Summary[,2:11])
  #Scale each column from TPR to IMAE so all of their values are rescaled between 0 and 100
  Threshold_Summary[,2:11] <- data.frame(lapply(Threshold_Summary[,2:11], function(x){scale(x, center = FALSE, scale = sum(x, na.rm = TRUE)/100)}))
  colSums(Threshold_Summary[,2:11])
  
  #Now pivot the dataframe longer and summarize a variable importance mean, stdev and stderr for each predictor
  Summarized_Importance <- Threshold_Summary %>% pivot_longer(cols = TPR_1:IMAE_1, names_to = "metric", values_to = "value") %>% dplyr::group_by(predictors) %>% dplyr::summarise(Mean_Importance = mean(value, na.rm = TRUE), StDev_Importance = sd(value, na.rm = TRUE), StErr_Importance = stats::sd(value, na.rm = TRUE)/sqrt(n()))
  #Calculate min and max values for error bars using stdev or sterr (I use stdev here)
  Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
  Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance
  
  occurrence_SI <- Summarized_Importance 

  occurrence_SI$type <- c('occurrence')
  #occurrence_SI <- occurrence_SI %>% select(c('Mean_Importance', 'StDev_Importance', 'predictors', 'type', 'min', 'max'))
  
  
#Formatting density data ################
  #Rename
  density_PI <- density_PI %>%
  rename(predictors = expl.var)

Variable_Importance <- density_PI

# Pivot and summarize variable importance by predictor
Summarized_Importance <- Variable_Importance %>%
  pivot_longer(cols = var.imp, names_to = "metric", values_to = "value") %>%
  dplyr::group_by(predictors) %>%
  dplyr::summarise(
    Mean_Importance = mean(value, na.rm = TRUE),
    StDev_Importance = sd(value, na.rm = TRUE),
    StErr_Importance = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Calculate the scaling factor based on the unscaled means
scaling_factor <- sum(Summarized_Importance$Mean_Importance, na.rm = TRUE) / 100

# Apply the same scaling factor to all importance metrics
Summarized_Importance$Mean_Importance <- Summarized_Importance$Mean_Importance / scaling_factor
Summarized_Importance$StDev_Importance <- Summarized_Importance$StDev_Importance / scaling_factor
Summarized_Importance$StErr_Importance <- Summarized_Importance$StErr_Importance / scaling_factor

# Calculate min and max for error bars using scaled standard deviation
Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance

# Check that the new scaled values add to 100
colSums(Summarized_Importance[, "Mean_Importance"])
  

  density_SI <- Summarized_Importance 
    
  density_SI$type <- c('density')

  
#Format ordinal data 
  #Rename
  ordinal_PI <- ordinal_PI %>%
  rename(predictors = expl.var)

Variable_Importance <- ordinal_PI

# Pivot and summarize variable importance by predictor
Summarized_Importance <- Variable_Importance %>%
  pivot_longer(cols = var.imp, names_to = "metric", values_to = "value") %>%
  dplyr::group_by(predictors) %>%
  dplyr::summarise(
    Mean_Importance = mean(value, na.rm = TRUE),
    StDev_Importance = sd(value, na.rm = TRUE),
    StErr_Importance = sd(value, na.rm = TRUE) / sqrt(n()),
    .groups = "drop"
  )

# Calculate the scaling factor based on the unscaled means
scaling_factor <- sum(Summarized_Importance$Mean_Importance, na.rm = TRUE) / 100

# Apply the same scaling factor to all importance metrics
Summarized_Importance$Mean_Importance <- Summarized_Importance$Mean_Importance / scaling_factor
Summarized_Importance$StDev_Importance <- Summarized_Importance$StDev_Importance / scaling_factor
Summarized_Importance$StErr_Importance <- Summarized_Importance$StErr_Importance / scaling_factor

# Calculate min and max for error bars using scaled standard deviation
Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance

# Check that the new scaled values add to 100
colSums(Summarized_Importance[, "Mean_Importance"])
  

  ordinal_SI <- Summarized_Importance 
    
  ordinal_SI$type <- c('relative abundance')
  

#Bind data together 
  data <- rbind(density_SI, occurrence_SI, ordinal_SI)


  red_data <- data
  red_data$species <- c('red')
  


```
###All predictor plots
```{r}
#Fix labels
predictor_labels <- c(
  do_10m_min      = "Minimum DO (bottom)",
  do_10m_mean     = "Mean DO (bottom)",
  temp_10m_mean   = "Mean Temperature (bottom)",
  temp_10m_max    = "Maximum Temperature (bottom)",
  temp_s_mean     = "Mean Temperature (surface)",
  temperature_s_max = "Max Temperature (surface)",
  REI             = "Relative Exposure Index (REI)",
  temp_s_min      = "Minimum Temperature (surface)",
  temp_s_max = "Maximum Temperature (surface)",
  salt_10m_mean   = "Mean Salinity (bottom)",
  temp_10m_min    = "Minimum Temperature (bottom)",
  TAlk_10m_mean   = "Mean Alkalinity (bottom)",
  salt_10m_min    = "Minimum Salinity (bottom)",
  Rock            = "Rockiness", 
  Slope = "Slope", 
  Depth = "Depth", 
  Current = "Current"
)

#Apply relabeling before combining with `type`
data <- data %>%
  mutate(type = factor(type, levels = c("occurrence", "density", "relative abundance")),
         predictors_clean = predictor_labels[predictors],
         y_lab = interaction(predictors_clean, type, sep = " - ")) %>%
  arrange(type, desc(Mean_Importance)) %>%
  mutate(y_lab = fct_rev(factor(y_lab, levels = unique(y_lab))))

# Step 3: Plot
Predictor_Plot <- ggplot(data) +
  geom_col(
    aes(x = Mean_Importance, y = y_lab, fill = type),
    width = 0.7
  ) +
  geom_errorbar(
    aes(xmin = min, xmax = max, y = y_lab),
    width = 0.3,
    size = 0.5
  ) +
  scale_fill_manual(
    name = "",
    values = c("occurrence" = "#760b08", "density" = "#e6360F", "relative abundance" = "#f8b4b2"),
    breaks = c("occurrence", "density", "relative abundance")
  ) +
  scale_y_discrete(labels = function(x) gsub(" - (occurrence|density|relative abundance)", "", x)) + 
  theme_cowplot() +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 26),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 22), 
    legend.position = "bottom",
    legend.text = element_text(size = 20),       # <-- Increase legend text size
    legend.title = element_text(size = 22)) +
  xlab("Predictor Importance (%)") +
  ylab("")

Predictor_Plot

Predictor_Plot_Red <- Predictor_Plot

#save
ggsave("./Plots/PI_Plots/RUR_Predictor_Plot.png", plot = Predictor_Plot, width = 10, height = 15, dpi = 300)
```

###Class plot
```{r}

data$Predictor_Class <- data$Predictor_Class
data <- data %>%
  mutate(
    Predictor_Class = case_when(
      grepl("temp", predictors) ~ "Temperature",
      grepl("salt", predictors) ~ "Salinity",
      #grepl("PAR", predictors) ~ "Light",
      grepl("Rock", predictors) ~ "Substrate",
      grepl("NH4|NO3", predictors) ~ "Nutrient",
      grepl("Current|REI", predictors) ~ "Current/Exposure",
      grepl("^do", predictors) ~ "Oxygen",
      grepl("TAlk", predictors, ignore.case = TRUE) ~ "Alkalinity",
      grepl("Slope|Depth|Aspect", predictors) ~ "Bathymetric",
      TRUE ~ "None"  # Default case
    )
  )


# Reverse factor levels to get occurrence on top
data$type <- factor(data$type, levels = c("relative abundance", "density", "occurrence"))

type_colors <- c(
  occurrence = "#760b08",
  density = "#e6360F",
  "relative abundance" = "#f8b4b2"
)

class_type_summary <- data %>%
  group_by(Predictor_Class, type) %>%
  summarise(Predictor_Imp = sum(Mean_Importance), .groups = "drop")


class_type_plot <- ggplot(class_type_summary, 
                          aes(x = Predictor_Imp, 
                              y = reorder(Predictor_Class, Predictor_Imp), 
                              fill = type)) +
  geom_col(position = position_dodge2(preserve = "single", padding = 0), width = 0.7) +
  scale_fill_manual(values = type_colors, 
    breaks = c("occurrence", "density", "relative abundance")) +
  theme_cowplot() +
  guides(x.sec = "axis", y.sec = "axis") +
  theme(axis.line.y.left = element_line(size = 0.75, color = "black"),
        axis.line.y.right = element_line(size = 0.75, color = "black"),
        axis.line.x.top = element_line(size = 0.75, color = "black"),
        axis.line.x.bottom = element_line(size = 0.75, color = "black"),
        axis.ticks.x.top = element_blank(), 
        axis.ticks.y.right = element_blank(),
        axis.text.x.top = element_blank(), 
        axis.text.y.right = element_blank(),
        axis.title = element_text(size = 26), 
        axis.text.x = element_text(size = 22), 
        axis.text.y = element_text(size = 22), 
        legend.position = "top",
    legend.text = element_text(size = 20),       # <-- Increase legend text size
    legend.title = element_text(size = 22), 
    plot.title = element_text(size = 28, face = "bold", hjust = 0.5)
) +
  xlab("Predictor Importance (%)") + 
  ylab(" ") + 
  labs(fill = "  ", title = "Red urchin")

class_type_plot


PredictorClass_Plot_Red <- class_type_plot

ggsave("./Plots/PI_Plots/RUR_PredictorClass_Plot.png", plot = class_type_plot,   width = 16,        
  height = 8,   dpi = 300)
```
#Purple Urchin 
```{r}
#Read in files ##################
occurrence_PI <- read.csv("./Generated_Data/Predictor_Importance_datasets/PUR_Predictor_Importance.csv")

#ordinal_PI <- read.csv("./Generated_Data/Evaluation_datasets/Ordinal/PUR/PURord_ensemble_variable_importance.csv")

#Formating occurrence data #############

Variable_Importance <- occurrence_PI

#Summarize a mean value across all threshold metric for each importance metric
  Threshold_Summary <- Variable_Importance %>% dplyr::group_by(predictors) %>% dplyr::summarise(across(TPR:IMAE, list(mean)))
  colSums(Threshold_Summary[,2:11])
  #Scale each column from TPR to IMAE so all of their values are rescaled between 0 and 100
  Threshold_Summary[,2:11] <- data.frame(lapply(Threshold_Summary[,2:11], function(x){scale(x, center = FALSE, scale = sum(x, na.rm = TRUE)/100)}))
  colSums(Threshold_Summary[,2:11])
  
  #Now pivot the dataframe longer and summarize a variable importance mean, stdev and stderr for each predictor
  Summarized_Importance <- Threshold_Summary %>% pivot_longer(cols = TPR_1:IMAE_1, names_to = "metric", values_to = "value") %>% dplyr::group_by(predictors) %>% dplyr::summarise(Mean_Importance = mean(value, na.rm = TRUE), StDev_Importance = sd(value, na.rm = TRUE), StErr_Importance = stats::sd(value, na.rm = TRUE)/sqrt(n()))
  #Calculate min and max values for error bars using stdev or sterr (I use stdev here)
  Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
  Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance
  
  occurrence_SI <- Summarized_Importance 

  occurrence_SI$type <- c('occurrence')
  #occurrence_SI <- occurrence_SI %>% select(c('Mean_Importance', 'StDev_Importance', 'predictors', 'type', 'min', 'max'))
  
  

#Format ordinal data 
  #Rename
  #ordinal_PI <- ordinal_PI %>%
  #rename(predictors = expl.var)

#Variable_Importance <- ordinal_PI

# Pivot and summarize variable importance by predictor
#Summarized_Importance <- Variable_Importance %>%
  #pivot_longer(cols = var.imp, names_to = "metric", values_to = "value") %>%
  #dplyr::group_by(predictors) %>%
  #dplyr::summarise(
   # Mean_Importance = mean(value, na.rm = TRUE),
   # StDev_Importance = sd(value, na.rm = TRUE),
   # StErr_Importance = sd(value, na.rm = TRUE) / sqrt(n()),
   # .groups = "drop"
 # )

# Calculate the scaling factor based on the unscaled means
#scaling_factor <- sum(Summarized_Importance$Mean_Importance, na.rm = TRUE) / 100

# Apply the same scaling factor to all importance metrics
#Summarized_Importance$Mean_Importance <- Summarized_Importance$Mean_Importance / scaling_factor
#Summarized_Importance$StDev_Importance <- Summarized_Importance$StDev_Importance / scaling_factor
#Summarized_Importance$StErr_Importance <- Summarized_Importance$StErr_Importance / scaling_factor

# Calculate min and max for error bars using scaled standard deviation
#Summarized_Importance$min <- Summarized_Importance$Mean_Importance - Summarized_Importance$StDev_Importance
#Summarized_Importance$max <- Summarized_Importance$Mean_Importance + Summarized_Importance$StDev_Importance

# Check that the new scaled values add to 100
#colSums(Summarized_Importance[, "Mean_Importance"])
  

 # ordinal_SI <- Summarized_Importance 
    
 # ordinal_SI$type <- c('relative abundance')
  

#Bind data together 
 # data <- rbind(occurrence_SI, ordinal_SI)


  purple_data <- occurrence_SI
  purple_data$species <- c('purple')
```

###All predictor plots
```{r}
predictor_labels <- c(
  do_10m_min      = "Minimum DO (bottom)",
  do_10m_mean     = "Mean DO (bottom)",
  temp_10m_mean   = "Mean Temperature (bottom)",
  temp_10m_max    = "Maximum Temperature (bottom)",
  temp_s_mean     = "Mean Temperature (surface)",
  temperature_s_max = "Max Temperature (surface)",
  REI             = "Relative Exposure Index (REI)",
  temp_s_min      = "Minimum Temperature (surface)",
  temp_s_max = "Maximum Temperature (surface)",
  salt_10m_mean   = "Mean Salinity (bottom)",
  temp_10m_min    = "Minimum Temperature (bottom)",
  TAlk_10m_mean   = "Mean Alkalinity (bottom)",
  salt_10m_min    = "Minimum Salinity (bottom)",
  Rock            = "Rockiness", 
  Slope = "Slope", 
  Depth = "Depth", 
  Current = "Current"
)

#Apply relabeling before combining with `type`
data <- purple_data %>%
  mutate(type = factor(type, levels = c("occurrence", "relative abundance")),
         predictors_clean = predictor_labels[predictors],
         y_lab = interaction(predictors_clean, type, sep = " - ")) %>%
  arrange(type, desc(Mean_Importance)) %>%
  mutate(y_lab = fct_rev(factor(y_lab, levels = unique(y_lab))))

# Step 3: Plot
Predictor_Plot <- ggplot(data) +
  geom_col(
    aes(x = Mean_Importance, y = y_lab, fill = type),
    width = 0.7
  ) +
  geom_errorbar(
    aes(xmin = min, xmax = max, y = y_lab),
    width = 0.3,
    size = 0.5
  ) +
  scale_fill_manual(
    name = "",
    values = c("occurrence" = "#420b7f", "relative abundance" = "#d8baf9"),
    breaks = c("occurrence", "relative abundance")
  ) +
  scale_y_discrete(labels = function(x) gsub(" - (occurrence|relative abundance)", "", x)) + 
  theme_cowplot() +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 26),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 22), 
    legend.position= "bottom"
  ) +
  xlab("Predictor Importance (%)") +
  ylab("Predictor Variable")

Predictor_Plot

#Remove ordinal
#Apply relabeling before combining with `type`
data <- data %>%
  filter(type == "occurrence") %>%
  mutate(type = factor(type, levels = c("occurrence")),
         predictors_clean = predictor_labels[predictors],
         y_lab = interaction(predictors_clean, type, sep = " - ")) %>%
  arrange(type, desc(Mean_Importance)) %>%
  mutate(y_lab = fct_rev(factor(y_lab, levels = unique(y_lab))))

# Step 3: Plot
Predictor_Plot <- ggplot(data) +
  geom_col(
    aes(x = Mean_Importance, y = y_lab, fill = type),
    width = 0.7
  ) +
  geom_errorbar(
    aes(xmin = min, xmax = max, y = y_lab),
    width = 0.3,
    size = 0.5
  ) +
  scale_fill_manual(
    name = "",
    values = c("occurrence" = "#a503fc"),#420b7f
    breaks = c("occurrence")
  ) +
  scale_y_discrete(labels = function(x) gsub(" - (occurrence)", "", x)) + 
  theme_cowplot() +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 26),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 22), 
    legend.position= "top"
  ) +
  xlab("Predictor Importance (%)") +
  ylab("Predictor Variable")

Predictor_Plot


Predictor_Plot_Purple <- Predictor_Plot

#save
ggsave("./Plots/PI_Plots/PUR_Predictor_Plot.png", plot = Predictor_Plot, width = 15, height = 10, dpi = 300)
```

###Class plot 
```{r}
data$Predictor_Class <- data$Predictor_Class
data <- purple_data %>%
  mutate(
    Predictor_Class = case_when(
      grepl("temp", predictors) ~ "Temperature",
      grepl("salt", predictors) ~ "Salinity",
      #grepl("PAR", predictors) ~ "Light",
      grepl("Rock", predictors) ~ "Substrate",
      grepl("NH4|NO3", predictors) ~ "Nutrient",
      grepl("Current|REI", predictors) ~ "Current/Exposure",
      grepl("^do", predictors) ~ "Oxygen",
      grepl("TAlk", predictors, ignore.case = TRUE) ~ "Alkalinity",
      grepl("Slope|Depth|Aspect", predictors) ~ "Bathymetric",
      TRUE ~ "None"  # Default case
    )
  )


# Reverse factor levels to get occurrence on top
data$type <- factor(data$type, levels = c("relative abundance", "occurrence"))

type_colors <- c(
  occurrence = "#420b7f",
  "relative abundance" = "#d8baf9"
)

class_type_summary <- data %>%
  group_by(Predictor_Class, type) %>%
  summarise(Predictor_Imp = sum(Mean_Importance), .groups = "drop")


class_type_plot <- ggplot(class_type_summary, 
                          aes(x = Predictor_Imp, 
                              y = reorder(Predictor_Class, Predictor_Imp), 
                              fill = type)) +
  geom_col(position = position_dodge2(preserve = "single", padding = 0), width = 0.7) +
  scale_fill_manual(values = type_colors, 
    breaks = c("occurrence", "relative abundance")) +
  theme_cowplot() +
  guides(x.sec = "axis", y.sec = "axis") +
  theme(axis.line.y.left = element_line(size = 0.75, color = "black"),
        axis.line.y.right = element_line(size = 0.75, color = "black"),
        axis.line.x.top = element_line(size = 0.75, color = "black"),
        axis.line.x.bottom = element_line(size = 0.75, color = "black"),
        axis.ticks.x.top = element_blank(), 
        axis.ticks.y.right = element_blank(),
        axis.text.x.top = element_blank(), 
        axis.text.y.right = element_blank(),
        axis.title = element_text(size = 26), 
        axis.text.x = element_text(size = 22), 
        axis.text.y = element_text(size = 22), 
        legend.position = "bottom") +
  xlab("Predictor Importance (%)") + 
  ylab(" ") + 
  labs(fill = "  ", labs(title = "Purple urchin"))

class_type_plot


data$Predictor_Class <- data$Predictor_Class
data <- data %>%
  mutate(
    Predictor_Class = case_when(
      grepl("temp", predictors) ~ "Temperature",
      grepl("salt", predictors) ~ "Salinity",
      #grepl("PAR", predictors) ~ "Light",
      grepl("Rock", predictors) ~ "Substrate",
      grepl("NH4|NO3", predictors) ~ "Nutrient",
      grepl("Current|REI", predictors) ~ "Current/Exposure",
      grepl("^do", predictors) ~ "Oxygen",
      grepl("TAlk", predictors, ignore.case = TRUE) ~ "Alkalinity",
      grepl("Slope|Depth|Aspect", predictors) ~ "Bathymetric",
      TRUE ~ "None"  # Default case
    )
  )


# Reverse factor levels to get occurrence on top
data$type <- factor(data$type, levels = c("relative abundance", "occurrence"))

data<- data %>% 
  filter(type == "occurrence")

type_colors <- c(
  occurrence = "#420b7f"
)

class_type_summary <- data %>%
  group_by(Predictor_Class, type) %>%
  summarise(Predictor_Imp = sum(Mean_Importance), .groups = "drop")


class_type_plot <- ggplot(class_type_summary, 
                          aes(x = Predictor_Imp, 
                              y = reorder(Predictor_Class, Predictor_Imp), 
                              fill = type)) +
  geom_col(position = position_dodge2(preserve = "single", padding = 0), width = 0.7) +
  scale_fill_manual(values = type_colors, 
    breaks = c("occurrence", "density", "relative abundance")) +
  theme_cowplot() +
  guides(x.sec = "axis", y.sec = "axis") +
  theme(axis.line.y.left = element_line(size = 0.75, color = "black"),
        axis.line.y.right = element_line(size = 0.75, color = "black"),
        axis.line.x.top = element_line(size = 0.75, color = "black"),
        axis.line.x.bottom = element_line(size = 0.75, color = "black"),
        axis.ticks.x.top = element_blank(), 
        axis.ticks.y.right = element_blank(),
        axis.text.x.top = element_blank(), 
        axis.text.y.right = element_blank(),
        axis.title = element_text(size = 26), 
        axis.text.x = element_text(size = 22), 
        axis.text.y = element_text(size = 22), 
        legend.position = "top",
    legend.text = element_text(size = 20),       # <-- Increase legend text size
    legend.title = element_text(size = 22), 
    plot.title = element_text(size = 28, face = "bold", hjust = 0.5)
) +
  xlab("Predictor Importance (%)") + 
  ylab(" ") + 
  labs(fill = "  ", title = "Purple urchin")

class_type_plot


PredictorClass_Plot_Purple <- class_type_plot

ggsave("./Plots/PI_Plots/PUR_PredictorClass_Plot.png", plot = class_type_plot,   width = 16,        
  height = 8,   dpi = 300)
```

#Plot em all together
```{r}
#Arrange plots side by side
combined_plot <-PredictorClass_Plot_Purple + PredictorClass_Plot_Green + PredictorClass_Plot_Red +
  plot_layout(ncol = 3, widths = c(1,1,1)) + 
  plot_annotation(tag_levels = 'a')

combined_plot <- combined_plot & 
  theme(
    plot.tag = element_text(
      size = 30, 
      face = "bold", 
      color = "black",   # to check if change is applied visually
      family = "Times",
      hjust = 0, 
      vjust = 1
    )
  )

ggsave("./Plots/PI_Plots/Combined_PredictorClass_Plot.png", plot = combined_plot,   width = 30,        
  height = 10,    dpi = 300)

#Arrange plots side by side
combined_plot <- PredictorClass_Plot_Purple

ggsave("./Plots/PI_Plots/Purple_PredictorClass_Plot.png", plot = combined_plot,   width = 30,        
  height = 10,    dpi = 300)


#Arrange plots side by side
combined_plot <- Predictor_Plot_Green + Predictor_Plot_Red + 
  plot_layout(ncol = 2) +      # 2 plots → 2 columns
  plot_annotation(tag_levels = 'a')  # add tags "a", "b"

combined_plot <- combined_plot & theme(
  plot.tag = element_text(size = 30, face = "bold", family = "Times", color = "black")
)
ggsave("./Plots/PI_Plots/Combined_Predictor_Plot.png", plot = combined_plot,   width = 25,        
  height = 16,    dpi = 300)

#Arrange plots side by side
combined_plot <- Predictor_Plot_Purple

combined_plot <- combined_plot & 
  theme(
    plot.tag = element_text(
      size = 30, 
      face = "bold", 
      color = "black",   # to check if change is applied visually
      family = "Times",
      hjust = 0, 
      vjust = 1
    )
  )

ggsave("./Plots/PI_Plots/Purple_Predictor_Plot.png", plot = combined_plot,   width = 20,        
  height = 16,    dpi = 300)

```


##Plotting all species on the same graph
```{r}

#Formatting data
  #combine all data
  all_data <- rbind(purple_data, green_data, red_data)
  

predictor_labels <- c(
  do_10m_min        = "Minimum DO (bottom)",
  do_10m_mean       = "Mean DO (bottom)",
  temp_10m_mean     = "Mean Temperature (bottom)",
  temp_10m_max      = "Maximum Temperature (bottom)",
  temp_s_mean       = "Mean Temperature (surface)",
  temperature_s_max = "Max Temperature (surface)",
  REI               = "Relative Exposure Index (REI)",
  temp_s_min        = "Minimum Temperature (surface)",
  temp_s_max        = "Maximum Temperature (surface)",
  salt_10m_mean     = "Mean Salinity (bottom)",
  temp_10m_min      = "Minimum Temperature (bottom)",
  TAlk_10m_mean     = "Mean Alkalinity (bottom)",
  salt_10m_min      = "Minimum Salinity (bottom)",
  Rock              = "Rockiness", 
  Slope             = "Slope", 
  Depth             = "Depth", 
  Current           = "Current"
)

# Replace values in the column using the named vector
all_data$predictors <- predictor_labels[as.character(all_data$predictors)]

  
#Plotting all occurrence ########################
  occurrence <- all_data %>% filter(type == "occurrence")

# Ensure 'predictors' is a factor to preserve order in the plot
predictor_order <- occurrence %>%
  dplyr::group_by(predictors) %>%
  dplyr::summarise(total_importance = sum(Mean_Importance, na.rm = TRUE)) %>%
  dplyr::arrange(total_importance)

occurrence$predictors <- factor(occurrence$predictors, levels = predictor_order$predictors)

# Create the plot 
occurrence <- ggplot(occurrence, aes(x = Mean_Importance, 
                     y = predictors, 
                     fill = species)) +
  geom_col(position = position_dodge(width = 0.8)) +  # Side-by-side bars
  geom_errorbar(aes(xmin = Mean_Importance - StDev_Importance,
                    xmax = Mean_Importance + StDev_Importance),
                position = position_dodge(width = 0.9),
                width = 0.3) +
  labs(x = "Mean Importance", y = "Predictor", fill = "Species", title = "Occurrence") +
  theme_cowplot() +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 26),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 10),
    legend.position = "bottom"  # Move legend below the plot
  ) + 
  xlab("Predictor Importance (%)") + 
  ylab("Predictor Variable") + 
  scale_fill_manual(
    name = "Species",
    values = c(
      "red" = "#cd654c",
      "green" = "#30a350",
      "purple" = "#894ccd"), 
      breaks = c("green", "purple", "red")
    )
  
#Plotting all density #################

density <- all_data %>% filter(type == "density")

# Ensure 'predictors' is a factor to preserve order in the plot
predictor_order <- density %>%
  dplyr::group_by(predictors) %>%
  dplyr::summarise(total_importance = sum(Mean_Importance, na.rm = TRUE)) %>%
  dplyr::arrange(total_importance)

density$predictors <- factor(density$predictors, levels = predictor_order$predictors)

# Create the plot
density <- ggplot(density, aes(x = Mean_Importance, 
                     y = predictors, 
                     fill = species)) +
  geom_col(position = position_dodge(width = 0.8)) +  # Side-by-side bars
  geom_errorbar(aes(xmin = Mean_Importance - StDev_Importance,
                    xmax = Mean_Importance + StDev_Importance),
                position = position_dodge(width = 0.9),
                width = 0.3) +
  labs(x = "Mean Importance", y = NULL, fill = "Species", title= "density") +
  theme_cowplot() +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 26),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 10), 
    legend.position = "none"
  ) + 
  xlab("Predictor Importance (%)") + 
  scale_fill_manual(
    name = "Species",
    values = c(
      "red" = "#cd654c",
      "green" = "#30a350"
     # "purple" = "#a503fc"
    )
  )

#Plotting all relative data #############
RA <- all_data %>% filter(type == "relative abundance")

# Ensure 'predictors' is a factor to preserve order in the plot
predictor_order <- RA %>%
  dplyr::group_by(predictors) %>%
  dplyr::summarise(total_importance = sum(Mean_Importance, na.rm = TRUE)) %>%
  dplyr::arrange(total_importance)

RA$predictors <- factor(RA$predictors, levels = predictor_order$predictors)

# Create the plot
relative <- ggplot(RA, aes(x = Mean_Importance, 
                     y = predictors, 
                     fill = species)) +
  geom_col(position = position_dodge(width = 0.8)) +  # Side-by-side bars
  geom_errorbar(aes(xmin = Mean_Importance - StDev_Importance,
                    xmax = Mean_Importance + StDev_Importance),
                position = position_dodge(width = 0.9),
                width = 0.3) +
  labs(x = "Mean Importance", y = NULL, fill = "Species", title = "Relative Abundance") +
  theme_cowplot() +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 26),
    axis.text.x = element_text(size = 22),
    axis.text.y = element_text(size = 10), 
    legend.position = "none"
  ) + 
  xlab("Predictor Importance (%)") + 
  scale_fill_manual(
    name = "Species",
    values = c(
      "red" = "#cd654c",
      "green" = "#30a350",
      "purple" = "#894ccd"
    )
  )


# Arrange plots side by side
combined_plot <- occurrence + density + relative + 
  plot_layout(ncol = 3)  # 3 plots in one row

# Display combined plot
print(combined_plot)

ggsave("./Plots/PI_Plots/Combined_Species_PI.png", plot = combined_plot, width = 20, height = 10, dpi = 300)

```

#Comparing Response Curves ####
#Presence 
```{r}
#Load data ###
source("General_Functions.R")

GUR_ES_Preds <- read.csv("./Generated_Data/ES_datasets/GUR_ES_Preds.csv")
RUR_ES_Preds <- read.csv("./Generated_Data/ES_datasets/RUR_ES_Preds.csv")
PUR_ES_Preds <- read.csv("./Generated_Data/ES_datasets/PUR_ES_Preds.csv")
#Set up dataframes for performing the evaluation strip method for any Urchin species 
Hindcast_df <- terra::as.data.frame(terra::rast("./processed_data/ENV_Layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif")) %>% na.omit() 

Hindcast_df <- Hindcast_df[, !names(Hindcast_df) %in% "Depth"]

gc()

#Calculate the means for each environmental variable
Enviro_Means <- colMeans(Hindcast_df)
#Calculate the max and min for each variable
Enviro_Max <- colMax(Hindcast_df)
Enviro_Min <- colMin(Hindcast_df)

#Create a dataframe for prediction to occur onto for each of the predictor variables of interest.
#Set up a base dataframe
Evaluation_Strip_Base <- data.frame()
Evaluation_Strip_Base_Max <- data.frame()
Evaluation_Strip_Base_Min <- data.frame()
for (i in 1:1000){
  Evaluation_Strip_Base[i,1:ncol(Hindcast_df)] <- Enviro_Means
  Evaluation_Strip_Base_Max[i,1:ncol(Hindcast_df)] <- Enviro_Max
  Evaluation_Strip_Base_Min[i,1:ncol(Hindcast_df)] <- Enviro_Min
}
#name the dataframe columns
colnames(Evaluation_Strip_Base) <- colnames(Hindcast_df)
colnames(Evaluation_Strip_Base_Max) <- colnames(Hindcast_df)
colnames(Evaluation_Strip_Base_Min) <- colnames(Hindcast_df)

#Set up a dataframe where values ranging the full environmental conditions for each variable are stored
Evaluation_Strip_Range <- Evaluation_Strip_Base
for (i in 1:ncol(Evaluation_Strip_Range)){
  Evaluation_Strip_Range[,i] <- seq(from = Enviro_Min[i], to = Enviro_Max[i], length.out = 1000)
}

#Load in hindcast climatology #CODEEDIT select for only the variables I want included in my models
Hindcast_Climatology <- terra::rast("./processed_data/ENV_layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif")
Hindcast_Climatology <- terra::subset(Hindcast_Climatology, names(Hindcast_Climatology) != "Depth")


#Running for a single species
Predictor_Names <- c("Talkbmax (mEq/m³)", "Sbmean (PSU)", "Sbmin (PSU)","Tsmean (\u00B0C)", "Tsmax (\u00B0C)", "Tsmin (\u00B0C)", "Tbmean (\u00B0C)", "Tbmax (\u00B0C)", "Tbmin (\u00B0C)", "Oxybmean (mmol/m^3)", "Oxybmin (mmol/m^3)", "Current (m/s)", "REI", "Percent Rock (%)", "Slope (deg)") #CODEEDIT: change this based on the predictors being used
Predictor_IDs <- names(Hindcast_Climatology)

#Plotting
#Plotting function
ES_CompareSpeciesPlotter <- function(GUR_ES, RUR_ES, PUR_ES, var_id, var_name){
  
  get_summary <- function(ES_Preds){
    ES_Preds %>%
      dplyr::select(all_of(var_id), ID, Hold, Model) %>%
      group_by(ID, Model) %>%
      summarise(Mean_Response = mean(.data[[var_id]]), .groups = "drop") %>%
      group_by(ID) %>%
      summarise(Mean = mean(Mean_Response), .groups = "drop") %>%
      cbind(Evaluation_Strip_Range)
  }
  
  # Generate summaries for each species
  GUR_df <- get_summary(GUR_ES) %>% mutate(Species = "GUR")
  RUR_df <- get_summary(RUR_ES) %>% mutate(Species = "RUR")
  PUR_df <- get_summary(PUR_ES) %>% mutate(Species = "PUR")
  
  # Combine all
  Combined_ES <- bind_rows(GUR_df, RUR_df, PUR_df)
  
  # Generate the plot
  p <- ggplot(Combined_ES, aes(x = .data[[var_id]], y = Mean, color = Species)) +
    geom_smooth(method = "loess", se = TRUE, linewidth = 1.2, alpha = 0.2) +
    scale_color_manual(
      values = c("RUR" = "#e6360F", "GUR" = "#55C667FF", "PUR" = "#a503fc"),
      labels = c("RUR" = "red", "GUR" = "green", "PUR" = "purple")
    ) +
    theme_cowplot() +
    labs(x = var_name, y = "LO", color = "Species") +
    theme(
      axis.title = element_text(size = 18),
      axis.text = element_text(size = 16)
    )
  
  return(p)
}

#Generate individual plots
Combined_ES_Plots <- list()
for(i in seq_along(Predictor_IDs)){
  plot_i <- ES_CompareSpeciesPlotter(GUR_ES_Preds, RUR_ES_Preds, PUR_ES_Preds,
                                     Predictor_IDs[i], Predictor_Names[i])
  Combined_ES_Plots[[i]] <- plot_i
  
  # Save individual plots
  ggsave2(filename = paste0("./Plots/ES_Plots/Combined/ES_Compare_", Predictor_IDs[i], ".jpeg"),
          plot = plot_i, scale = 1, height = 150, width = 225, units = "mm")
}

#Extract shared first legend
shared_legend <- get_legend(
  Combined_ES_Plots[[1]] +
    theme(
      legend.position = "right",
      legend.text = element_text(size = 18),
      legend.title = element_text(size = 20)
    )
)

#Remove legends
Combined_ES_Plots_no_legend <- lapply(Combined_ES_Plots, function(p) p + theme(legend.position = "none"))

#Combine plots and legend in grid
Combined_ES_Grid <- plot_grid(
  plot_grid(plotlist = Combined_ES_Plots_no_legend, ncol = 5),
  shared_legend,
  rel_widths = c(1, 0.1)
)

#Save plots
ggsave2(filename = "./Plots/ES_Plots/Combined_ES_Grid.jpeg",
        plot = Combined_ES_Grid, scale = 1, height = 450, width = 520, units = "mm")

```


#density
```{r}
# Load both datasets
density_GUR <- read.csv("./Generated_Data/ES_datasets/GUR_ES_Count.csv") %>%
  mutate(species = "GUR")
density_RUR <- read.csv("./Generated_Data/ES_datasets/RUR_ES_Count.csv") %>%
  mutate(species = "RUR")

# Combine datasets
ggdat <- bind_rows(density_GUR, density_RUR)
ggdat$id <- as.factor(ggdat$id)

# Recode variable names
recode_vars <- function(df) {
  df %>%
    mutate(
      expl.var.factor = recode(expl.var.factor,
                               "Current" = "Current (m/s)",
                               "Slope" = "Slope (deg)",
                               "Rock" = "Percent Rock (%)",
                               "TAlk_10m_mean" = "Talkbmax (mEq/m³)",
                               "salt_10m_min" = "Sbmin (PSU)",
                               "temp_s_max" = "Tsmax (°C)",
                               "temp_s_min" = "Tsmin (°C)",
                               "temp_10m_mean" = "Tbmean (°C)",
                               "temp_10m_min" = "Tbmin (°C)",
                               "do_10m_mean" = "Oxybmean (mmol/m^3)",
                               "do_10m_min" = "Oxybmin (mmol/m^3)",
                               "Depth" = "Depth (m)"
      ),
      expl.name = expl.var.factor
    )
}

ggdat <- recode_vars(ggdat)


# Define oxygen variables
oxy_vars <- c("Oxybmean (mmol/m^3)", "Oxybmin (mmol/m^3)")

# Subset data
ggdat_oxy <- ggdat %>% filter(expl.name %in% oxy_vars)
ggdat_other <- ggdat %>% filter(!expl.name %in% oxy_vars)

# Plot oxygen variables with y limits -200 to 200
p_oxy <- ggplot(ggdat_oxy, aes(x = expl.val, y = pred.val, color = species)) +
  geom_smooth(
    aes(group = species),
    method = "loess",
    se = TRUE,
    fill = "grey30",
    size = 1.2,
    alpha = 0.3
  ) +
  facet_wrap(~ expl.name, scales = "free_x") +
  scale_color_manual(values = c("GUR" = "#55C667FF", "RUR" = "#e6360F")) +
  coord_cartesian(ylim = c(-10, 50)) +
  labs(
    x = "",
    y = "Urchin density (m²)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    panel.grid = element_blank()
  )

# Plot other variables with y limits -4 to 10
p_other <- ggplot(ggdat_other, aes(x = expl.val, y = pred.val, color = species)) +
  geom_smooth(
    aes(group = species),
    method = "loess",
    se = TRUE,
    fill = "grey30",
    size = 1.2,
    alpha = 0.3
  ) +
  facet_wrap(~ expl.name, scales = "free_x") +
  scale_color_manual(values = c("GUR" = "#55C667FF", "RUR" = "#e6360F")) +
  coord_cartesian(ylim = c(-4, 10)) +
  labs(
    x = "Explanatory Variable",
    y = "Urchin density (m²)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    panel.grid = element_blank()
  )

# Combine with patchwork, sharing the legend and stacking vertically
combined_plot <- p_oxy / p_other + plot_layout(guides = "collect")

# Print combined plot
print(combined_plot)


ggsave("./Plots/ES_Plots/GUR_RUR_ES_Count_Combined.png", plot = combined_plot, width = 8, height = 10, dpi = 300)

```


#Ordinal 
```{r}
# Load both datasets
ord_GUR <- read.csv("./Generated_Data/ES_datasets/GUR_ES_Ordinal.csv") %>%
  mutate(species = "GUR")
ord_RUR <- read.csv("./Generated_Data/ES_datasets/RUR_ES_Ordinal.csv") %>%
  mutate(species = "RUR")

# Combine datasets
ggdat <- bind_rows(ord_GUR, ord_RUR)
ggdat$id <- as.factor(ggdat$id)

# Recode variable names
recode_vars <- function(df) {
  df %>%
    mutate(
      expl.var.factor = recode(expl.var.factor,
                               "Current" = "Current (m/s)",
                               "Slope" = "Slope (deg)",
                               "Rock" = "Percent Rock",
                               "TAlk_10m_mean" = "Talkbmax (mEq/m³)",
                               "salt_10m_min" = "Sbmin (PSU)",
                               "temp_s_max" = "Tsmax (°C)",
                               "temp_s_min" = "Tsmin (°C)",
                               "temp_10m_mean" = "Tbmean (°C)",
                               "temp_10m_min" = "Tbmin (°C)",
                               "do_10m_mean" = "Oxybmean (mmol/m^3)",
                               "do_10m_min" = "Oxybmin (mmol/m^3)",
                               "Depth" = "Depth (m)"
      ),
      expl.name = expl.var.factor
    )
}

ggdat <- recode_vars(ggdat)


# Define oxygen variables
oxy_vars <- c("Oxybmean (mmol/m^3)", "Oxybmin (mmol/m^3 )")

# Subset data
ggdat_oxy <- ggdat %>% filter(expl.name %in% oxy_vars)
ggdat_other <- ggdat %>% filter(!expl.name %in% oxy_vars)

# Plot oxygen variables with y limits -200 to 200
p_oxy <- ggplot(ggdat_oxy, aes(x = expl.val, y = pred.val, color = species)) +
  geom_smooth(
    aes(group = species),
    method = "loess",
    se = TRUE,
    fill = "grey30",
    size = 1.2,
    alpha = 0.3
  ) +
  facet_wrap(~ expl.name, scales = "free_x") +
  scale_color_manual(values = c("GUR" = "#55C667FF", "RUR" = "#e6360F")) +
  coord_cartesian(ylim = c(-200, 200)) +
  labs(
    x = "",
    y = "Relative Abundance"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    panel.grid = element_blank()
  )

# Plot other variables with y limits -4 to 10
p_other <- ggplot(ggdat_other, aes(x = expl.val, y = pred.val, color = species)) +
  geom_smooth(
    aes(group = species),
    method = "loess",
    se = TRUE,
    fill = "grey30",
    size = 1.2,
    alpha = 0.3
  ) +
  facet_wrap(~ expl.name, scales = "free_x") +
  scale_color_manual(values = c("GUR" = "#55C667FF", "RUR" = "#e6360F")) +
  coord_cartesian(ylim = c(-4, 10)) +
  labs(
    x = "Explanatory Variable",
    y = "Urchin Relative Abundance"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    panel.grid = element_blank()
  )

# Combine with patchwork, sharing the legend and stacking vertically
combined_plot <- p_oxy / p_other + plot_layout(guides = "collect")

# Print combined plot
print(combined_plot)


```
#all on same graph
```{r}
plot <- ggplot(ggdat, aes(x = expl.val, y = pred.val, color = species)) +
  geom_smooth(
    aes(group = species),
    method = "loess",
    se = TRUE,
    fill = "grey30",
    size = 1.2,
    alpha = 0.3
  ) +
  facet_wrap(~ expl.name, scales = "free_x") +
  scale_color_manual(values = c("GUR" = "#55C667FF", "RUR" = "#e6360F")) +
  #coord_cartesian(ylim = c(1, 4)) +
  labs(
    x = "",
    y = "Urchin Relative Abundance"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "none",
    strip.text = element_text(face = "bold"),
    panel.grid = element_blank()
  )

plot

ggsave("./Plots/ES_Plots/GUR_RUR_ES_Ordinal_Combined.png", plot = plot, width = 10, height = 15, dpi = 300)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
