---
title: "7_Comparing_Distributions"
author: "Tessa Rehill"
date: "2025-05-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load libraries 
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(cowplot)
library(terra)
library(sf)
library(ggspatial)
library(ggplot2)
library(cowplot)
library(dplyr)
library(forcats)
library(patchwork)
library(grid)
```

#Load land file 
```{r}
#geographic layers
Land <- terra::vect("./raw_data/BC_Buffer/Land.shp") 
 # Convert land to sf
  land_sf <- st_as_sf(Land)
  
  Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")
```

#Rotated map function 
```{r}
rotate_sf <- function(sf_obj, angle_deg, center = NULL) {
  angle_rad <- angle_deg * pi / 180
  
  #get centroid if center not provided
  if (is.null(center)) {
    center <- st_centroid(st_union(sf_obj))
    center_coords <- st_coordinates(center)
  } else {
    center_coords <- center
  }
  
  #apply rotation on each geometry separately
  rotated_geoms <- lapply(st_geometry(sf_obj), function(geom) {
    # Extract coordinates depending on geometry type
    coords_list <- list()
    
    #handle POLYGON or MULTIPOLYGON
    if (inherits(geom, "POLYGON")) {
      coords_list <- list(geom)
    } else if (inherits(geom, "MULTIPOLYGON")) {
      coords_list <- geom
    } else {
      stop("Unsupported geometry type")
    }
    
    #rotate each polygon (rings)
    rotated_polygons <- lapply(coords_list, function(polygon) {
      rotated_rings <- lapply(polygon, function(ring) {
        # matrix of coordinates
        coords <- ring
        
        # translate to origin (center)
        coords_centered <- coords - matrix(center_coords, nrow = nrow(coords), ncol = 2, byrow = TRUE)
        
        # rotation matrix
        rotation_matrix <- matrix(c(cos(angle_rad), -sin(angle_rad),
                                    sin(angle_rad), cos(angle_rad)),
                                  nrow = 2, byrow = TRUE)
        
        # rotate coords
        rotated <- t(rotation_matrix %*% t(coords_centered))
        
        # translate back
        rotated <- rotated + matrix(center_coords, nrow = nrow(rotated), ncol = 2, byrow = TRUE)
        
        # ensure polygon ring is closed: first point == last point
        if (!all(rotated[1,] == rotated[nrow(rotated), ])) {
          rotated <- rbind(rotated, rotated[1, ])
        }
        
        return(rotated)
      })
      return(rotated_rings)
    })
    
    #build new geometry with rotated polygons
    if (inherits(geom, "POLYGON")) {
      st_polygon(rotated_polygons[[1]])
    } else if (inherits(geom, "MULTIPOLYGON")) {
      st_multipolygon(rotated_polygons)
    }
  })
  
  #return sf with new rotated geometries
  new_sf <- sf_obj
  st_geometry(new_sf) <- st_sfc(rotated_geoms, crs = st_crs(sf_obj))
  return(new_sf)
}

rotate_points <- function(points, angle_deg, center) {
  angle_rad <- angle_deg * pi / 180
  coords_centered <- as.matrix(points) - matrix(center, nrow = nrow(points), ncol = 2, byrow = TRUE)
  rotation_matrix <- matrix(c(cos(angle_rad), -sin(angle_rad),
                              sin(angle_rad), cos(angle_rad)),
                            nrow = 2, byrow = TRUE)
  rotated <- t(rotation_matrix %*% t(coords_centered))
  rotated + matrix(center, nrow = nrow(points), ncol = 2, byrow = TRUE)
}


```

#Rotate land file 
```{r}
rotation_angle <- -30

#Rotate land
land_rotated <- rotate_sf(land_sf, rotation_angle)

#Define bounding box corners 
coords <- data.frame(
  x = c(1100000, 950000, 750000, 1300000),
  y = c(950000, 1320000, 230000, 450000)
)
#Rotate bounding box corners around the same centroid
#Use the centroid of land_sf for center of rotation
center <- st_coordinates(st_centroid(st_union(land_sf)))

rotated_bbox_coords <- rotate_points(coords, rotation_angle, center)

xlim <- range(rotated_bbox_coords[,1])
ylim <- range(rotated_bbox_coords[,2])


x_center <- mean(rotated_bbox_coords[,1])
y_range <- range(rotated_bbox_coords[,2])  # keep this unchanged

#Define a new, narrower x-range (e.g., 300 km wide instead of full range)
#Adjust the width to your liking
x_half_width <- 300000  # in meters

xlim <- c(x_center - x_half_width, x_center + x_half_width)
ylim <- y_range  # unchanged

#Shift center slightly left (e.g., by 75,000 m)
x_center_adjusted <- x_center - 75000  # Try 50000, 75000, or 100000 as needed

#keep the same x half-width
xlim <- c(x_center_adjusted - x_half_width, x_center_adjusted + x_half_width)

#plot the rotated land shapefile with rotated bounding box limits
ggplot() +
  geom_sf(data = land_rotated, fill = "grey70", color = "black") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  labs(title = "Land shapefile rotated")


#Rotated NORTH arrows
#determining angle of north arrow
arrow_x <- xlim[1] + 0.05 * diff(xlim)  # 5% in from the left
arrow_y <- ylim[1] + 0.05 * diff(ylim)  # 5% up from the bottom

arrow_length <- 50000  # arrow length in map units (e.g., meters)

#Rotate arrow
angle_deg <- 180 - (90 + (-rotation_angle))  
angle_rad <- angle_deg * pi / 180

#Endpoint of arrow
end_x <- arrow_x + arrow_length * cos(angle_rad)
end_y <- arrow_y + arrow_length * sin(angle_rad)

```

#Presence maps ####
##Load prediction layers, land files and threshold datasets 
```{r}
#LOC predictions 
GUR_BP <-terra::rast("./processed_data/LO_Predictions/GUR_LO_Predictions.tif")
PUR_BP <-terra::rast("./processed_data/LO_Predictions/PUR_LO_Predictions.tif")
RUR_BP <-terra::rast("./processed_data/LO_Predictions/RUR_LO_Predictions.tif")

#threshold datasets 
GUR_Thresholds <- read.csv("./Generated_Data/Threshold_datasets/GUR_Thresholds.csv")
RUR_Thresholds <- read.csv("./Generated_Data/Threshold_datasets/RUR_Thresholds.csv")
PUR_Thresholds <- read.csv("./Generated_Data/Threshold_datasets/PUR_Thresholds.csv")


```

##Green urchin 
```{r}
optimal_TSS_row <- GUR_Thresholds[which.max(GUR_Thresholds$TSS), ]
optimal_TSS_threshold <- optimal_TSS_row$Threshold

#Aggregate the raster based on sum
GUR_Agg <- terra::aggregate(GUR_BP, fact=5, fun="mean")

#Convert the raster to points
GUR_Points <- terra::as.points(GUR_Agg)

#Convert to a dataframe with coordinates and drop absences
GUR_df <- st_as_sf(GUR_Points) %>% dplyr::mutate(x = sf::st_coordinates(.)[,1], y = sf::st_coordinates(.)[,2]) %>% st_drop_geometry() %>%
  filter(meanw > optimal_TSS_threshold)

GUR_Point_Map <- ggplot() +    
  geom_sf(data = land_sf, fill = "grey80", color = NA) + 
  geom_point(data = GUR_df, aes(x = x, y = y, size = meanw, alpha = meanw), color = "#55C667FF") + 
  scale_size_continuous(range = c(.1,1.5), name = "LO") + 
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") + 
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true",
                         pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +    
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +    
  theme_cowplot() +    
  xlab(NULL) +    
  ylab(NULL) +    
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  )

GUR_Point_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_Point_Map.png",   # or another path/filename
  plot = GUR_Point_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```
##Green Urchin - rotated 
```{r}
#Rotate green urchin points using the same rotation parameters
GUR_df_rotated_coords <- rotate_points(
  points = GUR_df[, c("x", "y")], 
  angle_deg = rotation_angle, 
  center = center
)

#Add rotated coordinates back to the dataframe
GUR_df <- GUR_df %>%
  dplyr::mutate(
    x_rot = GUR_df_rotated_coords[,1],
    y_rot = GUR_df_rotated_coords[,2]
  )

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_BP)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_BP, coords_points)
values_points <- extracted$meanw  # raster values at points

GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

#Plot
q <- ggplot() +
  geom_point(data = GUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  # small red squares
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +
  geom_point(data = GUR_df, aes(x = x_rot, y = y_rot, size = meanw, alpha = meanw), color = "#55C667FF") +
  scale_size_continuous(range = c(0.1, 1.5), name = "LO") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave("./Plots/Maps/Cluster_maps/GUR_Point_Map_Rotated.png", plot = q, width = 10, height = 8, dpi = 300)

GUR_Point_Map <- q
```
```{r}
#Plot
q <- ggplot() +
  geom_point(data = GUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  # small red squares
  geom_sf(data = land_rotated, fill = 'white', color = 'black') +
  geom_point(data = GUR_df, aes(x = x_rot, y = y_rot, size = meanw, alpha = meanw), color = "#55C667FF") +
  scale_size_continuous(range = c(0.1, 1.5), name = "LO") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave("./Plots/Maps/Cluster_maps/GUR_Point_Map_Rotated_outline.png", plot = q, width = 10, height = 8, dpi = 300)
```

##Green urchin - continuous LO 
```{r}

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_BP)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_BP, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = GUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "LO", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave(
  filename = "./Plots/Maps/LO_Plots/GUR_LO_Map_Rotated.png",   # or another path/filename
  plot = g,
  width = 6,        
  height = 8,       
  dpi = 300         
)

GUR_Presence_Plot_Continuous  <- g
```

##Purple Urchin 
```{r}
optimal_TSS_row <- PUR_Thresholds[which.max(PUR_Thresholds$TSS), ]
optimal_TSS_threshold <- optimal_TSS_row$Threshold

#Aggregate the raster based on sum
PUR_Agg <- terra::aggregate(PUR_BP, fact=5, fun="mean")
#Convert the raster to points
PUR_Points <- terra::as.points(PUR_Agg)
#Convert to a dataframe with coordinates and drop absences
PUR_df <- st_as_sf(PUR_Points) %>% dplyr::mutate(x = sf::st_coordinates(.)[,1], y = sf::st_coordinates(.)[,2]) %>% st_drop_geometry() %>%
  filter(meanw > optimal_TSS_threshold)


PUR_Point_Map <- ggplot() +    
  geom_sf(data = land_sf, fill = "grey80", color = NA) + 
  geom_point(data = PUR_df, aes(x = x, y = y, size = meanw, alpha = meanw), color = "#a503fc") + 
  scale_size_continuous(range = c(.1,1.5), name = "LO") + 
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") + 
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true",
                         pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +    
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +    
  theme_cowplot() +    
  xlab(NULL) +    
  ylab(NULL) +    
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  )
PUR_Point_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/PUR_Point_Map.png",   # or another path/filename
  plot = PUR_Point_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```
##Purple urchin -rotated 
```{r}
#Rotate green urchin points using the same rotation parameters
PUR_df_rotated_coords <- rotate_points(
  points = PUR_df[, c("x", "y")], 
  angle_deg = rotation_angle, 
  center = center
)

#Add rotated coordinates back to the dataframe
PUR_df <- PUR_df %>%
  dplyr::mutate(
    x_rot = PUR_df_rotated_coords[,1],
    y_rot = PUR_df_rotated_coords[,2]
  )

#Adding buffer, rotated 
PUR_points_full <- terra::as.points(PUR_BP)
coords_points <- terra::crds(PUR_points_full)

extracted <- terra::extract(PUR_BP, coords_points)
values_points <- extracted$meanw  # raster values at points

PUR_df_full <- st_as_sf(PUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

rotated_coords <- rotate_points(
  points = PUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

PUR_df_full <- PUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

#Plot
q <- ggplot() +
  geom_point(data = PUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  # small red squares
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +
  geom_point(data = PUR_df, aes(x = x_rot, y = y_rot, size = meanw, alpha = meanw), color = "#a503fc") +
  scale_size_continuous(range = c(0.1, 1.5), name = "LO") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave("./Plots/Maps/Cluster_maps/PUR_Point_Map_Rotated.png", plot = q, width = 10, height = 8, dpi = 300)

PUR_Point_Map <- q
```

```{r}
#Plot
q <- ggplot() +
  geom_point(data = PUR_df_full, aes(x = x_rot, y = y_rot), color = "grey70", shape = 15, size = 0.3, alpha = 0.1) +  # small red squares
  geom_sf(data = land_rotated, fill = 'white', color = 'black') +
  geom_point(data = PUR_df, aes(x = x_rot, y = y_rot, size = meanw, alpha = meanw), color = "#a503fc") +
  scale_size_continuous(range = c(0.1, 1.5), name = "LO") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

ggsave("./Plots/Maps/Cluster_maps/PUR_Point_Map_Rotated_outline.png", plot = q, width = 10, height = 8, dpi = 300)

```

##Purple urchin - LO continuous
```{r}
#Adding buffer, rotated 
PUR_points_full <- terra::as.points(PUR_BP)
coords_points <- terra::crds(PUR_points_full)

extracted <- terra::extract(PUR_BP, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
PUR_df_full <- st_as_sf(PUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = PUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

PUR_df_full <- PUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = PUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "LO", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave(
  filename = "./Plots/Maps/LO_Plots/PUR_LO_Map_Rotated.png",   # or another path/filename
  plot = g,
  width = 6,        
  height = 8,       
  dpi = 300         
)

PUR_Presence_Plot_Continuous  <- g
```

##Red urchin 
```{r}
optimal_TSS_row <- RUR_Thresholds[which.max(RUR_Thresholds$TSS), ]
optimal_TSS_threshold <- optimal_TSS_row$Threshold

#Aggregate the raster based on sum
RUR_Agg <- terra::aggregate(RUR_BP, fact=5, fun="mean")
#Convert the raster to points
RUR_Points <- terra::as.points(RUR_Agg)
#Convert to a dataframe with coordinates and drop absences
RUR_df <- st_as_sf(RUR_Points) %>% dplyr::mutate(x = sf::st_coordinates(.)[,1], y = sf::st_coordinates(.)[,2]) %>% st_drop_geometry() %>%
  filter(meanw > optimal_TSS_threshold)


RUR_Point_Map <- ggplot() +    
  geom_sf(data = land_sf, fill = "grey80", color = NA) + 
  geom_point(data = RUR_df, aes(x = x, y = y, size = meanw, alpha = meanw), color = "#e6360F") + 
  scale_size_continuous(range = c(.1,1.5), name = "LO") + 
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") + 
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true",
                         pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +    
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +    
  theme_cowplot() +    
  xlab(NULL) +    
  ylab(NULL) +    
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  )
RUR_Point_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_Point_Map.png",   # or another path/filename
  plot = RUR_Point_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)

```
##Red urchin - rotated 
```{r}
#Rotate green urchin points using the same rotation parameters
RUR_df_rotated_coords <- rotate_points(
  points = RUR_df[, c("x", "y")], 
  angle_deg = rotation_angle, 
  center = center
)

#Add rotated coordinates back to the dataframe
RUR_df <- RUR_df %>%
  dplyr::mutate(
    x_rot = RUR_df_rotated_coords[,1],
    y_rot = RUR_df_rotated_coords[,2]
  )

#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_BP)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_BP, coords_points)
values_points <- extracted$meanw  # raster values at points

RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

#Plot
q <- ggplot() +
  geom_point(data = RUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  # small red squares
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +
  geom_point(data = RUR_df, aes(x = x_rot, y = y_rot, size = meanw, alpha = meanw), color = "#e6360F") +
  scale_size_continuous(range = c(0.1, 1.5), name = "LO") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave("./Plots/Maps/Cluster_maps/RUR_Point_Map_Rotated.png", plot = q, width = 10, height = 8, dpi = 300)

RUR_Point_Map <- q
```

```{r}
#Plot
q <- ggplot() +
  geom_point(data = RUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  # small red squares
  geom_sf(data = land_rotated,  fill = 'white', color = 'black') +
  geom_point(data = RUR_df, aes(x = x_rot, y = y_rot, size = meanw, alpha = meanw), color = "#e6360F") +
  scale_size_continuous(range = c(0.1, 1.5), name = "LO") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "LO") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave("./Plots/Maps/Cluster_maps/RUR_Point_Map_Rotated_outline.png", plot = q, width = 10, height = 8, dpi = 300)



```

##Red urchin - LOC continuous
```{r}
#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_BP)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_BP, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = RUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "LO", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave(
  filename = "./Plots/Maps/LO_Plots/RUR_LO_Map_Rotated.png",   # or another path/filename
  plot = g,
  width = 6,        
  height = 8,       
  dpi = 300         
)

RUR_Presence_Plot_Continuous  <- g
```


#Combine presence maps 
```{r}
presence_combined_plot <- PUR_Point_Map + GUR_Point_Map + RUR_Point_Map + 
  plot_layout(ncol = 3)  

presence_combined_plot

ggsave(
  filename = "./Plots/Maps/Cluster_maps/Combined_Presence_Map.png",   # or another path/filename
  plot = presence_combined_plot,
  width = 20,        
  height = 16,       
  dpi = 300         
)

######

presence_combined_plot <- PUR_Presence_Plot_Continuous + GUR_Presence_Plot_Continuous + RUR_Presence_Plot_Continuous + 
  plot_layout(ncol = 3)  

presence_combined_plot

ggsave(
  filename = "./Plots/Maps/LO_Plots/Combined_Presence_Continuous_Map.png",   # or another path/filename
  plot = presence_combined_plot,
  width = 20,        
  height = 16,       
  dpi = 300         
)


```

#LOC Uncertainty Maps ####
##Green urchin - presence uncertainty
```{r}
#read in uncertainty 
GUR_Uncertainty <-terra::rast("./processed_data/Uncertainty/GUR_Method_Uncertainty.tif")

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_Uncertainty)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_Uncertainty, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = GUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Uncertainty", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave("./Plots/Maps/Uncertainty_Plots/GUR_Model_Uncertainty_Map_Rotate.png", plot = g, width = 10, height = 8, dpi = 300)

GUR_Uncertainty_Plot  <- g
```

##Purple urchin - presence uncertainty
```{r}
#read in uncertainty 
PUR_Uncertainty <-terra::rast("./processed_data/Uncertainty/PUR_Method_Uncertainty.tif")

#Adding buffer, rotated 
PUR_points_full <- terra::as.points(PUR_Uncertainty)
coords_points <- terra::crds(PUR_points_full)

extracted <- terra::extract(PUR_Uncertainty, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
PUR_df_full <- st_as_sf(PUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = PUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

PUR_df_full <- PUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = PUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Uncertainty", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave("./Plots/Maps/Uncertainty_Plots/PUR_Model_Uncertainty_Map_Rotate.png", plot = g, width = 10, height = 8, dpi = 300)

PUR_Uncertainty_Plot <- g
```

##Red urchin - presence uncertainty 
```{r}
#read in uncertainty 
RUR_Uncertainty <-terra::rast("./processed_data/Uncertainty/RUR_Method_Uncertainty.tif")

#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_Uncertainty)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_Uncertainty, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = RUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Uncertainty", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave("./Plots/Maps/Uncertainty_Plots/RUR_Model_Uncertainty_Map_Rotate.png", plot = g, width = 10, height = 8, dpi = 300)

RUR_Uncertainty_Plot <- g
```

###Plot presence uncertainties together
```{r}
presence_uncertainty_combined_plot <- PUR_Uncertainty_Plot + GUR_Uncertainty_Plot + RUR_Uncertainty_Plot + 
  plot_layout(ncol = 3)  

presence_uncertainty_combined_plot

ggsave(
  filename = "./Plots/Maps/Uncertainty_Plots/Combined_Presence_Uncertainty_Map.png",   # or another path/filename
  plot = presence_uncertainty_combined_plot,
  width = 20,        
  height = 16,       
  dpi = 300         
)
```

#Count maps ####
##Load prediction layers (rasters) + land files
```{r}

#count predictions
GUR_Count_Raster <- rast("./models_biomod/count_projections/GUR/proj_GUR_current_count_ensembles/proj_GUR_current_count_ensembles_GUR_ensemble.tif")
RUR_Count_Raster <- rast("./models_biomod/count_projections/RUR/proj_RUR_current_count_ensembles/proj_RUR_current_count_ensembles_RUR_ensemble.tif")

#masks (extrapolated variables)
GURclamp <- rast("./models_biomod/count/GUR/proj_GUR_current_count/proj_GUR_current_count_ClampingMask.tif")

RURclamp <- rast("./models_biomod/count/RUR/proj_RUR_current_count/proj_RUR_current_count_ClampingMask.tif")

#Remove cells that extrapolate
GUR_Count_Raster_Mask <- mask(GUR_Count_Raster, GURclamp >= 1, maskvalue = TRUE)
RUR_Count_Raster_Mask <- mask(RUR_Count_Raster, RURclamp >= 1, maskvalue = TRUE)
```

##Abundance map (no aggregations)
```{r}
AC_Map <- function(AC_Raster, Legend_Label, BC_Shapefile) {
  
  # Convert raster to dataframe for ggplot
  AC_df <- as.data.frame(AC_Raster, xy = TRUE, na.rm = TRUE)
  colnames(AC_df)[3] <- "value"  # Rename third column to "value" for clarity

  # Convert land shapefile to sf
  BC_sf <- st_as_sf(BC_Shapefile)

  # Define breaks and labels based on raster values
  min_val <- round(minmax(AC_Raster)[1], 2)
  max_val <- round(minmax(AC_Raster)[2], 2)
  breaks_seq <- round(seq(min_val, max_val, length.out = 5), 2)

  # Plot
  AC_Plot <- ggplot() +
    geom_sf(data = BC_sf, fill = "grey80", color = NA) +
    geom_tile(data = AC_df, aes(x = x, y = y, fill = value)) +
    scale_fill_viridis_c(
      na.value = NA,
      name = Legend_Label,
      limits = c(min_val - 0.01, max_val + 0.01),
      breaks = breaks_seq,
      labels = as.character(breaks_seq)
    ) +
    coord_sf(
      crs = st_crs(3005),
      xlim = c(515000, 1255000),
      ylim = c(340000, 1222000)
    ) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(
      location = "bl", which_north = "true",
      pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
      style = north_arrow_fancy_orienteering
    ) +
    theme_cowplot() +
    labs(x = NULL, y = NULL) +
    theme(
      panel.background = element_rect(fill = "white"),
      legend.position = c(1, 1),
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 20),
      legend.justification = c("right", "top"),
      legend.box.just = "right"
    )
  
  return(AC_Plot)
}

#Green urchin map

GUR_Map <- AC_Map(GUR_Count_Raster, "Green Urchin Abundance", Land)

GUR_Map

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/GUR_Abundance_Map.png",   # or another path/filename
  plot = GUR_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)


#Red urchin map
RUR_Map <- AC_Map(RUR_Count_Raster, "Red Urchin Abundance", Land)

RUR_Map

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/RUR_Abundance_Map.png",   # or another path/filename
  plot = RUR_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)

```

##Green urchin count (continuous) rotated 
```{r}
#read in raster
GUR_Count_Raster_Mask

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_Count_Raster_Mask)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_Count_Raster_Mask, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = GUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Density (m²)", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/GUR_Abundance_Map_Rotated.png",   # or another path/filename
  plot = g,
  width = 6,        
  height = 8,       
  dpi = 300         
)

GUR_Abundance_Plot_Continuous  <- g
```

##Red urchin count (continuous) rotated 
```{r}
#read in raster
RUR_Count_Raster_Mask

#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_Count_Raster_Mask)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_Count_Raster_Mask, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = RUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Density (m²)", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/RUR_Abundance_Map_Rotated.png",   # or another path/filename
  plot = g,
  width = 6,        
  height = 8,       
  dpi = 300         
)

RUR_Abundance_Plot_Continuous  <- g
```

###Combine red and green continuous rotated maps 
```{r}
abundance_continuous_plots<- GUR_Abundance_Plot_Continuous + RUR_Abundance_Plot_Continuous + 
  plot_layout(ncol = 2)  

abundance_continuous_plots

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/Combined_Abundance_Continuous.png",   # or another path/filename
  plot = abundance_continuous_plots,
  width = 20,        
  height = 16,       
  dpi = 300         
)
```


##Green cluster abundance maps 
```{r}

#Import one of my raster prediction layers
GUR_BP <- GUR_Count_Raster_Mask

#Aggregate the raster based on sum
GUR_Agg <- terra::aggregate(GUR_BP, fact=5, fun="mean")

model_name <- c("GUR_EMwmeanByRsquared_mergedData_mergedRun_mergedAlgo")

#Convert the raster to points
GUR_Points <- terra::as.points(GUR_Agg)

#Convert to a dataframe with coordinates and drop absences
GUR_df <- st_as_sf(GUR_Points) %>% dplyr::mutate(x = sf::st_coordinates(.)[,1], y = sf::st_coordinates(.)[,2]) %>% st_drop_geometry() %>%
  filter(GUR_EMwmeanByRsquared_mergedData_mergedRun_mergedAlgo > 0)

GUR_df <- GUR_df %>%
  rename(Likelihood = GUR_EMwmeanByRsquared_mergedData_mergedRun_mergedAlgo)


GUR_Count_Map <- ggplot() +    
  geom_sf(data = land_sf, fill = "grey80", color = NA) + 
  geom_point(data = GUR_df, aes(x = x, y = y, size = Likelihood, alpha = Likelihood), color = "#55C667FF") + 
  scale_size_continuous(range = c(.1, 1.5), name = "Mean Density (m²)") + 
  scale_alpha_continuous(range = c(0.1, 0.5), name = "Mean Density (m²)") + 
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true",
                         pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +    
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +    
  theme_cowplot() +    
  xlab(NULL) +    
  ylab(NULL) +    
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  )

GUR_Count_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_Count_Map.png",   # or another path/filename
  plot = GUR_Count_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```

#Green cluster abundance map - rotated 
```{r}

names(GUR_BP) <- "value"

#Rotate green urchin points using the same rotation parameters
GUR_df_rotated_coords <- rotate_points(
  points = GUR_df[, c("x", "y")], 
  angle_deg = rotation_angle, 
  center = center
)

#Add rotated coordinates back to the dataframe
GUR_df <- GUR_df %>%
  dplyr::mutate(
    x_rot = GUR_df_rotated_coords[,1],
    y_rot = GUR_df_rotated_coords[,2]
  )

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_BP)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_BP, coords_points)
values_points <- extracted$value 


GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

#Plot
q <- ggplot() +
  geom_point(data = GUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +
  geom_point(data = GUR_df, aes(x = x_rot, y = y_rot, size = Likelihood, alpha = Likelihood), color = "#55C667FF") +
  scale_size_continuous(range = c(0.1, 1.5), name = "Mean Density (m²)") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "Mean Density (m²)") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_Count_Map_Rotated.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)

GUR_Count_Map <- q
```

```{r}
#Plot
q <- ggplot() +
  geom_point(data = GUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "white", color = 'black') +
  geom_point(data = GUR_df, aes(x = x_rot, y = y_rot, size = Likelihood, alpha = Likelihood), color = "#55C667FF") +
  scale_size_continuous(range = c(0.1, 1.5), name = "Mean Density (m²)") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "Mean Density (m²)") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_Count_Map_Rotated_outline.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```

##Red cluster abundance map 
```{r}
#Import one of my raster prediction layers
RUR_BP <- RUR_Count_Raster_Mask

#Aggregate the raster based on sum
RUR_Agg <- terra::aggregate(RUR_BP, fact=5, fun="mean")

model_name <- c("RUR_EMwmeanByRsquared_mergedData_mergedRun_mergedAlgo")

#Convert the raster to points
RUR_Points <- terra::as.points(RUR_Agg)

#Convert to a dataframe with coordinates and drop absences
RUR_df <- st_as_sf(RUR_Points) %>% dplyr::mutate(x = sf::st_coordinates(.)[,1], y = sf::st_coordinates(.)[,2]) %>% st_drop_geometry() %>%
  filter(RUR_EMwmeanByRsquared_mergedData_mergedRun_mergedAlgo > 0)

RUR_df <- RUR_df %>%
  rename(Likelihood = RUR_EMwmeanByRsquared_mergedData_mergedRun_mergedAlgo)


RUR_Count_Map <- ggplot() +    
  geom_sf(data = land_sf, fill = "grey80", color = NA) + 
  geom_point(data = RUR_df, aes(x = x, y = y, size = Likelihood, alpha = Likelihood), color = "#e6360F") + 
  scale_size_continuous(range = c(.1, 1.5), name = "Mean Density (m²)") + 
  scale_alpha_continuous(range = c(0.1, 0.5), name = "Mean Density (m²)") + 
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true",
                         pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +    
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +    
  theme_cowplot() +    
  xlab(NULL) +    
  ylab(NULL) +    
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA))  # No box around legend

RUR_Count_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_Count_Map.png",   # or another path/filename
  plot = RUR_Count_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```

#Red urchin count - rotated 
```{r}
names(RUR_BP) <- "value"

#Rotate green urchin points using the same rotation parameters
RUR_df_rotated_coords <- rotate_points(
  points = RUR_df[, c("x", "y")], 
  angle_deg = rotation_angle, 
  center = center
)

#Add rotated coordinates back to the dataframe
RUR_df <- RUR_df %>%
  dplyr::mutate(
    x_rot = RUR_df_rotated_coords[,1],
    y_rot = RUR_df_rotated_coords[,2]
  )

#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_BP)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_BP, coords_points)
values_points <- extracted$value 


RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

#Plot
q <- ggplot() +
  geom_point(data = RUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +
  geom_point(data = RUR_df, aes(x = x_rot, y = y_rot, size = Likelihood, alpha = Likelihood), color = "#e6360F") +
  scale_size_continuous(range = c(0.1, 1.5), name = "Mean Density (m²)") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "Mean Density (m²)") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_Count_Map_Rotated.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)

RUR_Count_Map <- q
```

```{r}
#Plot
q <- ggplot() +
  geom_point(data = RUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "white", color = 'black') +
  geom_point(data = RUR_df, aes(x = x_rot, y = y_rot, size = Likelihood, alpha = Likelihood), color = "#e6360F") +
  scale_size_continuous(range = c(0.1, 1.5), name = "Mean Density (m²)") +
  scale_alpha_continuous(range = c(0.1, 0.5), name = "Mean Density (m²)") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  ) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_Count_Map_Rotated_outline.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```

###Combine count maps 
```{r}
count_combined_plot <- GUR_Count_Map + RUR_Count_Map+ 
  plot_layout(ncol = 2)  

count_combined_plot

ggsave(
  filename = "./Plots/Maps/Cluster_maps/Combined_Count_Map.png",   # or another path/filename
  plot = count_combined_plot,
  width = 20,        
  height = 16,       
  dpi = 300         
)


```

#Uncertainty Count Maps ####
#Green urchin count uncertainty
```{r}
#read in uncertainty 
GUR_Uncertainty <-terra::rast("./processed_data/Uncertainty/GUR_Method_Uncertainty_Count.tif")

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_Uncertainty)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_Uncertainty, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = GUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Uncertainty (m²)", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave("./Plots/Maps/Uncertainty_Plots/GUR_Model_Uncertainty_Map_Count_Rotate.jpeg", plot = g, width = 10, height = 8, dpi = 300)

GUR_Uncertainty_Plot  <- g
```

#Red urchin count uncertainty 
```{r}
#read in uncertainty 
RUR_Uncertainty <-terra::rast("./processed_data/Uncertainty/RUR_Method_Uncertainty_Count.tif")

#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_Uncertainty)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_Uncertainty, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = RUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Uncertainty (m²)", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave("./Plots/Maps/Uncertainty_Plots/RUR_Model_Uncertainty_Map_Count_Rotate.jpeg", plot = g, width = 10, height = 8, dpi = 300)

RUR_Uncertainty_Plot <- g
```

###Combine uncertainty plots 
```{r}
count_uncertainty_combined_plot <- GUR_Uncertainty_Plot + RUR_Uncertainty_Plot + 
  plot_layout(ncol = 2)  

count_uncertainty_combined_plot

ggsave(
  filename = "./Plots/Maps/Uncertainty_Plots/Combined_Count_Uncertainty_Map.png",   # or another path/filename
  plot = count_uncertainty_combined_plot,
  width = 20,        
  height = 16,       
  dpi = 300         
)
```

#Relative abundance maps ####
```{r}
#PREDICTIONS ########

RUR_Ordinal_Raster <- rast("./models_biomod/ordinal/RUR/proj_RUR_current_ord_ensembles/proj_RUR_current_ord_ensembles_RUR_ensemble.tif")

GUR_Ordinal_Raster <- rast("./models_biomod/ordinal/GUR/proj_GUR_current_ord_ensembles/proj_GUR_current_ord_ensembles_GUR_ensemble.tif")


#CLAMPS #########
#masks (extrapolated variables)
#GURclamp <- rast("./models_biomod/ordinal/GUR/proj_GUR_current_ord/proj_GUR_current_ord_ClampingMask.tif")

#RURclamp <- rast("./models_biomod/ordinal/RUR/proj_RUR_current_ord/proj_RUR_current_ord_ClampingMask.tif")

#Remove cells that extrapolate
#GUR_Ordinal_Raster_Mask <- mask(GUR_Ordinal_Raster, GURclamp >= 1, maskvalue = TRUE)
#RUR_Ordinal_Raster_Mask <- mask(RUR_Ordinal_Raster, RURclamp >= 1, maskvalue = TRUE)


##
GUR_Ordinal_Raster_Mask <- GUR_Ordinal_Raster
RUR_Ordinal_Raster_Mask <- RUR_Ordinal_Raster


AC_df_GUR <- as.data.frame(GUR_Ordinal_Raster_Mask, xy = TRUE, na.rm = TRUE)
AC_df_RUR <- as.data.frame(RUR_Ordinal_Raster_Mask, xy = TRUE, na.rm = TRUE)

```

##Continuous 
```{r}
AC_Map <- function(AC_Raster, Legend_Label, BC_Shapefile) {
  # Convert raster to dataframe for ggplot
  AC_df <- as.data.frame(AC_Raster, xy = TRUE, na.rm = TRUE)
  colnames(AC_df)[3] <- "value"  # Rename third column to "value"

  # Convert land shapefile to sf
  BC_sf <- st_as_sf(BC_Shapefile)

  # Convert value to factor with labels
  AC_df$value <- factor(AC_df$value,
                        levels = c(1, 2, 3, 4),
                        labels = c("None", "Few", "Many", "Abundant"))

  # Plot
  AC_Plot <- ggplot() +
    geom_sf(data = BC_sf, fill = "grey80", color = NA) +
    geom_tile(data = AC_df, aes(x = x, y = y, fill = value)) +
    scale_fill_manual(
      values = viridis::viridis(4),
      name = Legend_Label,
      drop = FALSE
    ) +
    coord_sf(
      crs = st_crs(3005),
      xlim = c(515000, 1255000),
      ylim = c(340000, 1222000)
    ) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(
      location = "bl", which_north = "true",
      pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
      style = north_arrow_fancy_orienteering
    ) +
    theme_cowplot() +
    labs(x = NULL, y = NULL) +
    theme(
      panel.background = element_rect(fill = "white"),
      legend.position = c(1, 1),
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 20),
      legend.justification = c("right", "top"),
      legend.box.just = "right"
    )

  return(AC_Plot)
}





#Red urchin map
RUR_Map <- AC_Map(RUR_Ordinal_Raster, "Red Urchin Relative Abundance", Land)

RUR_Map

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/RUR_RelativeAbundance_Map.png",   # or another path/filename
  plot = RUR_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)



```
```{r}
AC_Map <- function(AC_Raster, Legend_Label, BC_Shapefile) {
  # Convert raster to dataframe for ggplot
  AC_df <- as.data.frame(AC_Raster, xy = TRUE, na.rm = TRUE)
  colnames(AC_df)[3] <- "value"  # Rename third column to "value"

  # Convert land shapefile to sf
  BC_sf <- st_as_sf(BC_Shapefile)

  # Convert value to factor with labels
  AC_df$value <- factor(AC_df$value,
                        levels = c(0, 1, 2, 3),
                        labels = c("None", "Few", "Many", "Abundant"))

  # Plot
  AC_Plot <- ggplot() +
    geom_sf(data = BC_sf, fill = "grey80", color = NA) +
    geom_tile(data = AC_df, aes(x = x, y = y, fill = value)) +
    scale_fill_manual(
      values = viridis::viridis(4),
      name = Legend_Label,
      drop = FALSE
    ) +
    coord_sf(
      crs = st_crs(3005),
      xlim = c(515000, 1255000),
      ylim = c(340000, 1222000)
    ) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(
      location = "bl", which_north = "true",
      pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
      style = north_arrow_fancy_orienteering
    ) +
    theme_cowplot() +
    labs(x = NULL, y = NULL) +
    theme(
      panel.background = element_rect(fill = "white"),
      legend.position = c(1, 1),
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 20),
      legend.justification = c("right", "top"),
      legend.box.just = "right"
    )

  return(AC_Plot)
}

#Green urchin map

GUR_Map <- AC_Map(GUR_Ordinal_Raster, "Green Urchin Relative Abundance", Land)
GUR_Map 

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/GUR_RelativeAbundance_Map.png",   # or another path/filename
  plot = GUR_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)

```



#Green continuous - rotated
```{r}
# Read in and convert raster to points
GUR_points_full <- terra::as.points(GUR_Ordinal_Raster_Mask)
coords_points <- terra::crds(GUR_points_full)
extracted <- terra::extract(GUR_Ordinal_Raster_Mask, coords_points)
values_points <- extracted[[1]]  # First and only column

# Build dataframe with coordinates and extracted values
GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[, 1],
    y = sf::st_coordinates(.)[, 2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

# Rotate coordinates
rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[, 1],
    y_rot = rotated_coords[, 2],
    value = factor(value, levels = c(0, 1, 2, 3), labels = c("None", "Few", "Many", "Abundant"))
  )

# Create plot
g <- ggplot() +
  # Raster points with category-based fill
  geom_point(
    data = GUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +

  # Rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # Discrete viridis scale
  scale_color_viridis_d(
    name = "Relative\nAbundance",
    option = "C"
  ) +

  # Customize legend key size
  guides(color = guide_legend(override.aes = list(size = 5))) +

  # Coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # Themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c(0.98, 0.98),
    legend.justification = c("right", "top")
    # legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # North arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

# Show plot
g

# Save to file
ggsave(
  filename = "./Plots/Maps/Abundance_Maps/GUR_RelativeAbundance_Map_Rotated.png",
  plot = g,
  width = 6,
  height = 8,
  dpi = 300
)

# Store for reuse
GUR_RA_Plot_Continuous <- g

```

##Red continuous - rotated
```{r}
# Read in and convert raster to points
RUR_points_full <- terra::as.points(RUR_Ordinal_Raster_Mask)
coords_points <- terra::crds(RUR_points_full)
extracted <- terra::extract(RUR_Ordinal_Raster_Mask, coords_points)
values_points <- extracted[[1]]  # First and only column

# Build dataframe with coordinates and extracted values
RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[, 1],
    y = sf::st_coordinates(.)[, 2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

# Rotate coordinates
rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[, 1],
    y_rot = rotated_coords[, 2],
    value = droplevels(factor(value, levels = c(1, 2, 3, 4), labels = c("None", "Few", "Many", "Abundant")))
  )


# Create plot
g <- ggplot() +
  # Raster points with category-based fill
  geom_point(
    data = RUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +

  # Rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # Discrete viridis scale
  scale_color_viridis_d(
    name = "Relative\nAbundance",
    option = "C"
  ) +

  # Customize legend key size
  guides(color = guide_legend(override.aes = list(size = 5))) +

  # Coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # Themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c(0.98, 0.98),
    legend.justification = c("right", "top")
    # legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # North arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

# Show plot
g

# Save to file
ggsave(
  filename = "./Plots/Maps/Abundance_Maps/RUR_RelativeAbundance_Map_Rotated.png",
  plot = g,
  width = 6,
  height = 8,
  dpi = 300
)

# Store for reuse
RUR_RA_Plot_Continuous <- g

```

###Combine continuous RA
```{r}
abundance_continuous_plots<- GUR_RA_Plot_Continuous + RUR_RA_Plot_Continuous + 
  plot_layout(ncol = 2)  

abundance_continuous_plots

ggsave(
  filename = "./Plots/Maps/Abundance_Maps/Combined_RA_Continuous.png",   # or another path/filename
  plot = abundance_continuous_plots,
  width = 20,        
  height = 16,       
  dpi = 300         
)
```


#Cluster maps 
```{r}
#Import one of my raster prediction layers
GUR_BP <- GUR_Ordinal_Raster_Mask

#Aggregate the raster based on sum
GUR_Agg <- terra::aggregate(GUR_BP, fact=1, fun="mean")

model_name <- c("GUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo")

#Convert the raster to points
GUR_Points <- terra::as.points(GUR_Agg)

#Convert to a dataframe with coordinates and drop absences
GUR_df <- st_as_sf(GUR_Points) %>%
  dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry() 

GUR_df <- filter(GUR_df, GUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo %in% c(1, 2, 3)) 
GUR_df <- rename(GUR_df, Likelihood = GUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo) 
GUR_df <- mutate(GUR_df, Likelihood = factor(Likelihood,
                             levels = c(1, 2, 3),
                             labels = c("Few", "Many", "Abundant")))


GUR_RA_Map <- ggplot() +
  geom_sf(data = land_sf, fill = "grey80", color = NA) +
  geom_point(data = GUR_df, aes(x = x, y = y, color = Likelihood), size = 0.3, alpha = 0.2) +
  scale_color_manual(
    name = "Relative Abundance",
    values = c("Few" = "#cfeec9", "Many" = "#55C667FF", "Abundant" = "#1b5817" )
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)) + 
    guides(color = guide_legend(override.aes = list(size = 4, alpha = 1)))  

GUR_RA_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_Relative_Map.png",   # or another path/filename
  plot = GUR_RA_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)



```
#Green urchin - rotated 
```{r}
names(GUR_BP) <- "value"

#Rotate green urchin points using the same rotation parameters
GUR_df_rotated_coords <- rotate_points(
  points = GUR_df[, c("x", "y")], 
  angle_deg = rotation_angle, 
  center = center
)

#Add rotated coordinates back to the dataframe
GUR_df <- GUR_df %>%
  dplyr::mutate(
    x_rot = GUR_df_rotated_coords[,1],
    y_rot = GUR_df_rotated_coords[,2]
  )

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_BP)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_BP, coords_points)
values_points <- extracted$value 


GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

#Plot
q <- ggplot() +
  geom_point(data = GUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +
  geom_point(data = GUR_df, aes(x = x_rot, y = y_rot, color = Likelihood), size = 0.3, alpha = 0.2) +
  scale_color_manual(
    name = "Relative Abundance",
    values = c("Few" = "#cfeec9", "Many" = "#55C667FF","Abundant" = "#1b5817" )
  ) +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)) + 
    guides(color = guide_legend(override.aes = list(size = 4, alpha = 1))) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_Relative_Map_Rotated.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)

GUR_RA_Map <- q
```

```{r}
#Plot
q <- ggplot() +
  geom_point(data = GUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "white", color = 'black') +
  geom_point(data = GUR_df, aes(x = x_rot, y = y_rot, color = Likelihood), size = 0.3, alpha = 0.2) +
  scale_color_manual(
    name = "Relative Abundance",
    values = c("Few" = "#cfeec9", "Many" = "#55C667FF","Abundant" = "#1b5817" )
  ) +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)) + 
    guides(color = guide_legend(override.aes = list(size = 4, alpha = 1))) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_Relative_Map_Rotated_outline.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```

#Red urchin
```{r}
#Import one of my raster prediction layers
RUR_BP <- RUR_Ordinal_Raster_Mask

#Aggregate the raster based on sum
RUR_Agg <- terra::aggregate(RUR_BP, fact=1, fun="mean")

model_name <- c("RUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo")

#Convert the raster to points
RUR_Points <- terra::as.points(RUR_Agg)

#Convert to a dataframe with coordinates and drop absences
RUR_df <- st_as_sf(RUR_Points) %>%
  dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry() %>%
  filter(RUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo %in% c(2, 3, 4)) %>%
  rename(Likelihood = RUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo) %>%
  mutate(Likelihood = factor(Likelihood,
                             levels = c(2, 3, 4),
                             labels = c("Few", "Many", "Abundant")))


RUR_RA_Map <- ggplot() +
  geom_sf(data = land_sf, fill = "grey80", color = NA) +
  geom_point(data = RUR_df, aes(x = x, y = y, color = Likelihood), size = 0.3, alpha = 0.2) +
  scale_color_manual(
    name = "Relative Density",
    values = c("Few" = "#f8b4b2", "Many" = "#e6360F", "Abundant"='#760b08')
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)) + 
    guides(color = guide_legend(override.aes = list(size = 4, alpha = 1)))  

  
RUR_RA_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_Relative_Map.png",   # or another path/filename
  plot = RUR_RA_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)

#Import one of my raster prediction layers
RUR_BP <- RUR_Ordinal_Raster_Mask

#Aggregate the raster based on sum
RUR_Agg <- terra::aggregate(RUR_BP, fact=1, fun="mean")

model_name <- c("RUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo")

#Convert the raster to points
RUR_Points <- terra::as.points(RUR_Agg)

#Convert to a dataframe with coordinates and drop absences
RUR_df <- st_as_sf(RUR_Points) %>%
  dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry() %>%
  filter(RUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo %in% c(2, 3, 4)) %>%
  rename(Likelihood = RUR_EMwmeanByF1_mergedData_mergedRun_mergedAlgo) %>%
  mutate(Likelihood = factor(Likelihood,
                             levels = c(2, 3, 4),
                             labels = c("Few", "Many", "Abundant")))


RUR_RA_Map <- ggplot() +
  geom_sf(data = land_sf, fill = "grey80", color = NA) +
  geom_point(data = RUR_df, aes(x = x, y = y, color = Likelihood), size = 0.3, alpha = 0.2) +
  scale_color_manual(
    name = "Relative Abundance",
    values = c("Few" = "#f8b4b2", "Many" = "#760b08")
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)) + 
    guides(color = guide_legend(override.aes = list(size = 4, alpha = 1)))  
  

RUR_RA_Map

```

#Red urchin - rotated
```{r}
names(RUR_BP) <- "value"

#Rotate green urchin points using the same rotation parameters
RUR_df_rotated_coords <- rotate_points(
  points = RUR_df[, c("x", "y")], 
  angle_deg = rotation_angle, 
  center = center
)

#Add rotated coordinates back to the dataframe
RUR_df <- RUR_df %>%
  dplyr::mutate(
    x_rot = RUR_df_rotated_coords[,1],
    y_rot = RUR_df_rotated_coords[,2]
  )

#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_BP)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_BP, coords_points)
values_points <- extracted$value 


RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

#Plot
q <- ggplot() +
  geom_point(data = RUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +
  geom_point(data = RUR_df, aes(x = x_rot, y = y_rot, color = Likelihood), size = 0.3, alpha = 0.2) +
  scale_color_manual(
    name = "Relative Abundance",
    values = c("Few" = "#f8b4b2", "Many" = "#e6360F","Abundant"='#760b08')
  ) +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)) + 
    guides(color = guide_legend(override.aes = list(size = 4, alpha = 1))) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_Relative_Map_Rotated.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)

RUR_RA_Map <- q
```

```{r}
#Plot
q <- ggplot() +
  geom_point(data = RUR_df_full, aes(x = x_rot, y = y_rot), color = "#e5e7e9", shape = 15, size = 0.3, alpha = 0.1) +  
  geom_sf(data = land_rotated, fill = "white", color = 'black') +
  geom_point(data = RUR_df, aes(x = x_rot, y = y_rot, color = Likelihood), size = 0.3, alpha = 0.2) +
  scale_color_manual(
    name = "Relative Abundance",
    values = c("Few" = "#f8b4b2", "Many" = "#e6360F","Abundant"='#760b08')
  ) +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)) + 
    guides(color = guide_legend(override.aes = list(size = 4, alpha = 1))) +
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

q

#Save 
ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_Relative_Map_Rotated_outline.png",   # or another path/filename
  plot = q,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```
 
#Combined relative maps
```{r}
ra_combined_plot <- GUR_RA_Map + RUR_RA_Map + 
  plot_layout(ncol = 2)  

ggsave(
  filename = "./Plots/Maps/Cluster_maps/Combined_RA_Map.png",   # or another path/filename
  plot = ra_combined_plot,
  width = 20,        
  height = 16,       
  dpi = 300            
)

```

#Relative Abundance Uncertainty Maps ####
```{r}
#read in uncertainty 
GUR_Uncertainty <-terra::rast("./processed_data/Uncertainty/GUR_Method_Uncertainty_RA.tif")

#Adding buffer, rotated 
GUR_points_full <- terra::as.points(GUR_Uncertainty)
coords_points <- terra::crds(GUR_points_full)

extracted <- terra::extract(GUR_Uncertainty, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
GUR_df_full <- st_as_sf(GUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = GUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

GUR_df_full <- GUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = GUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Uncertainty", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
    #legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave("./Plots/Maps/Uncertainty_Plots/GUR_Model_Uncertainty_Map_RA_Rotate.jpeg", plot = g, width = 10, height = 8, dpi = 300)

GUR_Uncertainty_Plot  <- g
```

#red urchin 
```{r}
#read in uncertainty 
RUR_Uncertainty <-terra::rast("./processed_data/Uncertainty/RUR_Method_Uncertainty_RA.tif")

#Adding buffer, rotated 
RUR_points_full <- terra::as.points(RUR_Uncertainty)
coords_points <- terra::crds(RUR_points_full)

extracted <- terra::extract(RUR_Uncertainty, coords_points)
values_points <- extracted[[1]]  # this is usually the first and only column

#build dataframe with coordinated values
RUR_df_full <- st_as_sf(RUR_points_full) %>%
  mutate(
    x = sf::st_coordinates(.)[,1],
    y = sf::st_coordinates(.)[,2],
    value = values_points
  ) %>%
  st_drop_geometry() %>%
  filter(!is.na(value))

#rotate coordinates
rotated_coords <- rotate_points(
  points = RUR_df_full[, c("x", "y")],
  angle_deg = rotation_angle,
  center = center
)

RUR_df_full <- RUR_df_full %>%
  mutate(
    x_rot = rotated_coords[,1],
    y_rot = rotated_coords[,2]
  )

g <- ggplot() +
  # raster points with value mapped to color
  geom_point(
    data = RUR_df_full,
    aes(x = x_rot, y = y_rot, color = value),
    shape = 15, size = 0.3, alpha = 0.7
  ) +
  
  # rotated land layer
  geom_sf(data = land_rotated, fill = "grey70", color = NA) +

  # viridis continuous color scale for uncertainty
  scale_color_viridis_c(name = "Uncertainty", option = "C") +

  # coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +

  # themes and styling
  theme_minimal() + theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
      legend.position = c(0.98, 0.98),   legend.justification = c("right", "top"),
   # legend.background = element_rect(fill = "white", color = 'black')
  ) +

  # north arrow
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  )

g

ggsave("./Plots/Maps/Uncertainty_Plots/RUR_Model_Uncertainty_Map_RA_Rotate.jpeg", plot = g, width = 10, height = 8, dpi = 300)

RUR_Uncertainty_Plot <- g
```

#combine RA uncertain
```{r}
RA_uncertainty_combined_plot <- GUR_Uncertainty_Plot + RUR_Uncertainty_Plot + 
  plot_layout(ncol = 2)  

RA_uncertainty_combined_plot

ggsave(
  filename = "./Plots/Maps/Uncertainty_Plots/Combined_RA_Uncertainty_Map.png",   # or another path/filename
  plot = RA_uncertainty_combined_plot,
  width = 20,        
  height = 16,       
  dpi = 300         
)
```

#Combine all maps ####
```{r}
total_combined <- PUR_Point_Map + plot_spacer() +  plot_spacer() + GUR_Point_Map + GUR_Count_Map + GUR_RA_Map + RUR_Point_Map + RUR_Count_Map + RUR_RA_Map + 
  plot_layout(ncol = 3, nrow = 3)  

ggsave(
  filename = "./Plots/Maps/Cluster_maps/Combined_Maps.png",   # or another path/filename
  plot = total_combined,
  width = 20,        
  height = 16,       
  dpi = 300            
)
```

#Maps by species 
```{r}
green_combined<- GUR_Point_Map + GUR_Count_Map + GUR_RA_Map +  plot_layout(ncol = 3)  

ggsave(
  filename = "./Plots/Maps/Cluster_maps/Green_Maps.png",   # or another path/filename
  plot = green_combined,
  width = 20,        
  height = 16,       
  dpi = 300            
)

red_combined<- RUR_Point_Map + RUR_Count_Map + RUR_RA_Map +  plot_layout(ncol = 3)  

ggsave(
  filename = "./Plots/Maps/Cluster_maps/Red_Maps.png",   # or another path/filename
  plot = red_combined,
  width = 20,        
  height = 16,       
  dpi = 300            
)



```

#Congruency Maps ####

##Green urchin
```{r}
# Load and threshold rasters (stay as rasters for now!)
GUR_Binary <- rast("./processed_data/Binary_Predictions/GUR_Binary_Predictions.tif")
GUR_Count_Raster <- rast("./models_biomod/count/GUR/proj_GUR_current_count_ensembles/proj_GUR_current_count_ensembles_GUR_ensemble.tif")
GUR_Count_Raster[GUR_Count_Raster > 0] <- 1  # convert all cells > 0 to 1

GUR_Ordinal_Raster <- rast("./models_biomod/ordinal/GUR/proj_GUR_current_ord_ensembles/proj_GUR_current_ord_ensembles_GUR_ensemble.tif")
GUR_Ordinal_Raster[GUR_Ordinal_Raster == 0] <- 0  # category 1 = None
GUR_Ordinal_Raster[GUR_Ordinal_Raster > 0] <- 1  # anything above = presence

# Your three binary rasters:
r1 <- GUR_Binary
r2 <- GUR_Count_Raster
r3 <- GUR_Ordinal_Raster

# Create logical presence/absence rasters
p1 <- r1 == 1; p1[is.na(p1)] <- FALSE
p2 <- r2 == 1; p2[is.na(p2)] <- FALSE
p3 <- r3 == 1; p3[is.na(p3)] <- FALSE

a1 <- r1 == 0; a1[is.na(a1)] <- FALSE
a2 <- r2 == 0; a2[is.na(a2)] <- FALSE
a3 <- r3 == 0; a3[is.na(a3)] <- FALSE

# Sum presence and absence
presence_sum <- p1 + p2 + p3
absence_sum <- a1 + a2 + a3

# Define agreement categories
full_presence <- presence_sum == 3
full_absence <- absence_sum == 3
partial_presence <- presence_sum == 2 & absence_sum == 1
partial_absence <- presence_sum == 1 & absence_sum == 2
no_alignment <- is.na(presence_sum) | is.na(absence_sum)

# Create an empty raster to hold alignment categories
alignment_raster <- rast(r1)
values(alignment_raster) <- NA

# Assign category values
values(alignment_raster)[which(values(full_presence) == 1)] <- 1
values(alignment_raster)[which(values(full_absence) == 1)] <- 2
values(alignment_raster)[which(values(partial_presence) == 1)] <- 3
values(alignment_raster)[which(values(partial_absence) == 1)] <- 4
values(alignment_raster)[which(values(no_alignment) == 1)] <- 5

# Add category labels
levels(alignment_raster) <- data.frame(
  ID = 1:5,
  category = c("full_presence", "full_absence", "partial_presence", "partial_absence", "no_alignment")
)

# Create buffer mask: include only where ALL models predict something (presence or absence)
buffer_mask <- (p1 | a1) & (p2 | a2) & (p3 | a3)

# Mask to buffer and optionally to bathymetry
alignment_raster <- mask(alignment_raster, buffer_mask)
alignment_raster <- mask(alignment_raster, Bathymetry)

# Aggregate based on modal value (no-op if fact = 1)
GUR_Agg <- terra::aggregate(alignment_raster, fact = 1, fun = "modal")

# Convert to points
GUR_Points <- terra::as.points(GUR_Agg)

# Convert to dataframe and extract coordinates
GUR_df <- st_as_sf(GUR_Points) %>%
  dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry()

# Rotate point coordinates to match rotated land
GUR_df[, c("x", "y")] <- rotate_points(GUR_df[, c("x", "y")], rotation_angle, center)

# Plot
GUR_Point_Map <- ggplot() +
  geom_sf(data = land_rotated, fill = "grey80", color = NA) +
  geom_point(data = GUR_df, aes(x = x, y = y, color = category), size = 0.0001, alpha = 0.1) +
  scale_color_manual(
    name = "",
    values = c(
      "full_presence" = "#ff7f00",
      "partial_presence" = "#fdbf6f",
      "partial_absence" = "#a6cee3",
      "full_absence" = "#1f78b4"
    )
  ) +

  coord_sf(xlim = xlim, ylim = ylim) +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = "none",
    legend.background = element_rect(fill = "white", color = NA)
  ) +
  guides(color = guide_legend(override.aes = list(size = 4, alpha = 1)))

GUR_Point_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_AlignPA_Map.png",   # or another path/filename
  plot = GUR_Point_Map,
  width = 10,        
  height = 8,       
  dpi = 300         
)
```

##Red urchin
```{r}
# Load and threshold rasters (stay as rasters for now!)
RUR_Binary <-terra::rast("./processed_data/Binary_Predictions/RUR_Binary_Predictions.tif")
RUR_Count_Raster <- rast("./models_biomod/count/RUR/proj_RUR_current_count_ensembles/proj_RUR_current_count_ensembles_RUR_ensemble.tif")
RUR_Count_Raster[RUR_Count_Raster > 0] <- 1  # convert all cells > 0 to 1

RUR_Ordinal_Raster <- rast("./models_biomod/ordinal/RUR/proj_RUR_current_ord_ensembles/proj_RUR_current_ord_ensembles_RUR_ensemble.tif")
RUR_Ordinal_Raster[RUR_Ordinal_Raster == 1] <- 0  # category 1 = None
RUR_Ordinal_Raster[RUR_Ordinal_Raster >= 1] <- 1  # anything above = presence

# Your three binary rasters:
r1 <- RUR_Binary
r2 <- RUR_Count_Raster
r3 <- RUR_Ordinal_Raster

# Create logical presence/absence rasters
p1 <- r1 == 1; p1[is.na(p1)] <- FALSE
p2 <- r2 == 1; p2[is.na(p2)] <- FALSE
p3 <- r3 == 1; p3[is.na(p3)] <- FALSE

a1 <- r1 == 0; a1[is.na(a1)] <- FALSE
a2 <- r2 == 0; a2[is.na(a2)] <- FALSE
a3 <- r3 == 0; a3[is.na(a3)] <- FALSE

# Sum presence and absence
presence_sum <- p1 + p2 + p3
absence_sum <- a1 + a2 + a3

# Define agreement categories
full_presence <- presence_sum == 3
full_absence <- absence_sum == 3
partial_presence <- presence_sum == 2 & absence_sum == 1
partial_absence <- presence_sum == 1 & absence_sum == 2
no_alignment <- is.na(presence_sum) | is.na(absence_sum)

# Create an empty raster to hold alignment categories
alignment_raster <- rast(r1)
values(alignment_raster) <- NA

# Assign category values
values(alignment_raster)[which(values(full_presence) == 1)] <- 1
values(alignment_raster)[which(values(full_absence) == 1)] <- 2
values(alignment_raster)[which(values(partial_presence) == 1)] <- 3
values(alignment_raster)[which(values(partial_absence) == 1)] <- 4
values(alignment_raster)[which(values(no_alignment) == 1)] <- 5

# Add category labels
levels(alignment_raster) <- data.frame(
  ID = 1:5,
  category = c("full_presence", "full_absence", "partial_presence", "partial_absence", "no_alignment")
)

# Create buffer mask: include only where ALL models predict something (presence or absence)
buffer_mask <- (p1 | a1) & (p2 | a2) & (p3 | a3)

# Mask to buffer and optionally to bathymetry
alignment_raster <- mask(alignment_raster, buffer_mask)
alignment_raster <- mask(alignment_raster, Bathymetry)

# Aggregate based on modal value (no-op if fact = 1)
RUR_Agg <- terra::aggregate(alignment_raster, fact = 1, fun = "modal")

# Convert to points
RUR_Points <- terra::as.points(RUR_Agg)

# Convert to dataframe and extract coordinates
RUR_df <- st_as_sf(RUR_Points) %>%
  dplyr::mutate(x = sf::st_coordinates(.)[,1],
                y = sf::st_coordinates(.)[,2]) %>%
  st_drop_geometry()

# Rotate point coordinates to match rotated land
RUR_df[, c("x", "y")] <- rotate_points(RUR_df[, c("x", "y")], rotation_angle, center)

# Plot
RUR_Point_Map <- ggplot() +
  geom_sf(data = land_rotated, fill = "grey80", color = NA) +
  geom_point(data = RUR_df, aes(x = x, y = y, color = category), size = 0.0001, alpha = 0.1) +
  scale_color_manual(
    name = "",  # legend title
    values = c(
      "full_presence" = "#ff7f00",
      "partial_presence" = "#fdbf6f",
      "partial_absence" = "#a6cee3",
      "full_absence" = "#1f78b4"
    ),
    breaks = c("full_presence", "partial_presence", "partial_absence", "full_absence"),
    labels = c("Full Presence", "Partial Presence", "Partial Absence", "Full Absence")
  ) +

  coord_sf(xlim = xlim, ylim = ylim) +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = "right",
    legend.background = element_rect(fill = "white", color = NA)
  ) +
  guides(color = guide_legend(override.aes = list(size = 4, alpha = 1)))

RUR_Point_Map

ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_AlignPA_Map.png",   # or another path/filename
  plot = RUR_Point_Map,
  width = 10,        
  height = 8,       
  dpi = 300         
)
```

#Combine
```{r}
presenceabsence_combined<- GUR_Point_Map + RUR_Point_Map +  plot_layout(ncol = 2)  

ggsave(
  filename = "./Plots/Maps/Cluster_maps/Align_Presence_Absence_Maps.png",   # or another path/filename
  plot = presenceabsence_combined,
  width = 20,        
  height = 16,       
  dpi = 300            
)
```

#Presence only alignment
##Green urchin 
```{r}

#Converting binary to only presence
GUR_Binary <-terra::rast("./processed_data/Binary_Predictions/GUR_Binary_Predictions.tif")
GUR_Binary[GUR_Binary == 0] <- NA

#Converting count to only presence (>0)
GUR_Count_Raster <- rast("./models_biomod/count/GUR/proj_GUR_current_count_ensembles/proj_GUR_current_count_ensembles_GUR_ensemble.tif")
GUR_Count_Raster[GUR_Count_Raster == 0] <- NA
GUR_Count_Raster[GUR_Count_Raster >= 0] <- 1

#Converting relative density to only presence (>None)
GUR_Ordinal_Raster <- rast("./models_biomod/ordinal/GUR/proj_GUR_current_ord_ensembles/proj_GUR_current_ord_ensembles_GUR_ensemble.tif")
GUR_Ordinal_Raster[GUR_Ordinal_Raster == 0] <- NA
GUR_Ordinal_Raster[GUR_Ordinal_Raster >= 1] <- 1

#Stack 
GUR_stack <- c(GUR_Binary, GUR_Count_Raster, GUR_Ordinal_Raster)

#Sum them cell-by-cell
GUR_sum <- sum(GUR_stack, na.rm = FALSE)  # Only sum if all are not NA

#Keep only cells where all 3 rasters agree on presence (i.e., value == 3)
GUR_agreement <- GUR_sum == 3

#Convert 0s to NA so only full-agreement areas show
GUR_agreement[GUR_agreement == 0] <- NA

# Convert raster to dataframe for plotting
GUR_agreement_df <- as.data.frame(GUR_agreement, xy = TRUE, na.rm = TRUE)
colnames(GUR_agreement_df) <- c("x", "y", "presence")

GUR_Agreement_Map <- ggplot() +
  geom_sf(data = land_sf, fill = "grey80", color = NA) +
  geom_point(
    data = GUR_agreement_df,
    aes(x = x, y = y), 
    size = 0.02, 
    alpha = 0.3,
    color = "#55C667FF"
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +
  theme_cowplot() +
  xlab(NULL) +
  ylab(NULL) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5)
  )

# Show the plot
GUR_Agreement_Map


ggsave(
  filename = "./Plots/Maps/Cluster_maps/GUR_AlignPresence_Map.png",   # or another path/filename
  plot = GUR_Agreement_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```

##Red urchin 
```{r}
#Converting binary to only presence
RUR_Binary <-terra::rast("./processed_data/Binary_Predictions/RUR_Binary_Predictions.tif")
RUR_Binary[RUR_Binary == 0] <- NA

#Converting count to only presence (>0)
RUR_Count_Raster <- rast("./models_biomod/count/RUR/proj_RUR_current_count_ensembles/proj_RUR_current_count_ensembles_RUR_ensemble.tif")
RUR_Count_Raster[RUR_Count_Raster == 0] <- NA
RUR_Count_Raster[RUR_Count_Raster >= 0] <- 1

#Converting relative density to only presence (>None)
RUR_Ordinal_Raster <- rast("./models_biomod/ordinal/RUR/proj_RUR_current_ord_ensembles/proj_RUR_current_ord_ensembles_RUR_ensemble.tif")
RUR_Ordinal_Raster[RUR_Ordinal_Raster == 1] <- NA
RUR_Ordinal_Raster[RUR_Ordinal_Raster > 1] <- 1

#Stack 
RUR_stack <- c(RUR_Binary, RUR_Count_Raster, RUR_Ordinal_Raster)

#Sum them cell-by-cell
RUR_sum <- sum(RUR_stack, na.rm = FALSE)  # Only sum if all are not NA

#Keep only cells where all 3 rasters agree on presence (i.e., value == 3)
RUR_agreement <- RUR_sum == 3

#Convert 0s to NA so only full-agreement areas show
RUR_agreement[RUR_agreement == 0] <- NA

# Convert raster to dataframe for plotting
RUR_agreement_df <- as.data.frame(RUR_agreement, xy = TRUE, na.rm = TRUE)
colnames(RUR_agreement_df) <- c("x", "y", "presence")

RUR_Agreement_Map <- ggplot() +
  geom_sf(data = land_sf, fill = "grey80", color = NA) +
  geom_point(
    data = RUR_agreement_df,
    aes(x = x, y = y), 
    size = 0.02, 
    alpha = 0.3,
    color = "#e6360F"
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +
  theme_cowplot() +
  xlab(NULL) +
  ylab(NULL) +
  theme(
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5)
  )

# Show the plot
RUR_Agreement_Map


ggsave(
  filename = "./Plots/Maps/Cluster_maps/RUR_AlignPresence_Map.png",   # or another path/filename
  plot = RUR_Agreement_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```
#Mapping survey methods across datatypes
```{r}
#load libraries
library(ggplot2)
library(tidyr)
library(dplyr)
library(cowplot)
library(terra)
library(sf)
library(ggspatial)
library(ggplot2)
library(cowplot)
library(dplyr)
library(forcats)
library(patchwork)
library(grid)

#Read in different datasets #############
presence_data <- read.csv("./processed_data/Survey_Datasets/presence_only.csv")
RA_data <- read.csv("./processed_data/Survey_Datasets/relative_density.csv")
count_data <- read.csv("./processed_data/Survey_Datasets/actual_density.csv")

#Rename surveys - presence
#Rename DFO surveys 
presence_data <- presence_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("RSU_RESM", "RSU_RES", "RSU_RMS", "RSU_RBS", "RSU_RBB") ~ "DFO/RSU",
    TRUE ~ survey_type
  ))

presence_data <- presence_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("GSU_GSS") ~ "DFO/GSU",
    TRUE ~ survey_type
  ))

presence_data <- presence_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("DFO_MSP", "Multispecies_Multispecies", "DFO_MSP") ~ "DFO/MSP",
    TRUE ~ survey_type
  ))

presence_data <- presence_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("DFO_BHM") ~ "DFO/BHM",
    TRUE ~ survey_type
  ))

presence_data <- presence_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("RL_Survey") ~ "Reef_Life_Survey",
    TRUE ~ survey_type
  ))

#Rename surveys - count
count_data <- count_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("RSU_RESM", "RSU_RES", "RSU_RMS", "RSU_RBS", "RSU_RBB") ~ "DFO/RSU",
    TRUE ~ survey_type
  ))


count_data <- count_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("GSU_GSS") ~ "DFO/GSU",
    TRUE ~ survey_type
  ))

count_data <- count_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("DFO_MSP", "Multispecies_Multispecies", "DFO_MSP") ~ "DFO/MSP",
    TRUE ~ survey_type
  ))

count_data <- count_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("DFO_BHM") ~ "DFO/BHM",
    TRUE ~ survey_type
  ))

count_data <- count_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("RL_Survey") ~ "Reef_Life_Survey",
    TRUE ~ survey_type
  ))

#Rename surveys - RA
RA_data <- RA_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("RSU_RESM", "RSU_RES", "RSU_RMS", "RSU_RBS", "RSU_RBB") ~ "DFO/RSU",
    TRUE ~ survey_type
  ))

RA_data <- RA_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("GSU_GSS") ~ "DFO/GSU",
    TRUE ~ survey_type
  ))

RA_data <- RA_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("DFO_MSP", "Multispecies_Multispecies", "DFO_MSP") ~ "DFO/MSP",
    TRUE ~ survey_type
  ))

RA_data <- RA_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("DFO_BHM") ~ "DFO/BHM",
    TRUE ~ survey_type
  ))

RA_data <- RA_data %>%
  mutate(survey_type = case_when(
    survey_type %in% c("RL_Survey") ~ "Reef_Life_Survey",
    TRUE ~ survey_type
  ))



#Load land file ############################
#geographic layers
Land <- terra::vect("./raw_data/BC_Buffer/Land.shp") 
# Convert land to sf
land_sf <- st_as_sf(Land)

Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Rotate shapefile function #########################
rotate_sf <- function(sf_obj, angle_deg, center = NULL) {
  angle_rad <- angle_deg * pi / 180
  
  #get centroid if center not provided
  if (is.null(center)) {
    center <- st_centroid(st_union(sf_obj))
    center_coords <- st_coordinates(center)
  } else {
    center_coords <- center
  }
  
  #apply rotation on each geometry separately
  rotated_geoms <- lapply(st_geometry(sf_obj), function(geom) {
    # Extract coordinates depending on geometry type
    coords_list <- list()
    
    #handle POLYGON or MULTIPOLYGON
    if (inherits(geom, "POLYGON")) {
      coords_list <- list(geom)
    } else if (inherits(geom, "MULTIPOLYGON")) {
      coords_list <- geom
    } else {
      stop("Unsupported geometry type")
    }
    
    #rotate each polygon (rings)
    rotated_polygons <- lapply(coords_list, function(polygon) {
      rotated_rings <- lapply(polygon, function(ring) {
        # matrix of coordinates
        coords <- ring
        
        # translate to origin (center)
        coords_centered <- coords - matrix(center_coords, nrow = nrow(coords), ncol = 2, byrow = TRUE)
        
        # rotation matrix
        rotation_matrix <- matrix(c(cos(angle_rad), -sin(angle_rad),
                                    sin(angle_rad), cos(angle_rad)),
                                  nrow = 2, byrow = TRUE)
        
        # rotate coords
        rotated <- t(rotation_matrix %*% t(coords_centered))
        
        # translate back
        rotated <- rotated + matrix(center_coords, nrow = nrow(rotated), ncol = 2, byrow = TRUE)
        
        # ensure polygon ring is closed: first point == last point
        if (!all(rotated[1,] == rotated[nrow(rotated), ])) {
          rotated <- rbind(rotated, rotated[1, ])
        }
        
        return(rotated)
      })
      return(rotated_rings)
    })
    
    #build new geometry with rotated polygons
    if (inherits(geom, "POLYGON")) {
      st_polygon(rotated_polygons[[1]])
    } else if (inherits(geom, "MULTIPOLYGON")) {
      st_multipolygon(rotated_polygons)
    }
  })
  
  #return sf with new rotated geometries
  new_sf <- sf_obj
  st_geometry(new_sf) <- st_sfc(rotated_geoms, crs = st_crs(sf_obj))
  return(new_sf)
}

rotate_points <- function(points, angle_deg, center) {
  angle_rad <- angle_deg * pi / 180
  coords_centered <- as.matrix(points) - matrix(center, nrow = nrow(points), ncol = 2, byrow = TRUE)
  rotation_matrix <- matrix(c(cos(angle_rad), -sin(angle_rad),
                              sin(angle_rad), cos(angle_rad)),
                            nrow = 2, byrow = TRUE)
  rotated <- t(rotation_matrix %*% t(coords_centered))
  rotated + matrix(center, nrow = nrow(points), ncol = 2, byrow = TRUE)
}

#Rotate land shapefile ###############
rotation_angle <- -30

#Rotate land
land_rotated <- rotate_sf(land_sf, rotation_angle)

#Define bounding box corners 
coords <- data.frame(
  x = c(1100000, 950000, 750000, 1300000),
  y = c(950000, 1320000, 230000, 450000)
)
#Rotate bounding box corners around the same centroid
#Use the centroid of land_sf for center of rotation
center <- st_coordinates(st_centroid(st_union(land_sf)))

rotated_bbox_coords <- rotate_points(coords, rotation_angle, center)

xlim <- range(rotated_bbox_coords[,1])
ylim <- range(rotated_bbox_coords[,2])


x_center <- mean(rotated_bbox_coords[,1])
y_range <- range(rotated_bbox_coords[,2])  # keep this unchanged

#Define a new, narrower x-range (e.g., 300 km wide instead of full range)
#Adjust the width to your liking
x_half_width <- 300000  # in meters

xlim <- c(x_center - x_half_width, x_center + x_half_width)
ylim <- y_range  # unchanged

#Shift center slightly left (e.g., by 75,000 m)
x_center_adjusted <- x_center - 75000  # Try 50000, 75000, or 100000 as needed

#keep the same x half-width
xlim <- c(x_center_adjusted - x_half_width, x_center_adjusted + x_half_width)

#plot the rotated land shapefile with rotated bounding box limits
ggplot() +
  geom_sf(data = land_rotated, fill = "grey70", color = "black") +
  coord_sf(xlim = xlim, ylim = ylim) +
  theme_minimal() +
  labs(title = "Land shapefile rotated")


#Rotated NORTH arrows
#determining angle of north arrow
arrow_x <- xlim[1] + 0.05 * diff(xlim)  # 5% in from the left
arrow_y <- ylim[1] + 0.05 * diff(ylim)  # 5% up from the bottom

arrow_length <- 50000  # arrow length in map units (e.g., meters)

#Rotate arrow
angle_deg <- 180 - (90 + (-rotation_angle))  
angle_rad <- angle_deg * pi / 180

#Endpoint of arrow
end_x <- arrow_x + arrow_length * cos(angle_rad)
end_y <- arrow_y + arrow_length * sin(angle_rad)

#Make datasets into shapefiles and rotate ##########
#PRESENCE 
# Convert presence_data to sf if not already
presence_sf <- st_as_sf(presence_data, coords = c("start_lon", "start_lat"), crs = 4326)
presence_sf <- st_transform(presence_sf, crs = st_crs(land_sf))  # match CRS

# Extract coordinates
presence_coords <- st_coordinates(presence_sf)

# Rotate coordinates around the land centroid
center <- st_coordinates(st_centroid(st_union(land_sf)))
rotated_coords <- rotate_points(presence_coords, rotation_angle, center)

# Replace geometry with rotated coordinates
presence_sf_rotated <- st_as_sf(
  presence_sf %>% st_drop_geometry() %>% 
    mutate(X_rot = rotated_coords[,1], Y_rot = rotated_coords[,2]),
  coords = c("X_rot", "Y_rot"),
  crs = st_crs(land_sf)
)

#COUNT 
# Convert count_data to sf if not already
count_sf <- st_as_sf(count_data, coords = c("start_lon", "start_lat"), crs = 4326)
count_sf <- st_transform(count_sf, crs = st_crs(land_sf))  # match CRS

# Extract coordinates
count_coords <- st_coordinates(count_sf)

# Rotate coordinates around the land centroid
center <- st_coordinates(st_centroid(st_union(land_sf)))
rotated_coords <- rotate_points(count_coords, rotation_angle, center)

# Replace geometry with rotated coordinates
count_sf_rotated <- st_as_sf(
  count_sf %>% st_drop_geometry() %>% 
    mutate(X_rot = rotated_coords[,1], Y_rot = rotated_coords[,2]),
  coords = c("X_rot", "Y_rot"),
  crs = st_crs(land_sf)
)

#RA
# Convert RA_data to sf if not already
RA_sf <- st_as_sf(RA_data, coords = c("start_lon", "start_lat"), crs = 4326)
RA_sf <- st_transform(RA_sf, crs = st_crs(land_sf))  # match CRS

# Extract coordinates
RA_coords <- st_coordinates(RA_sf)

# Rotate coordinates around the land centroid
center <- st_coordinates(st_centroid(st_union(land_sf)))
rotated_coords <- rotate_points(RA_coords, rotation_angle, center)

# Replace geometry with rotated coordinates
RA_sf_rotated <- st_as_sf(
  RA_sf %>% st_drop_geometry() %>% 
    mutate(X_rot = rotated_coords[,1], Y_rot = rotated_coords[,2]),
  coords = c("X_rot", "Y_rot"),
  crs = st_crs(land_sf)
)

#PLOTTING ################
my_colors <- c(
  "Baum_ROV" = "#1b9e77",
  "Hakai" = "#d95f02",
  "Reef_Life_Survey" = "#7570b3",
  "DFO/RSU" = "#e7298a",
  "DFO/GSU" = "#66a61e",
  "DFO/BHM" = "#e6ab02",
  "DFO/MSP" = "#1f78b4"
)

#PRESENCE
# Rotated survey map
Method_Point_Map<- ggplot() +
  # Rotated land
  geom_sf(data = land_rotated, fill = "grey80", color = NA) +
  
  # Rotated survey points
  geom_sf(
    data = presence_sf_rotated,
    aes(color = survey_type),
    size = 1,
    alpha = 0.2
  ) +
  
  # Color scale
  scale_color_manual(name = "Survey Type", values = my_colors) +
  
  # North arrow and scale
  annotation_scale(location = "bl", width_hint = 0.5) +
  
  # Add rotated north arrow manually
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  ) +
  
  # Coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +
  
  # Styling
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = "bottom",
    legend.background = element_rect(fill = "white", color = NA)
  )

# Plot
Method_Point_Map

ggsave(
  filename = "./Plots/Maps/Methods/Presence_Method_Point_Map.png",   # file name
  plot = Method_Point_Map,             # the ggplot object
  width = 10,                          # width in inches
  height = 8,                          # height in inches
  dpi = 300                            # resolution
)

#COUNT
# Rotated survey map
Method_Point_Map <- ggplot() +
  # Rotated land
  geom_sf(data = land_rotated, fill = "grey80", color = NA) +
  
  # Rotated survey points
  geom_sf(
    data = count_sf_rotated,
    aes(color = survey_type),
    size = 1,
    alpha = 0.2
  ) +
  
  # Color scale
  scale_color_manual(name = "Survey Type", values = my_colors) +
  
  # North arrow and scale
  annotation_scale(location = "bl", width_hint = 0.5) +
  
  # Add rotated north arrow manually
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  ) +
  
  # Coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +
  
  # Styling
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = "bottom",
    legend.background = element_rect(fill = "white", color = NA)
  )

# Plot
Method_Point_Map

ggsave(
  filename = "./Plots/Maps/Methods/Count_Method_Point_Map.png",   # file name
  plot = Method_Point_Map,             # the ggplot object
  width = 10,                          # width in inches
  height = 8,                          # height in inches
  dpi = 300                            # resolution
)

#RA
#COUNT
# Rotated survey map
Method_Point_Map <- ggplot() +
  # Rotated land
  geom_sf(data = land_rotated, fill = "grey80", color = NA) +
  
  # Rotated survey points
  geom_sf(
    data = RA_sf_rotated,
    aes(color = survey_type),
    size = 1,
    alpha = 0.2
  ) +
  
  # Color scale
  scale_color_manual(name = "Survey Type", values = my_colors) +
  
  # North arrow and scale
  annotation_scale(location = "bl", width_hint = 0.5) +
  
  # Add rotated north arrow manually
  geom_segment(
    aes(x = arrow_x, y = arrow_y, xend = end_x, yend = end_y),
    arrow = arrow(length = unit(0.3, "cm")),
    size = 1,
    color = "black"
  ) +
  annotate(
    "text",
    x = end_x + 20000 * cos(angle_rad),
    y = end_y + 20000 * sin(angle_rad),
    label = "N",
    size = 5,
    fontface = "bold"
  ) +
  
  # Coordinate limits
  coord_sf(xlim = xlim, ylim = ylim) +
  
  # Styling
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = "bottom",
    legend.background = element_rect(fill = "white", color = NA)
  )

# Plot
Method_Point_Map


ggsave(
  filename = "./Plots/Maps/Methods/RA_Method_Point_Map.png",   # file name
  plot = Method_Point_Map,             # the ggplot object
  width = 10,                          # width in inches
  height = 8,                          # height in inches
  dpi = 300                            # resolution
)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
