---
title: "Model_comparisons"
author: "Tessa Rehill"
date: "2025-05-13"
output: html_document
---

#comparing model outputs between occurrence, and count models 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#load packages

```{r}
  library(terra)
  library(sf)
  library(dplyr)
  library(ggplot2)
  library(cowplot)
  library(ggspatial)
  library(sf)
  library(tibble)
  library(knitr)
  library(kableExtra)
  library(patchwork)


```

#load data 
```{r}
#presence predictions
GUR_LO_Raster <- rast("./processed_data/LO_Predictions/GUR_LO_Predictions.tif")
RUR_LO_Raster <- rast("./processed_data/LO_Predictions/RUR_LO_Predictions.tif")
PUR_LO_Raster <- rast("./processed_data/LO_Predictions/PUR_LO_Predictions.tif")

#count predictions
GUR_Count_Raster <- rast("./models_biomod/count/GUR/proj_GUR_current_count_ensembles/proj_GUR_current_count_ensembles_GUR_ensemble.tif")
RUR_Count_Raster <- rast("./models_biomod/count/RUR/proj_RUR_current_count_ensembles/proj_RUR_current_count_ensembles_RUR_ensemble.tif")

#masks (extrapolated variables)
GURclamp <- rast("./models_biomod/count/GUR/proj_GUR_current_count/proj_GUR_current_count_ClampingMask.tif")
RURclamp <- rast("./models_biomod/count/RUR/proj_RUR_current_count/proj_RUR_current_count_ClampingMask.tif")

#Remove cells that extrapolate
#GUR_Count_Raster <- mask(GUR_Count_Raster, GURclamp >= 1, maskvalue = TRUE)
#RUR_Count_Raster <- mask(RUR_Count_Raster, RURclamp >= 1, maskvalue = TRUE)


#predictions
#PUR_Ordinal_Raster <- rast("./models_biomod/ordinal/PUR/PUR/proj_PUR_current_ord_ensembles/proj_PUR_current_ord_ensembles_PUR_ensemble.tif")
RUR_Ordinal_Raster <- rast("./models_biomod/ordinal/RUR/proj_RUR_current_ord_ensembles/proj_RUR_current_ord_ensembles_RUR_ensemble.tif")
GUR_Ordinal_Raster <- rast("./models_biomod/ordinal/GUR/proj_GUR_current_ord_ensembles/proj_GUR_current_ord_ensembles_GUR_ensemble.tif")

#masks (extrapolated variables)
#GURclamp <- rast("./models_biomod/ordinal/GUR/GUR/proj_GUR_current_ord/proj_GUR_current_ord_ClampingMask.tif")
#RURclamp <- rast("./models_biomod/ordinal/RUR/RUR/proj_RUR_current_ord/proj_RUR_current_ord_ClampingMask.tif")
#PURclamp <- rast("./models_biomod/ordinal/PUR/PUR/proj_PUR_current_ord/proj_PUR_current_ord_ClampingMask.tif")


#Remove cells that extrapolate
#GUR_Ordinal_Raster <- mask(GUR_Ordinal_Raster, GURclamp >= 1, maskvalue = TRUE)
#RUR_Ordinal_Raster <- mask(RUR_Ordinal_Raster, RURclamp >= 1, maskvalue = TRUE)
#PUR_Ordinal_Raster <- mask(PUR_Ordinal_Raster, PURclamp >= 1, maskvalue = TRUE)


#geographic layers
Land <- terra::vect("./raw_data/BC_Buffer/Land.shp") 
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")


```


#Comparing likelihood of occurrence to count predictions ####
#Method: Spearmans correlation coefficient
##Green urchin
```{r}

# --- RASTER ALIGNMENT ---
GUR_LO_Raster <- resample(GUR_LO_Raster, GUR_Count_Raster, method = "bilinear")  # Match resolution

# --- STACK AND PREP DATA ---
stacked <- c(GUR_Count_Raster, GUR_LO_Raster)
names(stacked) <- c("count", "likelihood")

vals_df <- as.data.frame(stacked, na.rm = TRUE)

# --- CHECK ASSUMPTIONS ---

#Check normality - not normal 
qqnorm(vals_df$count)
qqline(vals_df$count, col = "blue")

qqnorm(vals_df$likelihood)
qqline(vals_df$likelihood, col = "red")


#Check monotonicity 
ggplot(vals_df, aes(x = likelihood, y = count)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "blue", se = FALSE) +
  theme_minimal() +
  labs(title = "Check for monotonicity between count and likelihood",
       x = "Likelihood of Occurrence",
       y = "Urchin Count")


# 2. Check for excessive ties (count data tends to be rounded or integer-valued)
summary(as.factor(vals_df$count))  # Optional: assess frequency of repeated values
length(unique(vals_df$count))      # Fewer unique values = more ties (Spearman can tolerate some)

# 3. Independence assumption: Visual inspection (no autocorrelation check here, just note)
# If you're concerned about spatial autocorrelation, Moran's I is an option 

# --- RUN SPEARMAN'S RANK CORRELATION TEST ---
spearman_result <- cor.test(vals_df$count, vals_df$likelihood, method = "spearman")
print(spearman_result)

GUR_LOC_v_count <- spearman_result

GUR_LOC_v_count

```

#Red urchin
```{r}
# --- RASTER ALIGNMENT ---
RUR_LO_Raster <- resample(RUR_LO_Raster, RUR_Count_Raster, method = "bilinear")  # Match resolution

# --- STACK AND PREP DATA ---
stacked <- c(RUR_Count_Raster, RUR_LO_Raster)
names(stacked) <- c("count", "likelihood")

vals_df <- as.data.frame(stacked, na.rm = TRUE)

# --- CHECK ASSUMPTIONS ---
#Check normality - not normal 
qqnorm(vals_df$count)
qqline(vals_df$count, col = "blue")

qqnorm(vals_df$likelihood)
qqline(vals_df$likelihood, col = "red")


ggplot(vals_df, aes(x = likelihood, y = count)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "blue", se = FALSE) +
  theme_minimal() +
  labs(title = "Check for monotonicity between count and likelihood",
       x = "Likelihood of Occurrence",
       y = "Urchin Count")


# 2. Check for excessive ties (count data tends to be rounded or integer-valued)
summary(as.factor(vals_df$count))  # Optional: assess frequency of repeated values
length(unique(vals_df$count))      # Fewer unique values = more ties (Spearman can tolerate some)

# 3. Independence assumption: Visual inspection (no autocorrelation check here, just note)
# If you're concerned about spatial autocorrelation, Moran's I is an option 

# --- RUN SPEARMAN'S RANK CORRELATION TEST ---
spearman_result <- cor.test(vals_df$count, vals_df$likelihood, method = "spearman")
print(spearman_result)

RUR_LOC_v_count <- spearman_result


```

#Comparing count to relative abundance prediction ####
##Green urchin 
```{r}
# --- RASTER ALIGNMENT ---
GUR_Count_Raster <- resample(GUR_Count_Raster, GUR_Ordinal_Raster, method = "bilinear")  # Match resolution

# --- STACK AND PREP DATA ---
stacked <- c(GUR_Count_Raster, GUR_Ordinal_Raster)
names(stacked) <- c("count", "density")

vals_df <- as.data.frame(stacked, na.rm = TRUE)

# --- CHECK ASSUMPTIONS ---
#Check normality - not normal 
qqnorm(vals_df$count)
qqline(vals_df$count, col = "blue")

qqnorm(vals_df$density)
qqline(vals_df$density, col = "red")

#Check monotonicity
ggplot(vals_df, aes(x = density, y = count)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "blue", se = FALSE) +
  theme_minimal() +
  labs(title = "Check for monotonicity between count and relative density",
       x = "Relative Density",
       y = "Urchin Count")


# 2. Check for excessive ties (count data tends to be rounded or integer-valued)
summary(as.factor(vals_df$count))  # Optional: assess frequency of repeated values
length(unique(vals_df$count))      # Fewer unique values = more ties (Spearman can tolerate some)

summary(as.factor(vals_df$density))  
length(unique(vals_df$density))   

# 3. Independence assumption: Visual inspection (no autocorrelation check here, just note)
# If you're concerned about spatial autocorrelation, Moran's I is an option 

# --- RUN SPEARMAN'S RANK CORRELATION TEST ---
spearman_result <- cor.test(vals_df$count, vals_df$density, method = "spearman")
print(spearman_result)

GUR_RA_v_count <- spearman_result


```

##Red urchin 
```{r}
# --- RASTER ALIGNMENT ---
RUR_Count_Raster <- resample(RUR_Count_Raster, RUR_Ordinal_Raster, method = "bilinear")  # Match resolution

# --- STACK AND PREP DATA ---
stacked <- c(RUR_Count_Raster, RUR_Ordinal_Raster)
names(stacked) <- c("count", "density")

vals_df <- as.data.frame(stacked, na.rm = TRUE)

# --- CHECK ASSUMPTIONS ---

qqnorm(vals_df$count)
qqline(vals_df$count, col = "blue")

qqnorm(vals_df$density)
qqline(vals_df$density, col = "red")

ggplot(vals_df, aes(x = density, y = count)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "blue", se = FALSE) +
  theme_minimal() +
  labs(title = "Check for monotonicity between count and relative density",
       x = "Relative Density",
       y = "Urchin Count")


# 2. Check for excessive ties (count data tends to be rounded or integer-valued)
summary(as.factor(vals_df$count))  # Optional: assess frequency of repeated values
length(unique(vals_df$count))      # Fewer unique values = more ties (Spearman can tolerate some)

summary(as.factor(vals_df$density))  
length(unique(vals_df$density))   

# 3. Independence assumption: Visual inspection (no autocorrelation check here, just note)
# If you're concerned about spatial autocorrelation, Moran's I is an option 

# --- RUN SPEARMAN'S RANK CORRELATION TEST ---
spearman_result <- cor.test(vals_df$count, vals_df$density, method = "spearman")
print(spearman_result)

RUR_RA_v_count <- spearman_result

```


#Comparing likelihood of occurrence to relative abundance prediction ####
#Green urchin 
```{r}
# --- RASTER ALIGNMENT ---
GUR_Ordinal_Raster <- resample(GUR_Ordinal_Raster, GUR_LO_Raster,  method = "bilinear")  # Match resolution

# --- STACK AND PREP DATA ---
stacked <- c(GUR_LO_Raster, GUR_Ordinal_Raster)
names(stacked) <- c("occurrence", "density")

vals_df <- as.data.frame(stacked, na.rm = TRUE)

# --- CHECK ASSUMPTIONS ---
qqnorm(vals_df$occurrence)
qqline(vals_df$occurrence, col = "blue")

qqnorm(vals_df$density)
qqline(vals_df$density, col = "red")


ggplot(vals_df, aes(x = density, y = occurrence)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "blue", se = FALSE) +
  theme_minimal() +
  labs(title = "Check for monotonicity between occurrence and relative density",
       x = "Relative Density",
       y = "Urchin LOC")


# 2. Check for excessive ties (count data tends to be rounded or integer-valued)
summary(as.factor(vals_df$occurrence))  # Optional: assess frequency of repeated values
length(unique(vals_df$occurrence))      # Fewer unique values = more ties (Spearman can tolerate some)

summary(as.factor(vals_df$density))  
length(unique(vals_df$density))   

# 3. Independence assumption: Visual inspection (no autocorrelation check here, just note)
# If you're concerned about spatial autocorrelation, Moran's I is an option 

# --- RUN SPEARMAN'S RANK CORRELATION TEST ---
spearman_result <- cor.test(vals_df$occurrence, vals_df$density, method = "spearman")
print(spearman_result)

GUR_RA_v_LOC <- spearman_result
```

#Red urchin 
```{r}
# --- RASTER ALIGNMENT ---
RUR_Ordinal_Raster <- resample(RUR_Ordinal_Raster, RUR_LO_Raster, method = "bilinear")  # Match resolution

# --- STACK AND PREP DATA ---
stacked <- c(RUR_LO_Raster, RUR_Ordinal_Raster)
names(stacked) <- c("occurrence", "density")

vals_df <- as.data.frame(stacked, na.rm = TRUE)

# --- CHECK ASSUMPTIONS ---
qqnorm(vals_df$occurrence)
qqline(vals_df$occurrence, col = "blue")

qqnorm(vals_df$density)
qqline(vals_df$density, col = "red")

ggplot(vals_df, aes(x = density, y = occurrence)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "gam", formula = y ~ s(x), color = "blue", se = FALSE) +
  theme_minimal() +
  labs(title = "Check for monotonicity between occurrence and relative density",
       x = "Relative Density",
       y = "Urchin LOC")


# 2. Check for excessive ties (count data tends to be rounded or integer-valued)
summary(as.factor(vals_df$occurrence))  # Optional: assess frequency of repeated values
length(unique(vals_df$occurrence))      # Fewer unique values = more ties (Spearman can tolerate some)

summary(as.factor(vals_df$density))  
length(unique(vals_df$density))   

# 3. Independence assumption: Visual inspection (no autocorrelation check here, just note)
# If you're concerned about spatial autocorrelation, Moran's I is an option 

# --- RUN SPEARMAN'S RANK CORRELATION TEST ---
spearman_result <- cor.test(vals_df$occurrence, vals_df$density, method = "spearman")
print(spearman_result)

RUR_RA_v_LOC <- spearman_result
```



```{r}
# Create a summary table of Spearman correlation test results

summary_table <- tibble(
  Species = c("Green Urchin", "Red Urchin"),
  
  `LOC vs Count (rho)` = c(
    round(GUR_LOC_v_count$estimate, 3),
    round(RUR_LOC_v_count$estimate, 3)
  ),
  `LOC vs Count (p-value)` = c(
    format.pval(GUR_LOC_v_count$p.value, digits = 3),
    format.pval(RUR_LOC_v_count$p.value, digits = 3)
  ),
  
  `Count vs RA (rho)` = c(
    round(GUR_RA_v_count$estimate, 3),
    round(RUR_RA_v_count$estimate, 3)
  ),
  `Count vs RA (p-value)` = c(
    format.pval(GUR_RA_v_count$p.value, digits = 3),
    format.pval(RUR_RA_v_count$p.value, digits = 3)
  ),
  
  `LOC vs RA (rho)` = c(
    round(GUR_RA_v_LOC$estimate, 3),
    round(RUR_RA_v_LOC$estimate, 3)
  ),
  `LOC vs RA (p-value)` = c(
    format.pval(GUR_RA_v_LOC$p.value, digits = 3),
    format.pval(RUR_RA_v_LOC$p.value, digits = 3)
  )
)

summary_table %>% 
  kable("html", caption = "Spearman Correlations Among Models by Urchin Species") %>% 
  kable_styling(full_width = FALSE, position = "center") %>% 
  column_spec(1, bold = TRUE, border_right = TRUE)



write.csv(summary_table, "./Generated_Data/Correlation_data/urchin_correlations_summary.csv", row.names = FALSE)

```





Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
