---
title: "Green Urchin LO"
author: "Matt Csordas" and "Tessa Rehill" 
date: "2024-12-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmarkdown file runs through the species distribution modeling process for Green urchins from dataset creation to final model evaluation and interpretation

GUR = Green urchin

#Install the flexsdm package
```{r}
# For Windows and Mac OS operating systems
#remotes::install_github("sjevelazco/flexsdm")
```

# Load all necessary packages
```{r}
library(tidyverse)
library(terra)
library(sf)
library(cowplot)
library(tidyterra)
library(ggspatial)
library(blockCV)
library(gbm)
library(dismo)
library(mop)
library(foreach)
library(doParallel)
library(caret)
library(CAST)
library(viridis)
library(ggExtra)
library(gridExtra)
library(flexsdm)
library(kernlab)
library(mgcv)
library(enmSdmX)
```

# Data preparation

## Load all initally required datasets
```{r}
#Load general functions and general spatial mapping plotting scripts 
source("General_Functions.R")
source("Spatial_Mapping_Functions_flexSDM.R")

#Load the BC buffer used to ensure all training data is contained within BC
BC_Buffer <- terra::vect("./raw_data/BC_Buffer/BC_Buffer.shp") %>% terra::project(terra::crs("epsg:3005"))

#Import the shapefile we will be using to represent land/the coastline and project to the correct crs/mask to NC
Coastline <- vect("./raw_data/BC_Buffer/GSHHS_f_L1.shp") %>% terra::project("epsg:3005")
Land <- Coastline %>% terra::mask(BC_Buffer)

#Load in the prepped presence data from DFO transect surveys
Urchin_Transects_Filtered <- sf::st_read("./processed_data/Urchin_Presence/Urchin_Occurrence_Transects_Filtered.shp")
colnames(Urchin_Transects_Filtered)[4] <- "Year"
colnames(Urchin_Transects_Filtered)[11] <- "Species"
```

## Create the final presence dataset for GUR
```{r}
#Load in the scripts containing functions we will use to do this
source("Presence_Point_Creator.R")
source("Presence_Map_Creator.R")

#Generate a list of years for which occurrence data is available for GUR
GUR_Year_list <- unique((Urchin_Transects_Filtered %>% dplyr::filter(Species == "GUR"))$Year)

#Load bathymetry layer
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#For each year in the GUR year list generate presence points, a presence raster, and a map of where presences were observed
for (i in 1:length(GUR_Year_list)){
  GUR_Output <- Point_Creator(Urchin_Transects_Filtered %>% dplyr::filter(Year == GUR_Year_list[i]), "GUR", Bathymetry)
  GUR_Presence_Points <- GUR_Output[[2]]
  GUR_Raster <- GUR_Output[[1]]
  
  terra::writeVector(GUR_Presence_Points, paste0("./processed_data/Urchin_Presence/GUR_Presence_Points_", GUR_Year_list[i],".shp"), overwrite = TRUE)
  terra::writeRaster(GUR_Raster, paste0("./processed_data/Urchin_Presence/GUR_Presence_Raster_", GUR_Year_list[i], ".tif"), overwrite = TRUE)
  GUR_Pres_Map <- Pres_Map_GUR(Land, GUR_Presence_Points %>% terra::project(crs(Land)))
  ggsave(paste0("./Plots/Maps/Species_Presence_Maps/GUR_Pres_Map_", GUR_Year_list[i], ".jpeg"), GUR_Pres_Map, units = "mm", width = 120, height = 140)
  print(i)
}

#Do the same for the entire time period 
GUR_Output <- Point_Creator(Urchin_Transects_Filtered, "GUR", Bathymetry)
GUR_Presence_Points <- GUR_Output[[2]]
GUR_Raster <- GUR_Output[[1]]
terra::writeVector(GUR_Presence_Points, "./processed_data/Urchin_Presence/GUR_Presence_Points.shp", overwrite = TRUE)
terra::writeRaster(GUR_Raster, "./processed_data/Urchin_Presence/GUR_Presence_Raster.tif", overwrite = TRUE)

GUR_Pres_Map <- Pres_Map_GUR(Land, GUR_Presence_Points %>% terra::project(crs(Land)))
ggsave("./Plots/Maps/Species_Presence_Maps/GUR_Pres_Map.jpeg", GUR_Pres_Map, units = "mm", width = 120, height = 140)
```

## Generate pseudo-absences for each year, associate all data points with their environmental conditions, and output the final training dataset
```{r}
gc()
#Load all of the required scripts containing functions to do this
source("Pseudo-absence_Creator_flexSDM.R")

#Set up lists to hold the environmental rasters/dataframes, survey locations, and species presence points/rasters for each year
Enviro_list <- list()
Enviro_df_list <- list()
Enviro_df_mask_list <- list()
Enviro_df_long_list <- list()
Buffer_list <- list()
GUR_point_list <- list()
GUR_rast_list <- list()

#Load the environmental layers, survey buffers, and species presence points from each relevant year of the hindcast. In the loop create masked environmental dataframes for each year of data based on a buffer around survey locations from that year.
for (i in 1:length(GUR_Year_list)){
  #Load the environmental layers for year i
Enviro_list[[i]] <- terra::rast(paste0("./processed_data/ENV_layers/BCCM_Rasters/Predictor_hindcast_climatology_", GUR_Year_list[i], ".tif"))[[names(terra::rast(paste0("./processed_data/ENV_layers/BCCM_Rasters/Predictor_hindcast_climatology_", GUR_Year_list[i], ".tif")))[!grepl("Depth", names(terra::rast(paste0("./processed_data/ENV_layers/BCCM_Rasters/Predictor_hindcast_climatology_", GUR_Year_list[i], ".tif"))))]]]

  #Load the survey buffers for year i
  Buffer_list[[i]] <- terra::vect(paste0("./processed_data/Survey_Buffers/Survey_Buffer_", GUR_Year_list[i], ".shp"))
  #Create an environmental dataframe for year i
   Enviro_df_list[[i]] <- terra::as.data.frame(Enviro_list[[i]], xy = TRUE) %>% na.omit()
  #Create a long-form version of the above dataframe where instead of a column for each predictor there is a column    denoting which values are for which predictor
  Enviro_df_long_list[[i]] <- Enviro_df_list[[i]] %>% pivot_longer(cols = everything(), names_to = "Type", values_to = "Value")
  Enviro_df_long_list[[i]]$DataType <- "BC-wide"
  #Create a masked environmental dataframe based on the survey locations in year i
  Enviro_df_mask_list[[i]] <- terra::as.data.frame(terra::mask(Enviro_list[[i]], Buffer_list[[i]]), xy=TRUE) %>% na.omit()
  #Load in species presence points and rasters
  GUR_point_list[[i]] <- terra::vect(paste0("./processed_data/Urchin_Presence/GUR_Presence_Points_", GUR_Year_list[i], ".shp")) %>% terra::mask(BC_Buffer)
  GUR_rast_list[[i]] <- terra::rast(paste0("./processed_data/Urchin_Presence/GUR_Presence_Raster_", GUR_Year_list[i], ".tif")) %>% terra::mask(BC_Buffer)
  gc()
}

#Generate pseudo-absence points for each year of data for GUR
#Read in the overall GUR Presence Raster
GUR_Pres_Rast <- terra::rast("./processed_data/Urchin_Presence/GUR_Presence_Raster.tif")

#Set up a list to hold the data
GUR_PA_list <- list()
#Generate PA_num sets of presence-pseudoabsence points for each year
#for (i in 1:length(GUR_Year_list)){
#  GUR_PA_list[[i]] <- PA_Creation(GUR_rast_list[[i]], GUR_point_list[[i]], GUR_Pres_Rast, Enviro_df_mask_list[[i]], dim(GUR_point_list[[i]])[1])
#}

#select PA outside of survey buffers 
for (i in 1:length(GUR_Year_list)){
  GUR_PA_list[[i]] <- PA_Creation(GUR_rast_list[[i]], GUR_point_list[[i]], GUR_Pres_Rast, Enviro_df_list[[i]], dim(GUR_point_list[[i]])[1])
}

```

## Associate presence/pseudo-absence datasets with the appropriate training predictors and combine data from across years into one final dataset for each PA dataset created
```{r}
#Associate all data with environmental data from the appropriate location and year
GUR_PA_Enviro_List <- list()
for (i in 1:length(GUR_PA_list)){
  GUR_PA_Enviro_List[[i]] <- terra::extract(x = Enviro_list[[i]], y = GUR_PA_list[[i]], method = "simple", bind = TRUE) %>% terra::na.omit("")
}

#Combine independent years of data together for each PA dataset so that there is one final dataset for each 
GUR_training_data <- GUR_PA_Enviro_List[[1]]
for (j in 2:length(GUR_PA_Enviro_List)){
  GUR_training_data <- rbind(GUR_training_data, GUR_PA_Enviro_List[[j]])
}



```

#Save as RDS (for inputs into the Alliance)
```{r}
#Save
# Save R objects to .rds files
saveRDS(Enviro_list, "outputs/GUR_Enviro_list.rds")
saveRDS(Enviro_df_list, "outputs/GUR_Enviro_df_list.rds")
saveRDS(Enviro_df_mask_list, "outputs/GUR_Enviro_df_mask_list.rds")
saveRDS(Enviro_df_long_list, "outputs/GUR_Enviro_df_long_list.rds")
saveRDS(Buffer_list, "outputs/GUR_Buffer_list.rds")
saveRDS(GUR_point_list, "outputs/GUR_point_list.rds")
saveRDS(GUR_rast_list, "outputs/GUR_rast_list.rds")
saveRDS(GUR_PA_list, "outputs/GUR_PA_list.rds")
saveRDS(GUR_PA_Enviro_List, "outputs/GUR_PA_Enviro_List.rds")
saveRDS(GUR_training_data, "outputs/GUR_training_data.rds")


# Load previously saved data objects
Enviro_list <- readRDS("outputs/GUR_Enviro_list.rds")
Enviro_df_list <- readRDS("outputs/GUR_Enviro_df_list.rds")
Enviro_df_mask_list <- readRDS("outputs/GUR_Enviro_df_mask_list.rds")
Enviro_df_long_list <- readRDS("outputs/GUR_Enviro_df_long_list.rds")
Buffer_list <- readRDS("outputs/GUR_Buffer_list.rds")
GUR_point_list <- readRDS("outputs/GUR_point_list.rds")
GUR_rast_list <- readRDS("outputs/GUR_rast_list.rds")
GUR_PA_list <- readRDS("outputs/GUR_PA_list.rds")
GUR_PA_Enviro_List <- readRDS("outputs/GUR_PA_Enviro_List.rds")
GUR_training_data <- readRDS("outputs/GUR_training_data.rds")
```


## Divide datasets into CV folds by spatial blocking
```{r}
gc()
#Load all of the required scripts containing functions to do this
source("CV_Functions_flexSDM.R")

#Set the size of spatial blocks based on previously completed autocorrelation analysis
Block_Size <-90000 #CODEEDIT to adjust block size for the species if needed 
#Set the k-value for k-fold cross-validation: number of spatial folds
k <- 10

#Load in hindcast climatology #CODEEDIT select for only the variables I want included in my models
Hindcast_Climatology <- terra::rast("./processed_data/ENV_layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif")
Hindcast_Climatology <- terra::subset(Hindcast_Climatology, names(Hindcast_Climatology) != "Depth")

#Turn the final training dataset into an sf object
GUR_training_data_sf <- sf::st_as_sf(GUR_training_data, coords = c("x", "y"), crs = 3005)

#Set seed
set.seed(157)
#Spatially block based on the optimal size suggested above
GUR_training_blocks <- blockCV::cv_spatial(GUR_training_data_sf, column = "Occurrence", hexagon = FALSE, k = k+1, size = Block_Size, selection = "random", iteration = 200)

#Add the training fold IDs to the dataset
GUR_training_data$fold <- GUR_training_blocks$folds_ids
GUR_training_data_sf$fold <- GUR_training_blocks$folds_ids

#Check to see how many presence points ended up in each block
Presence_summary <- terra::as.data.frame(GUR_training_data) %>% dplyr::group_by(fold) %>% dplyr::summarise(Presence_num = sum(Occurrence))

#Adjust folds if needed by reshuffling presence points from folds with a number of presence points over some cutoff based on the total number of points in the dataset
GUR_training_data_sf <- CV_redistribute_new(GUR_training_data_sf, k+1, 150) #adjust this cutoff to be reasonable based on the number of presence points and folds you are working with (the idea is to better balance presences within folds, but not to make things completely equal)
GUR_training_data <- terra::vect(GUR_training_data_sf)
GUR_training_data_df <- terra::as.data.frame(GUR_training_data)

#Save the dataset for GUR
write.csv(GUR_training_data_df, "./Generated_Data/Blocked_datasets/GUR_blocks.csv")
terra::writeVector(GUR_training_data, "./processed_data/Blocked_Spatial_Data/GUR_blocks.shp", overwrite = TRUE)

#Check to see how many presence points ended up in each block
Presence_summary <- GUR_training_data_df %>% dplyr::group_by(fold) %>% dplyr::summarise(Presence_num = sum(Occurrence))
Point_summary <- GUR_training_data_df %>% dplyr::group_by(fold) %>% dplyr::summarise(Point_num = n())

#Merge the two data frames by the 'fold' column
GUR_Point_numbers <- merge(Presence_summary, Point_summary, by = "fold")
#Create the 'Absence_num' column
#GUR_Point_numbers$Absence_num <- Point_numbers$Point_num - Point_numbers$Presence_num


GUR_training_data_df <- GUR_training_data_df[!is.na(GUR_training_data_df$fold), ]
GUR_training_data_sf <- GUR_training_data_sf[!is.na(GUR_training_data_sf$fold), ]

#Look at the CAST plot to see how well folds represent predicting to the entire predictor space
GUR_CAST_output <- CAST::geodist(x = GUR_training_data_sf, modeldomain = Hindcast_Climatology, type = "feature", cvfolds = GUR_training_data_sf$fold, samplesize = 2000, sampling = "random")
GUR_CAST_plot <- ggplot() + geom_density(data = GUR_CAST_output, aes(x = dist, fill = what), alpha = 0.4) + 
      theme_classic() + scale_fill_manual(values = c("#404788FF", "#238A8DFF", "#FDE725FF"), name = "Compared features") + 
      ylab("Density") + xlab("Distance in feature space") + 
      theme(legend.position = c(0.7,0.7)) + guides(x.sec = "axis", y.sec = "axis")+ 
      theme(axis.line.y.left = element_line(size = 0.75, color = "black"), axis.line.y.right = element_line(size = 0.75, color = "black"), axis.line.x.top = element_line(size = 0.75, color = "black"), axis.line.x.bottom = element_line(size = 0.75, color = "black"),axis.ticks.x.top = element_blank(), axis.ticks.y.right = element_blank(), axis.text.x.top = element_blank(), axis.text.y.right = element_blank())

GUR_CAST_plot
ggsave2(filename = "./Plots/CAST_Plots/GUR_CAST_Plot.jpeg", plot = GUR_CAST_plot, scale = 1, height = 75, width = 100, units = "mm")
```

## Plot the environmental and geographic coverage of the dataset
```{r}
gc()
#Load all scripts containing functions required for this step
source("Predictor_Coverage_Plotter_flexSDM.R")
source("PA_Map_Creator_flexSDM.R")

#Calculate the number of predictors from one of the environmental dataframes
Predictor_num <- ncol(Enviro_df_list[[1]]) 
Predictor_num <- Predictor_num - 2

#Plot the environmental coverage of presence and pseudo-absence datasets across all years and PA datasets compared to the overall environmental space
GUR_Predictor_Coverage_Plot <- Enviro_plot(Enviro_df_long_list, GUR_training_data_df, Predictor_num)
ggsave2(filename = "./Plots/Predictor_Coverage_Plots/GUR_Predictor_Plot.jpeg", plot = GUR_Predictor_Coverage_Plot, scale = 1, height = 400, width = 400, units = "mm")

#Geographically plot presences and pseudo-absences for each PA dataset (all years together on same plot)
GUR_PA_Map <- Geog_Plot_GUR(Land, GUR_training_data)
ggsave("./Plots/PA_Geog_Plots/GUR_PA_Geog_Plot.jpeg", GUR_PA_Map, units = "mm", width = 120, height = 140)

###
#Build density plots comparing all variable ranges within presence points of each data fold
###
Fold_Predictor_Data_long <- dplyr::filter(GUR_training_data_df, Occurrence == 1)[,2:(Predictor_num+2)] %>% pivot_longer(cols = TAlk_10m_mean:Slope, values_to = "Value", names_to = "Predictor")
Fold_Predictor_Data_long$fold <- as.factor(Fold_Predictor_Data_long$fold)

Predictor_density_plot <- ggplot(Fold_Predictor_Data_long, aes(x = fold, y = Value, fill = fold)) + geom_violin(alpha = 0.8) + theme_cowplot() + facet_wrap(Predictor~., scales = "free") + theme(legend.position = "none", axis.text.x = element_text(), legend.title = element_text(size = 32), legend.text = element_text(size =28)) + ylab(NULL) + xlab(NULL) + theme(strip.background = element_blank(), strip.placement = "outside", strip.text = element_text(size=14)) + scale_fill_manual(values = c("#bfbfbf","#e76bf3", "#11b0f7", "#09c07c", "#a2a500", "#ffa32c", "#f8766d", "pink", "brown", "white", "black"))
Predictor_density_plot
ggsave2(filename = "./Plots/Density_Plots/GUR_Fold_Predictor_density_plot.jpeg", plot = Predictor_density_plot, scale = 1, height = 400, width = 400, units = "mm")
```

## Split the spatially blocked data into the k training folds and the k+1 final testing dataset
```{r}


#Create a final training df from the first k folds
GUR_training_df <- GUR_training_data_df %>% dplyr::filter(fold != (k+1))

#Create a final testing df from the K+1 fold
GUR_testing_df <- GUR_training_data_df %>% dplyr::filter(fold == (k+1))


```

# Model Fitting

## Fit an ensemble SDM to the training data
```{r}
gc()

#BRT
BRT_tune_grid <- expand.grid(
  n.trees = c(1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800, 1900, 2000, 2100, 2200),
  shrinkage = c(0.05, 0.01, 0.005, 0.001, 0.0005),
  n.minobsinnode = c(5, 10, 15, 20, 40))

GUR_BRT <- flexsdm::tune_gbm(data = GUR_training_df, response = "Occurrence", predictors = colnames(GUR_training_data_df[2:(Predictor_num+1)]), partition = "fold", grid = BRT_tune_grid, thr = c("equal_sens_spec"), metric = "AUC", n_cores = 10)

#MaxEnt
Max_tune_grid <- expand.grid(
  regmult = c(0.1, 0.5, 1, 2, 3, 5),
  classes = c("l", "lq", "lqh"))

GUR_max <- flexsdm::tune_max(data = GUR_training_df %>% dplyr::filter(Occurrence == 1), response = "Occurrence", predictors = colnames(GUR_training_data_df[2:(Predictor_num+1)]), partition = "fold", background = GUR_training_df %>% dplyr::filter(Occurrence == 0), grid = Max_tune_grid, thr = c("equal_sens_spec"), metric = "AUC", clamp = FALSE, pred_type = "cloglog", n_cores = 10)

#GLM
GUR_glm <- flexsdm::fit_glm(data = GUR_training_df, response = "Occurrence", predictors = colnames(GUR_training_data_df[2:(Predictor_num+1)]), select_pred = TRUE, partition = "fold", thr = c("equal_sens_spec"), poly = 1, inter_order = 0)

#GAM
GUR_gam <- flexsdm::fit_gam(data = GUR_training_df, response = "Occurrence", predictors = colnames(GUR_training_data_df[2:(Predictor_num+1)]), select_pred = TRUE, partition = "fold", thr = c("equal_sens_spec"), k = 3)

# #RF
# RF_tune_grid <- expand.grid(
#   ntree = c(1000, 1100, 1200, 1300, 1400, 1500, 1600, 1700, 1800, 1900, 2000, 2100, 2200),
#   mtry = c(2, 3, 4, 5, 6, 7, 8, 9, 10, 15, 20))
# GUR_RF <- flexsdm::tune_raf(data = GUR_training_df, response = "Occurrence", predictors = colnames(GUR_training_data_df[2:(Predictor_num+1)]), partition = "fold", grid = RF_tune_grid, thr = c("equal_sens_spec"), metric = "AUC", n_cores = 10)
# 
# #SVM
# SVM_tune_grid <- expand.grid(
#   C = c(2, 4, 6, 10, 15),
#   sigma = c(0.001, 0.05, 0.1, 1, 3, 5, 10))
# GUR_svm <- flexsdm::tune_svm(data = GUR_training_df, response = "Occurrence", predictors = colnames(GUR_training_data_df[2:(Predictor_num+1)]), partition = "fold", grid = SVM_tune_grid, thr = c("equal_sens_spec"), metric = "AUC", n_cores = 10)

#Fit an ensemble of all methods
GUR_ENS <- flexsdm::fit_ensemble(models = list(GUR_BRT, GUR_max, GUR_glm, GUR_gam), ens_method = "meanw", thr = c("equal_sens_spec"), thr_model = "equal_sens_spec", metric = "AUC")

```

# Model Evaluation and Prediction

## For the final ensemble model, calculate metrics of model overfitting (Omission Error with a 10% omission threshold, and AUCdiff) and discriminatory ability (AUC), compare to null models.
```{r}
gc()
#Load all scripts containing functions required during this step
source("Model_Evaluation_flexSDM.R")
source("Model_Evaluation_Plots_flexSDM.R")

#create a dataframe of environmental conditions in cells within the survey buffer to be used when making random draws for null models using previously created Enviro_df_mask_list

#Survey_Enviro_df <- Enviro_df_mask_list[[1]]
#for (i in 2:length(Enviro_df_mask_list)){
#  Survey_Enviro_df_mask <- rbind(Survey_Enviro_df_mask, Enviro_df_mask_list[[i]])
#}

#write.csv(Survey_Enviro_df_mask, "./Generated_data/Interim_datasets/GUR_Survey_Enviro_df_mask.csv", row.names = FALSE)

Survey_Enviro_df <- Enviro_df_list[[1]]
for (i in 2:length(Enviro_df_list)){
  Survey_Enviro_df <- rbind(Survey_Enviro_df, Enviro_df_list[[i]])
}


#save 
write.csv(Survey_Enviro_df, "./Generated_data/Interim_datasets/GUR_Survey_Enviro_df.csv", row.names = FALSE)

GUR_Survey_Enviro_df <- Survey_Enviro_df

Survey_Enviro_df <- Survey_Enviro_df %>% dplyr::select(-x,-y)

#Evaluate models and 1000 null models for comparison
#When depth is NOT included Predictor + 2, when depth is included PRedictor num + 1 (in Model_Eval function needs to change)
GUR_Eval <- Model_Eval_NEW(GUR_ENS, GUR_training_df, GUR_testing_df, Survey_Enviro_df, GUR_BRT, GUR_max, GUR_glm, GUR_gam, Predictor_num)

#Save the model evaluation and null evaluation dataframes as CSV files
write.csv(GUR_Eval[[1]], "./Generated_Data/Evaluation_datasets/GUR_Eval.csv")
write.csv(GUR_Eval[[2]], "./Generated_Data/Evaluation_datasets/GUR_Null.csv")


#Plot the model and null model evaluation metrics
GUR_Eval_Plot <- Eval_Plot(GUR_Eval)
GUR_Eval_Plot
ggsave2(filename = "./Plots/Eval_Plots/GUR_Eval_Plot.jpeg", plot = GUR_Eval_Plot, scale = 1, height = 75, width = 100, units = "mm")

#Create and display the Standard Effect Size Plot for GUR
GUR_SES_Plot <- Effect_Size_Plot(GUR_Eval)
GUR_SES_Plot
ggsave2(filename = "./Plots/SES_Plots/GUR_SES_Plot.jpeg", plot = GUR_SES_Plot, scale = 1, height = 75, width = 100, units = "mm")


```

#plotting null comparison
```{r}
#GUR_Null <- read.csv("./Generated_Data/Evaluation_datasets/GUR_Null -FINAL.csv")
#GUR_Models <- read.csv("./Generated_Data/Evaluation_datasets/GUR_Eval.csv") 

#make null metrics long 
#Create long-form dataframes for the model and null model metrics for plotting
#GUR_Null_long <- GUR_Null %>% pivot_longer(cols = c("AUC_Train", "AUC_Test", "AUCdiff", "OErr"), names_to = "Metric", 
                                        #   values_to = "Value") %>% dplyr::select('Metric', 'Value')

#GUR_Model_long <- GUR_Models %>% pivot_longer(cols = c("AUC_Train", "AUC_Test", "AUCdiff", "OErr"), names_to = "Metric", 
#                                              values_to = "Value") %>% dplyr::select('Metric', 'Value')

# Set factor level order
#GUR_Null_long$Metric <- factor(GUR_Null_long$Metric, levels = c("AUC_Train", "AUC_Test", "AUCdiff", "OErr"))
#GUR_Model_long$Metric <- factor(GUR_Model_long$Metric, levels = c("AUC_Train", "AUC_Test", "AUCdiff", "OErr"))

#Plot the model and null model evaluation metrics
#GUR_Eval_Plot <- ggplot() + geom_boxplot(data = GUR_Null_long, aes(x = Metric, y = Value, fill = Metric)) + 
#  scale_fill_manual(values = c("#613f74", "#91B3B6", "#F4E76E", "#EDAF97"), labels = c("AUC train", "AUC test", "AUC diff", #"OErr")) + 
#  stat_summary(data = GUR_Model_long, aes(x = Metric, y = Value), fun.y = mean, geom = "point", colour = "#e6360F", size = 3) #+ 
 # xlab("Evaluation Metric") + ylab("Value") + theme_classic() + 
#  guides(x.sec = "axis", y.sec = "axis")+ theme(axis.line.y.left = element_line(size = 0.75, color = "black"), axis.line.y.right = element_line(size = 0.75, color = "black"), axis.line.x.top = element_line(size = 0.75, color = "black"), axis.line.x.bottom = element_line(size = 0.75, color = "black"),axis.ticks.x.top = element_blank(), axis.ticks.x.bottom = element_blank(), axis.ticks.y.right = element_blank(), axis.text.x.top = element_blank(), axis.text.y.right = element_blank()) + 
 # theme(legend.position = "none", legend.justification = "center") + 
#  theme(axis.text = element_text(size=8), axis.title = element_text(size = 12))


#GUR_Eval_Plot

#ggsave2(filename = "./Plots/Eval_Plots/GUR_Eval_Plot -FINAL.jpeg", plot = GUR_Eval_Plot, scale = 1, height = 75, width = 100, units = "mm")

```

## Generate final model likelihood of occurrence (LO) predictions and associated uncertainty
```{r}
gc()
#Load all scripts containing functions required during this step
source("Model_Prediction_flexSDM_V2.R")
source("Spatial_Mapping_Functions_flexSDM.R")

#Generate final LO predictions and uncertainty metrics
GUR_Outputs <- Prediction(GUR_ENS, GUR_training_df, Hindcast_Climatology, GUR_BRT, GUR_max, GUR_glm, GUR_gam, k, Predictor_num)

#Save the overall predictions and Uncertainty to raster layers for examination
terra::writeRaster(GUR_Outputs[[1]], "./processed_data/LO_Predictions/GUR_LO_Predictions.tif", overwrite = TRUE)
#terra::writeRaster(GUR_Outputs[[2]], "./processed_data/Uncertainty/GUR_Overall_Uncertainty.tif", overwrite = TRUE)
terra::writeRaster(GUR_Outputs[[2]], "./processed_data/Uncertainty/GUR_Method_Uncertainty.tif", overwrite = TRUE)
#terra::writeRaster(GUR_Outputs[[4]], "./processed_data/Uncertainty/GUR_CV_Uncertainty.tif", overwrite = TRUE)

#Generate maps of these metrics

#Create a map for the overall predictions as percent likelihood of occurrence for present day
GUR_Overall_Predicitions_Plot <- RC_Map(GUR_Outputs[[1]], "Likelihood of \nOccurence", "meanw", Land)
ggsave2(filename = "./Plots/Maps/LO_Plots/GUR_LO_Map.jpeg", plot = GUR_Overall_Predicitions_Plot, scale = 1, height = 140, width = 115, units = "mm")
#Create maps for each individual uncertainty metric
#GUR_Uncertainty_Plot <- RC_Map(GUR_Outputs[[2]], "Overall\nPrediction\nUncertainty", "std", Land)
#ggsave2(filename = "./Plots/Maps/Uncertainty_Plots/GUR_Overall_Uncertainty_Map.jpeg", plot = GUR_Uncertainty_Plot, scale = 1, height = 140, width = 115, units = "mm")
GUR_Method_Uncertainty_Plot <- RC_Map(GUR_Outputs[[2]], "Model\nUncertainty", "gbm", Land)
ggsave2(filename = "./Plots/Maps/Uncertainty_Plots/GUR_Model_Uncertainty_Map.jpeg", plot = GUR_Method_Uncertainty_Plot, scale = 1, height = 140, width = 115, units = "mm")

#GUR_CV_Uncertainty_Plot <- RC_Map(GUR_Outputs[[4]], "CV\nUncertainty", "std", Land)
#ggsave2(filename = "./Plots/Maps/Uncertainty_Plots/GUR_CV_Uncertainty_Map.jpeg", plot = GUR_CV_Uncertainty_Plot, scale = 1, height = 140, width = 115, units = "mm")
```

## Determine the optimal threshold for converting (LO) predictions to binary presence/absence and convert
```{r}
#Load all scripts containing functions required during this step
source("Binary_Prediction_flexSDM.R")

#Create an overall binary prediction layer
GUR_Thresholds <- Optimize_Threshold(GUR_training_data_df, GUR_BRT, GUR_max, GUR_glm, GUR_gam)
GUR_Thresholds[[1]]
#Save the Threshold Plot
ggsave2(filename = "./Plots/Threshold_Plots/GUR_Threshold_Plot.jpeg", plot = GUR_Thresholds[[1]], scale = 1, height = 100, width = 150, units = "mm")

#For GUR it appears that the optimal threshold is the optimal TSS threshold as this also keeps an omission error of 0 for the overall model
GUR_Optimal_Threshold <- GUR_Thresholds[[2]]

#Save the threshold data to a csv file for use when adjusting thresholds later
write.csv(GUR_Thresholds[[9]], "./Generated_Data/Threshold_datasets/GUR_Thresholds.csv")
write.csv(GUR_Thresholds[[4]], "./Generated_Data/Prediction_datasets/GUR_Predictions.csv")

#Read the predictions back in for use in plotting
GUR_Prediction_dataset <- read.csv("./Generated_Data/Prediction_datasets/GUR_Predictions.csv")

#Create a density plot of false TP, FP, TN, FN
GUR_Density_Plot <- Density_Plot(GUR_Prediction_dataset, GUR_Optimal_Threshold)
GUR_Density_Plot
ggsave2(filename = "./Plots/Density_Plots/GUR_Density_Plot.jpeg", plot = GUR_Density_Plot, scale = 1, height = 150, width = 225, units = "mm")
#Create the final binary prediction layer
GUR_Binary_Predictions <- GUR_Outputs[[1]] >= GUR_Optimal_Threshold
#Save this raster 
terra::writeRaster(GUR_Binary_Predictions, "./processed_data/Binary_Predictions/GUR_Binary_Predictions.tif", overwrite = TRUE)
gc()

#Generate a map of the final binary predictions
#Create a map of present day PA predictions 
GUR_PA_Plot <- RD_Map_GUR(GUR_Binary_Predictions, "Green \nurchin", Land)
ggsave2(filename = "./Plots/Maps/Binary_Occurence_Plots/GUR_Binary_Predictions_Map.jpeg", plot = GUR_PA_Plot, scale = 1, height = 140, width = 115, units = "mm")
```

# Model/Predictor Interpretation

## Assess variable imporatance using the evaluation strip method and relative importance 
```{r}
#Load all scripts containing functions required during this step
source("Predictor_Importance_flexSDM.R")
source("Predictor_Evaluation_Strip_flexSDM.R")

###
#Variable Importance
###

#Calculate relative predictor importance and generate plots representing this and aggregated importance for larger predictor classes
GUR_Pred_Imp <- Predictor_Importance_GUR(GUR_ENS, GUR_training_df, Predictor_num)

#Save a copy of the predictor importance dataframe in case needed later
write.csv(GUR_Pred_Imp[[1]],"./Generated_Data/Predictor_Importance_datasets/GUR_Predictor_Importance.csv")
#Save the plots
ggsave2(filename = "./Plots/PI_Plots/GUR_PI_Plot.jpeg", plot = GUR_Pred_Imp[[2]], scale = 1, height = 200, width = 175, units = "mm")
ggsave2(filename = "./Plots/PI_Plots/GUR_Class_PI_Plot.jpeg", plot = GUR_Pred_Imp[[3]], scale = 1, height = 200, width = 175, units = "mm")
gc()

###
#Set up for evaluation strip
###

#Set up dataframes for performing the evaluation strip method for any Urchin species 
Hindcast_df <- terra::as.data.frame(terra::rast("./processed_data/ENV_Layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif")) %>% na.omit() 

Hindcast_df <- Hindcast_df[, !names(Hindcast_df) %in% "Depth"]


gc()
#Calculate the means for each environmental variable
Enviro_Means <- colMeans(Hindcast_df)
#Calculate the max and min for each variable
Enviro_Max <- colMax(Hindcast_df)
Enviro_Min <- colMin(Hindcast_df)

#Create a dataframe for prediction to occur onto for each of the predictor variables of interest.
#Set up a base dataframe
Evaluation_Strip_Base <- data.frame()
Evaluation_Strip_Base_Max <- data.frame()
Evaluation_Strip_Base_Min <- data.frame()
for (i in 1:1000){
  Evaluation_Strip_Base[i,1:ncol(Hindcast_df)] <- Enviro_Means
  Evaluation_Strip_Base_Max[i,1:ncol(Hindcast_df)] <- Enviro_Max
  Evaluation_Strip_Base_Min[i,1:ncol(Hindcast_df)] <- Enviro_Min
}
#name the dataframe columns
colnames(Evaluation_Strip_Base) <- colnames(Hindcast_df)
colnames(Evaluation_Strip_Base_Max) <- colnames(Hindcast_df)
colnames(Evaluation_Strip_Base_Min) <- colnames(Hindcast_df)

#Set up a dataframe where values ranging the full environmental conditions for each variable are stored
Evaluation_Strip_Range <- Evaluation_Strip_Base
for (i in 1:ncol(Evaluation_Strip_Range)){
  Evaluation_Strip_Range[,i] <- seq(from = Enviro_Min[i], to = Enviro_Max[i], length.out = 1000)
}

###
#Evaluation Strip
###

#Create predictions for each variable all others held at their mean, max, and min value using evaluation strip
GUR_ES_Preds <- ES_Predictions(Hindcast_df, Evaluation_Strip_Base, Evaluation_Strip_Base_Max, Evaluation_Strip_Base_Min, Evaluation_Strip_Range, GUR_BRT, GUR_max, GUR_glm, GUR_gam)
#Save ES predictions for later use
write.csv(GUR_ES_Preds,"./Generated_Data/ES_datasets/GUR_ES_Preds.csv")

#Define a list of names for all of the predictor variables to be used by the evaluation strip plotting function
Predictor_Names <- c("Talkbmax (mEq/m^3)", "Sbmean (PSU)", "Sbmin (PSU)","Tsmean (\u00B0C)", "Tsmax (\u00B0C)", "Tsmin (\u00B0C)", "Tbmean (\u00B0C)", "Tbmax (\u00B0C)", "Tbmin (\u00B0C)", "Oxybmean (mmol/m^3)", "Oxybmin (mmol/m^3)", "Current (m/s)", "REI", "Percent Rock", "Slope (deg)") #CODEEDIT: change this based on the predictors being used
Predictor_IDs <- names(Hindcast_Climatology)
#Create all of the evaluation strip plots for my environmental variables
GUR_ES_Plots <- list()
GUR_ES_Plots_All <- list()
for(i in 1:length(Predictor_IDs)){
  ES_Plots_Temp <- ES_Plotter(GUR_ES_Preds, Predictor_IDs[i], Predictor_Names[i])
  ggsave2(filename = paste0("./Plots/ES_Plots/GUR/GUR_", Predictor_IDs[i],"_ES_Plot.jpeg"), plot = ES_Plots_Temp[[1]], scale = 1, height = 150, width = 225, units = "mm")
  ggsave2(filename = paste0("./Plots/ES_Plots/GUR/GUR_", Predictor_IDs[i],"_All_ES_Plot.jpeg"), plot = ES_Plots_Temp[[2]], scale = 1, height = 150, width = 225, units = "mm")
  GUR_ES_Plots[[i]] <- ES_Plots_Temp[[1]]
  GUR_ES_Plots_All[[i]] <- ES_Plots_Temp[[2]]
}

#Create a plot of all ES Plots
GUR_ES_Plot_Grid <- do.call(grid.arrange, c(GUR_ES_Plots, list(ncol = 5)))
ggsave2(filename = "./Plots/ES_Plots/GUR_ES_Plot.jpeg", plot = GUR_ES_Plot_Grid, scale = 1, height = 450, width = 500, units = "mm")

gc()
```

## Create predictor masks for GUR current day predictions based on where predictors are extrapolated beyond the training dataset
```{r}
#Create layers of a continuous extrapolation score and a binary yes/no of extrapolation
GUR_Extrap <- extra_eval(training_data = GUR_training_df, pr_ab = "Occurrence", projection_data = Hindcast_Climatology, univar_comb = TRUE, n_cores = 1)
#Adjust these to have NA values where they should
Mask_Mult <- Hindcast_Climatology$REI*0+1
GUR_Extrap_Continuous <- GUR_Extrap$extrapolation$extrapolation*Mask_Mult
GUR_Extrap_Discrete <- GUR_Extrap$uni_comb$uni_comb*Mask_Mult
#Save the layers
terra::writeRaster(GUR_Extrap_Continuous, "./processed_data/Extrapolation_Masks/GUR_Continuous_Mask.tif", overwrite = TRUE)
terra::writeRaster(GUR_Extrap_Discrete, "./processed_data/Extrapolation_Masks/GUR_Discrete_Mask.tif", overwrite = TRUE)

#Create maps of these extrapolation masks
GUR_Continuous_Mask_Plot <- RC_Map(GUR_Extrap_Continuous, "Degree of\nExtrapolation", "extrapolation", Land)
ggsave2(filename = "./Plots/Maps/Mask_Plots/GUR_Continuous_Mask_Map.jpeg", plot = GUR_Continuous_Mask_Plot, scale = 1, height = 140, width = 115, units = "mm")
GUR_Discrete_Mask_Plot <- ggplot(Land) + geom_sf() + geom_raster(data = as.factor(GUR_Extrap_Discrete), mapping = aes(x = x, y=y, fill = uni_comb)) + coord_sf(crs = st_crs(3005)) + annotation_scale(location = "bl", width_hint = 0.5) + annotation_north_arrow(location = "bl", which_north = "true", pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"), style = north_arrow_fancy_orienteering) + theme_cowplot() + xlim(c(650000, 890000)) + ylim(c(780000, 1225000)) + scale_fill_manual(values = c("#613f74","#F4E76E"), na.value = NA, name = element_blank(), labels = c("Not Extrapolated","Extrapolated"), na.translate = F) + xlab(element_blank()) + ylab(element_blank()) + theme(legend.position = c(0.6,0.9), legend.title=element_text(size=11), legend.text = element_text(size=10), legend.text.align = 0, legend.title.align = 0) + 
theme(panel.background = element_rect(fill = "#878c8c"))
ggsave2(filename = "./Plots/Maps/Mask_Plots/GUR_Discrete_Mask_Map.jpeg", plot = GUR_Discrete_Mask_Plot, scale = 1, height = 140, width = 115, units = "mm")
```
