---
title: "Red_Urchin_Count_Final"
author: "Tessa Rehill"
date: "2025-04-18"
output: html_document
---

This script runs abundance models for green urchin using biomod2 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install packages
```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(terra)
library(sf)
library(ggspatial)
library(cowplot)
library(biomod2)
library(ggtext)
library(usdm)
library(gridExtra)
library(ggsci)
library(blockCV)

```

#Call necessary functions 

```{r}
source("Count_Mapping_Functions.R")
source("CV_Functions_flexSDM.R")
```

#Creating count dataset
```{r}
#Import the mask which will be important for masking any unmasked data at this point
BC_Mask <- terra::vect("./raw_data/BC_Buffer/BC_Buffer.shp") %>% terra::project("epsg:3005")

#Import the shapefile we will be using to represent land/the coastline and project to the correct crs/mask to NC
Coastline <- vect("./raw_data/BC_Buffer/GSHHS_f_L1.shp") %>% terra::project("epsg:3005")
Land <- Coastline %>% terra::mask(BC_Mask)

#Load in the prepped Count data 
Urchin_Transects_Filtered <- sf::st_read("./processed_data/Urchin_Count/Urchin_Count_Transects_Filtered.shp")
```


##Create red urchin shapefiles
```{r}

selected_species <- "RUR"

filtered_data <- Urchin_Transects_Filtered %>% filter(species == selected_species)

RUR_Year_list <- unique((filtered_data %>% dplyr::filter(species == selected_species))$year)
# Load bathymetry layer
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")


#For each year in the RUR year list generate Count points, a Count raster, and a map of where Counts were observed
for (i in 1:length(RUR_Year_list)){
  Shapefile_Output <- Shapefile_Creator(filtered_data %>% dplyr::filter(year == RUR_Year_list[i]), Bathymetry, selected_species)
  terra::writeVector(Shapefile_Output, paste0("./processed_data/Urchin_Count/RUR_Count_Points_", RUR_Year_list[i],".shp"), overwrite = TRUE)
}


```

## Create point and maps for RUR
```{r}
#Generate a list of years for which occurrence data is available for NT
RUR_Year_list <- unique((Urchin_Transects_Filtered %>% dplyr::filter(species == "RUR"))$year)

#Load bathymetry layer
Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#For each year in the RUR year list generate Count points, a Count raster, and a map of where Counts were observed
for (i in 1:length(RUR_Year_list)){
  RUR_Output <- Point_Creator(Urchin_Transects_Filtered %>% dplyr::filter(year == RUR_Year_list[i]), "RUR", Bathymetry)
  RUR_Count_Points <- RUR_Output[[2]]
  RUR_Raster <- RUR_Output[[1]]
  terra::writeVector(RUR_Count_Points, paste0("./processed_data/Urchin_Count/RUR_AD_Points_", RUR_Year_list[i],".shp"), overwrite = TRUE)
  terra::writeRaster(RUR_Raster, paste0("./processed_data/Urchin_Count/RUR_AD_Raster_", RUR_Year_list[i], ".tif"), overwrite = TRUE)
  RUR_Pres_Map <- Pres_Map(Land, RUR_Count_Points %>% terra::project(crs(Land)))
  ggsave(paste0("./Plots/Maps/Species_Count_Maps/RUR_AD_Map_", RUR_Year_list[i], ".jpeg"), RUR_Pres_Map, units = "mm", width = 120, height = 140)
}

#Do the same for the entire time period 
RUR_Output <- Point_Creator(Urchin_Transects_Filtered, "RUR", Bathymetry)
RUR_Count_Points <- RUR_Output[[2]]
RUR_Raster <- RUR_Output[[1]]
terra::writeVector(RUR_Count_Points, "./processed_data/Urchin_Count/RUR_AD_Points.shp", overwrite = TRUE)
terra::writeRaster(RUR_Raster, "./processed_data/Urchin_Count/RUR_AD_Raster.tif", overwrite = TRUE)

RUR_Count_Points <- terra::vect("./processed_data/Urchin_Count/RUR_AD_Points.shp")
RUR_Pres_Map <- Pres_Map(Land, RUR_Count_Points %>% terra::project(crs(Land)))

ggsave("./Plots/Maps/Species_Count_Maps/RUR_AD_Map.jpeg", RUR_Pres_Map, units = "mm", width = 120, height = 140)
```

#BIOMOD Modelling
current version 10 

#Load in data for the Alliance 
```{r}
Land <- terra::vect("./raw_data/BC_Buffer/Land.shp")
```

#Format data to make dataframe of density points per year per environment
```{r}
Urchin_Transects_Filtered <- sf::st_read("./processed_data/Urchin_Count/Urchin_Count_Transects_Filtered.shp")
colnames(Urchin_Transects_Filtered)[4] <- "Year"
colnames(Urchin_Transects_Filtered)[11] <- "Species"


RUR_Transects <- terra::subset(Urchin_Transects_Filtered, Species == "RUR")
RUR_Year_list <- unique(RUR_Transects$Year)

#Set up dataframes to hold 
resp_df <- data.frame(mean = numeric())
env_df <- data.frame(
  ID = numeric(),
  TAlk_10m_mean = numeric(),
  salt_10m_mean = numeric(),
  salt_10m_min = numeric(),
  temp_s_mean = numeric(),
  temp_s_max = numeric(),
  temp_s_min = numeric(),
  temp_10m_mean = numeric(),
  temp_10m_max = numeric(),
  temp_10m_min = numeric(),
  do_10m_mean = numeric(),
  do_10m_min = numeric(),
  Current = numeric(),
  REI = numeric(),
  Rock = numeric(),
  Slope = numeric(),
  Depth = numeric()
)


#Create an empty data frame to store x,y
xy <- data.frame(
  x = numeric(),
  y = numeric()
)


for (i in 1:length(RUR_Year_list)){
  #Load the environmental layers for year i
  env.var <- terra::rast(paste0("./processed_data/ENV_layers/BCCM_Rasters/Predictor_hindcast_climatology_", RUR_Year_list[i], ".tif"))
  
  #Load shapefile for year i 
  resp.var <- terra::vect(paste0("./processed_data/Urchin_Count/RUR_Count_Points_", RUR_Year_list[i], ".shp"))
  
  #Rasterize shapefile, using env raster, take average density
  resp_raster <- terra::rasterize(resp.var, env.var, field = "density", fun = "mean") #average density per cell, creates raster
  
  #Resample shapefile, to env raster
  resp_raster <- terra::resample(resp_raster, env.var) #align resolutions and extents to environmental raster
  
  #Convert raster to point (using as.points) 
  points <- terra::as.points(resp_raster) #now each point has a location and averaged density value 
  points_df <- as.data.frame(points)
  
  
  #Extract coordinates of shapefile (using geom)
  coordinates <- terra::geom(points, df=TRUE) %>% dplyr::select(x, y)
  
  #Extract environmental variables for these coordinates 
  env_values <- terra::extract(env.var, points)
  
  #Append points_df to resp_df
  resp_df <- rbind(resp_df, points_df)
  
  #Append env_values to env_df
  env_df <- rbind(env_df, env_values)
  
  #Append coordinate to xy dataframe using rbind
  xy <- rbind(xy, coordinates)
}


#create full training dataframe 
print("Finished: Create dataframes BY year")
```
#Plot dataset 
```{r}
Land <- terra::vect("./raw_data/BC_Buffer/Land.shp") 
 # Convert land to sf
  land_sf <- st_as_sf(Land)
  
map_df <- cbind(resp_df, xy)# %>%
  #filter(mean < 100)


RUR_Count_Map <- ggplot() +    
  geom_sf(data = land_sf, fill = "grey80", color = NA) + 
  geom_point(data = map_df, aes(x = x, y = y, size = mean, alpha = mean), color = "#e6360F") + 
  scale_size_continuous(range = c(.1, 1.5), name = "Mean Density (m²)") + 
  scale_alpha_continuous(range = c(0.1, 0.5), name = "Mean Density (m²)") + 
  annotation_scale(location = "bl", width_hint = 0.5) + 
  annotation_north_arrow(location = "bl", which_north = "true",
                         pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                         style = north_arrow_fancy_orienteering) +    
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +    
  theme_cowplot() +    
  xlab(NULL) +    
  ylab(NULL) +    
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),  # Remove default x/y axis lines
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),  # Clean black box
    legend.position = c('top'),
    legend.background = element_rect(fill = "white", color = NA)  # No box around legend
  )

RUR_Count_Map

ggsave(
  filename = "./processed_data/Urchin_Count/RUR_CountDataset_Map.png",   # or another path/filename
  plot = RUR_Count_Map,
  width = 6,        
  height = 8,       
  dpi = 300         
)
```

#Remove collinear environmental variables
```{r}
env_df <- env_df[, !(colnames(env_df) %in% c("ID"))]

predictors <- env_df %>% na.omit()
varCor <- cor(predictors, use = "na.or.complete")
corrplot::corrplot(varCor)
Corr_Plot <- recordPlot()


#identify collinear/correlated variables and exclude them 
vif <- vifstep(env_df, th = 10, method = 'pearson') #checks collinearity based on VIF 
vif

#4 variables from 15 have collinearity problems (do_10m_mean, temp_s_mean ,temp_10m_max, salt_10m_min)
#keep do_10m_mean because most important predictor in occurence models 
env_df <- env_df[, !(colnames(env_df) %in% c("temp_s_mean", "salt_10m_mean", "temp_10m_max"))] 

#exclude these variables from our raster 
#env_df <- exclude(env_df, vif)

cor <- vifcor(env_df, th=0.7, method = 'pearson') #identify corelated variables
cor

#2 variables = do_10m_mean, salt_10m_min

#exclude these variables from our df (we keep both here)
#new_expl <- exclude(new_expl, cor) 

summary(env_df)
names(env_df)

#Read in projection layers, and remove the same variables

proj.var <- rast("./processed_data/ENV_layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif") 

proj.var <- proj.var[[!names(proj.var) %in% c("temp_s_mean", "temp_10m_max", "salt_10m_mean")]] 

names(proj.var)

print("Finished: Running collinearity") 
```

#Create final dataset (of all data)
```{r}
#Create final training dataset
RUR_training_data <- cbind(xy, resp_df, env_df) %>% na.omit() #includes all data, remove NA
#3275 observations


#Create cross validation fold table


#Set the size of spatial blocks based on previously completed autocorrelation analysis
Block_Size <-90000 #CODEEDIT to adjust block size for the species if needed 
#Set the k-value for k-fold cross-validation: number of spatial folds
k <- 10

#Load in hindcast climatology #CODEEDIT select for only the variables I want included in my models
Hindcast_Climatology <- terra::rast("./processed_data/ENV_layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif")

Hindcast_Climatology <- terra::subset(Hindcast_Climatology, names(Hindcast_Climatology) != c("temp_s_mean", "temp_10m_max", "salt_10m_min"))

#Turn the final training dataset into an sf object
RUR_training_data_sf <- sf::st_as_sf(RUR_training_data, coords = c("x", "y"), crs = 3005)

#Set seed
set.seed(157)
#Spatially block based on the optimal size suggested above
RUR_training_blocks <- blockCV::cv_spatial(RUR_training_data_sf, column = "mean", hexagon = FALSE, k = k+1, size = Block_Size, selection = "random", iteration = 200)

RUR_training_data$fold <- RUR_training_blocks$folds_ids
RUR_training_data_sf$fold <- RUR_training_blocks$folds_ids


Point_summary <-RUR_training_data %>%
  group_by(fold, mean) %>%
  summarise(count = n()) %>%
  arrange(fold, desc(mean))

# Categorize the mean values
plot_data <- RUR_training_data %>%
  mutate(mean_cat = case_when(
    mean == 0 ~ "none",
    mean > 0 & mean <= 10 ~ "few",
    mean > 10 & mean <= 20 ~ "many",
    mean > 20 ~ "abundant"
  )) %>%
  group_by(fold, mean_cat) %>%
  summarise(count = n(), .groups = "drop")

# Plot
plot <- ggplot(plot_data, aes(x = factor(fold), y = count, fill = mean_cat)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Fold", y = "Count", fill = "Abundance Category",
       title = "Abundance Categories per Fold") +
  scale_fill_manual(values = c("none" = "grey70", "few" = "lightblue", "many" = "orange", "abundant" = "darkred")) +
  theme_minimal()

plot

ggsave("./Plots/Count_models/RURcount_point_summary.png", width = 8, height = 6, dpi = 300)


#Check to see how many  points ended up in each block
Point_summary <- terra::as.data.frame(RUR_training_data) %>%
  dplyr::group_by(fold) %>%
  dplyr::summarise(Point_num = dplyr::n())

write.csv(Point_summary, "./Generated_Data/Interim_datasets/RURcount_point_summary.csv")

#Adjust folds if needed by reshuffling presence points from folds with a number of presence points over some cutoff based on the total number of points in the dataset
#RUR_training_data_sf <- CV_redistribute_count(RUR_training_data_sf, k+1, 80) #adjust this cutoff to be reasonable based on the number of presence points and folds you are working with (the idea is to better balance presences within folds, but not to make things completely equal)

RUR_training_data_vec <- terra::vect(RUR_training_data_sf)
RUR_training_data_df <- terra::as.data.frame(RUR_training_data)

#Save the dataset for RUR
write.csv(RUR_training_data_df, "./Generated_Data/Blocked_datasets/RURcount_blocks.csv")
terra::writeVector(RUR_training_data_vec, "./processed_data/Blocked_Spatial_Data/RUR_count_blocks.shp", overwrite = TRUE)

#Check to see how many presence points ended up in each block
Point_summary <- RUR_training_data_df %>% dplyr::group_by(fold) %>% dplyr::summarise(Point_num = dplyr::n())
Point_summary


#ggsave2(filename = "./Plots/Alliance/RURcount_CAST_Plot.jpeg", plot = RUR_CAST_plot, scale = 1, height = 75, width = 100, units = "mm")
print("Finished: Cross Validation Set Up")
```

#Create final training and evaluation dataset
```{r}
#Create a final training df from the first k folds
RUR_training_df <- RUR_training_data_df %>% dplyr::filter(fold != (3))

train_resp_df <- RUR_training_df$mean
train_xy <- RUR_training_df %>% dplyr::select(x, y)
train_env <- RUR_training_df %>% dplyr::select(TAlk_10m_mean, salt_10m_min, temp_s_max, temp_s_min, temp_10m_mean, temp_10m_min, do_10m_mean, do_10m_min, Current, REI, Rock, Slope, Depth)

#Create a final evaluation df from the K+1 fold
RUR_evaluation_df <- RUR_training_data_df %>% dplyr::filter(fold == (3))

eval_resp_df <- RUR_evaluation_df$mean
eval_xy <- RUR_evaluation_df %>% dplyr::select(x, y)
eval_env <- RUR_evaluation_df %>% dplyr::select(TAlk_10m_mean, salt_10m_min, temp_s_max, temp_s_min, temp_10m_mean, temp_10m_min, do_10m_mean, do_10m_min, Current, REI, Rock, Slope, Depth)

print("Finished: Creating final training and evaluation datasets")
```


#Format data with biomod
```{r}
#define species to model (ch)
sp <- c('RUR')

#define models to run 
all_models <- c("GLM","GBM","RF", "XGBOOST" ) #no GBM for ordinal data, no MAXENT for any abundance

#format
myBiomodData <- BIOMOD_FormatingData(resp.var = train_resp_df, #as SpatVector without associated data, for a single species
                                     expl.var = train_env, #SpatRaster containing explanatory variables in layers
                                     resp.name = sp, #character corresponding to species name
                                     resp.xy = train_xy, #if resp.var is vector, 2 column matrix associated xy coordinates
                                     dir.name = c('./models_biomod/count'), 
                                     data.type = c("abundance"), #can also be 'ordinal'
                                     eval.expl.var = eval_env, #SpatRaster without associated data, independent data to evaluate
                                     eval.resp.var = eval_resp_df, #SpatVector without associated data, independent data to evaluate 
                                     eval.resp.xy = eval_xy, #if resp.vect is vector, 2 column matrix of xy coordinates
                                     na.rm = TRUE, 
                                     filter.raster = TRUE, #filter points in same cell
                                     seed.val = 1234
) 


print("Finished: BIOMOD formatting data function")
    
```

#Create CV method 
```{r}
#using this cv strategy
cv <- bm_CrossValidation(bm.format = myBiomodData,
                         strategy = "kfold", #options are random, k-fold, strat, block, env, or user-defined
                         nb.rep = 1, #integer, number of repetitions of CV points to be drawn 
                         k = 10, #number of partitions                                         
                         do.full.models = TRUE #calibrated and validated over the whole dataset (and pseudo-absence datasets) or not
) 


```

#Model tuning
```{r}
#Tuning models and setting model options
opt.d <- bm_ModelingOptions(data.type = 'abundance', #character corresponding to data type to be used (see below)
                            models = all_models,
                            strategy = 'default',
                            bm.format = myBiomodData)

#boosted regression tree tuning 
tuned.gbm <- bm_Tuning(model = "GBM",
                       tuning.fun = "gbm", 
                       bm.options = opt.d@options$GBM.abundance.gbm.gbm,
                       do.formula = TRUE,
                       bm.format = myBiomodData,   
                       #calib.lines = cv,
                       metric.eval = c("Rsquared"), 
                       params.train = list(GBM.n.trees = c(1000, 2000, 3000),
                                           GBM.interaction.depth = c(3, 5, 7), 
                                           GBM.shrinkage = c(0.001, 0.005, 0.01),
                                           GBM.n.minobsinnode = c(5, 10))
)

print("Finished: GBM tuning")

#tuning GAM model 
#tuned.gam <- bm_Tuning(model = 'GAM',
#                       tuning.fun = 'gam', ## see in ModelsTable
#                       do.formula = TRUE,
#                       bm.options = opt.d@options$GAM.abundance.mgcv.gam,
#                       bm.format = myBiomodData,
#                       metric.eval=c("Rsquared"),
#                       #calib.lines = cv,
#                       params.train = list(GAM.select =c('TRUE'), GAM.method = c('GCV.Cp'), GAM.knots=c(3)))

#print("Finished: GAM tuning")

#Tuning CTA
#tuned.cta <- bm_Tuning(model = 'CTA',
#                       tuning.fun = 'rpart', ## see in ModelsTable
#                       do.formula = TRUE,
#                       bm.options = opt.d@options$CTA.abundance.rpart.rpart,
#                       bm.format = myBiomodData,
#                       metric.eval=c("Rsquared"),
                       #calib.lines = cv,
#                       params.train = list(CTA.maxdepth =c(5,10,20)))

#print("Finished: CTA tuning")


#Tuning RF
tuned.rf <- bm_Tuning(
  model = 'RF',
  tuning.fun = 'rf',
  do.formula = TRUE, #not sure if this should be true?, when run true = throws error, also TRUE = slower
  bm.options = opt.d@options$RF.abundance.randomForest.randomForest, #should be an object returned bybm_ModellingOptions,
  bm.format = myBiomodData, 
  #calib.lines = cv,
  metric.eval = c("Rsquared"), #Recall, Precision, and F1
  params.train = list(RF.mtry=c(2, 3, 4, 5))
)

print("Finished: RF tuning")

#Tuning XGBOOST 
tuned.xgboost <- bm_Tuning(
  model = 'XGBOOST',
  tuning.fun = 'xgbTree',
  do.formula = TRUE, #not sure if this should be true?, when run true = throws error, also TRUE = slower
  bm.options = opt.d@options$XGBOOST.abundance.xgboost.xgboost, #should be an object returned bybm_ModellingOptions,
  bm.format = myBiomodData, 
  #calib.lines = cv,
  metric.eval = c("Rsquared"), #Recall, Precision, and F1
  params.train = list(XGBOOST.nrounds = c(3,4,5), XGBOOST.max_depth =c(2,3,4), XGBOOST.eta =c(0.01,0.01, 1), XGBOOST.gamma=c(0,0.1)))

#increase gamma if you're seeing overfitting
#can also change colsample_bytree - fraction of features used per tree, lower = better for overfitting
#subsample - fraction of rows used per tree, prevents overfitting

print("Finished: XGBOOST tuning")

#tuned.mars <- bm_Tuning(
 # model = 'MARS',
 # tuning.fun = 'earth',
 # do.formula = TRUE, #not sure if this should be true?, when run true = throws error, also TRUE = slower
 # bm.options = opt.d@options$MARS.abundance.earth.earth, #should be an object returned bybm_ModellingOptions,
 # bm.format = myBiomodData, 
  #calib.lines = cv,
 # metric.eval = c("Rsquared"), #Recall, Precision, and F1
 # params.train = list(MARS.degree = c(1,2), MARS.nprune =c(20,30,40)))

#increase gamma if you're seeing overfitting
#can also change colsample_bytree - fraction of features used per tree, lower = better for overfitting
#subsample - fraction of rows used per tree, prevents overfitting

#print("Finished: MARS tuning")



#Setting tuning options
opt.t <- bm_ModelingOptions(data.type = 'abundance',
                            models = all_models,
                            strategy = "user.defined",
                            user.val = list(#GAM.nonbinary.mgcv.gam = tuned.gam,
                                            GBM.abundance.gbm.gbm = tuned.gbm, 
                                            RF.abundance.randomForest.randomForest = tuned.rf,
                                            #MARS.nonbinary.earth.earth=tuned.mars,
                                            #CTA.nonbinary.rpart.rpart=tuned.cta,
                                            XGBOOST.abundance.xgboost.xgboost = tuned.xgboost ),
                            user.base = "bigboss",
                            bm.format = myBiomodData,
                            calib.lines = cv)

names(opt.t@options)

print("Finished: Modeling Options")

```

#Model species 
```{r}
mod_id <- c('RUR')
myBiomodModelOut <- BIOMOD_Modeling(bm.format = myBiomodData,
                                    modeling.id = mod_id,
                                    models = all_models,
                                    CV.strategy = "user.defined",
                                    CV.user.table = cv,
                                    OPT.strategy = "user-defined",
                                    OPT.user = opt.t,
                                    metric.eval = c("MAE", "Rsquared"),
                                    var.import = 100,
                                    seed.val = 42,
                                    nb.cpu = 14)


#for ordinal data this includes: Accuracy, Recall, Precision and F1 
print("Finished: Modeling (out)")

```

#Single model evaluation scores 
```{r}
eval <- get_evaluations(myBiomodModelOut)

summary(eval)

write.csv(eval, "./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_single_model_scores.csv", row.names = FALSE)

var <- get_variables_importance(myBiomodModelOut)

summary(var)

write.csv(var, "./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_single_variable_importance.csv", row.names = FALSE)

print("Finished: Saving single model metrics")
    
```

#Graph single model evaluation scores
```{r}
#model scores plot 1
plot_output <-  bm_PlotEvalMean(bm.out = myBiomodModelOut) 
summary(plot_output)

ggdat <- plot_output$tab
summary(ggdat)


names(ggdat) <- c('algo', 'MAE_mean','Rsquared_mean', 'MAE_sd', "Rsquared_sd") ###################EDIT HERE 
ggdat$algo <- as.factor(ggdat$algo)

gg1 <-ggplot(ggdat, aes(x = MAE_mean, y = Rsquared_mean, color = algo)) + 
  geom_point() + 
  geom_errorbarh(aes(xmax = MAE_mean + MAE_sd, xmin = MAE_mean - MAE_sd),height=0)+ 
  geom_linerange(aes(ymax= Rsquared_mean + Rsquared_sd, ymin=Rsquared_mean - Rsquared_sd)) +
  theme_bw()+ theme(legend.title = element_blank(), 
                    legend.key = element_rect(fill = "white"), 
                    legend.position = "bottom",
                    axis.text.x = element_text(angle = 45, hjust = 1))+ 
  theme_classic() + xlab("MAE")+ ylab("Rsquared") +
  ggtitle(paste0("Comparison of model calibration metrics for ", sp," models" ))+
  scale_color_npg() 


gg1

# Save the plot
output_path <- file.path("./Plots/Eval_Plots/Count/RURcount_mean_model_scores.png")
ggsave(output_path, plot = gg1, width = 8, height = 6, dpi = 300)


#model scores plot 2
plot_output <- bm_PlotEvalBoxplot(bm.out = myBiomodModelOut, group.by = c('algo', 'algo'))
ggdat <- plot_output$tab
MAE_data <- ggdat %>% filter(metric.eval == "MAE")
rsq_data <- ggdat %>% filter(metric.eval == "Rsquared")

# Create MAE calibration plot
gg_MAE_cal <- ggplot(MAE_data, aes(y = calibration, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("MAE (Calibration)") +
  xlab("Algorithm") +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("MAE Calibration Metrics for ", sp, " Models"))

# Create R-squared calibration plot with y-axis from 0 to 1
gg_rsq_cal <- ggplot(rsq_data, aes(y = calibration, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("R-squared (Calibration)") +
  xlab("Algorithm") +
  ylim(0, 1) +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("R-squared Calibration Metrics for ", sp, " Models"))

# Create MAE validation plot
gg_MAE_val <- ggplot(MAE_data, aes(y = validation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("MAE (Validation)") +
  xlab("Algorithm") +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("MAE Validation Metrics for ", sp, " Models"))

# Create R-squared validation plot with y-axis from 0 to 1
gg_rsq_val <- ggplot(rsq_data, aes(y = validation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  ylab("R-squared (Validation)") +
  xlab("Algorithm") +
  ylim(0, 1) +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("R-squared Validation Metrics for ", sp, " Models"))


# Create MAE validation plot
gg_MAE_ev <- ggplot(MAE_data, aes(y = evaluation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("MAE (Evaluation)") +
  xlab("Algorithm") +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("MAE Evaluation Metrics for ", sp, " Models"))

# Create R-squared validation plot with y-axis from 0 to 1
gg_rsq_ev <- ggplot(rsq_data, aes(y = evaluation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  ylab("R-squared (Evaluation)") +
  xlab("Algorithm") +
  ylim(0, 1) +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("R-squared Evaluation Metrics for ", sp, " Models"))

# Combine all four plots into one figure
gg2 <- grid.arrange(gg_MAE_cal, gg_rsq_cal, gg_MAE_val, gg_rsq_val, gg_MAE_ev, gg_rsq_ev, ncol = 2, nrow = 3)

gg2

# Save the combined plot
output_path <- file.path("./Plots/Eval_Plots/Count/RURcount_combined_boxplot_model_scores.png")
ggsave(output_path, plot = gg2, width = 12, height = 10, dpi = 300)

#variable importance 

plot_output <- bm_PlotVarImpBoxplot(bm.out = myBiomodModelOut, group.by = c('expl.var', 'algo', 'algo')) 
ggdat <- plot_output$tab
ggdat <- ggdat %>%
  dplyr::mutate(expl.var.factor=as.factor(expl.var))%>%
  group_by(expl.var.factor,algo)%>%
  dplyr::summarize(mean = mean(var.imp),SD=sd(x=var.imp,na.rm=TRUE),N=sum(var.imp),SE= SD/sqrt(N))

gg3 <-ggplot(ggdat, aes(y = mean, x = reorder(expl.var.factor, -mean), color = algo, fill= algo)) + 
  geom_bar(position = "dodge",stat="identity") +
  geom_errorbar(aes(ymax=mean + SE, ymin = mean - SE),position="dodge",stat="identity",color="grey10")+
  ylim(-0.05,1)+
  theme_classic() + 
  theme(legend.title = element_blank(), 
        legend.key = element_rect(fill = "white"), 
        axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "bottom") + 
  ylab("Variable Importance")+ 
  xlab("Calibration")+ scale_color_npg()+ scale_fill_npg()+
  ggtitle(paste0("Variable importance for ", sp," models" ))
gg3 <- grid.arrange(gg3)

gg3

# Save the plot
output_path <- file.path("./Plots/Eval_Plots/Count/RURcount_variable_importance.png")
ggsave(output_path, plot = gg3, width = 8, height = 6, dpi = 300)

#response curves 
# Get response data
plot_output <- bm_PlotResponseCurves(
  bm.out = myBiomodModelOut,
  models.chosen = "all",
  fixed.var = 'mean',
  do.bivariate = FALSE,
  do.plot = FALSE,
  do.progress = FALSE
)

ggdat <- plot_output$tab %>%
  dplyr::mutate(expl.var.factor = as.factor(expl.name))

# Assign algorithm labels based on prediction name
ggdat$algo <- ifelse(grepl("_GLM", ggdat$pred.name, ignore.case = TRUE), "GLM",
                     ifelse(grepl("_GBM", ggdat$pred.name, ignore.case = TRUE), "BRT",
                            ifelse(grepl("_RF", ggdat$pred.name, ignore.case = TRUE), "RF",
                                   ifelse(grepl("_CTA", ggdat$pred.name, ignore.case = TRUE), "CTA",
                                          ifelse(grepl("_MARS", ggdat$pred.name, ignore.case = TRUE), "MARS",
                                                 ifelse(grepl("_XGBOOST", ggdat$pred.name, ignore.case = TRUE), "XGBOOST",
                                                        ifelse(grepl("_GAM", ggdat$pred.name, ignore.case = TRUE), "GAM", "NA")))))))


ggdat$algo <- as.factor(ggdat$algo)
ggdat$id <- as.factor(ggdat$id)

# Set custom colors
algo_colors <- c(
  "BRT" = "#FDE725FF",
  "GAM" = "#EAD7D1",
  "GLM" = "#238A8DFF",
  "MARS"= "red",
  "RF"= "blue",
  "CTA"="purple",
  "XGBOOST"="green",
  "Ensemble" = "#301934"
)

# Plot
gg4 <- ggplot(ggdat, aes(x = expl.val, y = pred.val, color = algo)) +
  geom_smooth(aes(group = algo), fill = "grey30", se = TRUE, method = "loess", alpha = 0.01) +  # By algorithm
  geom_smooth(color = algo_colors["Ensemble"], fill = "grey30", se = TRUE, method = "loess") + # Ensemble line
  facet_wrap(~ expl.var.factor, scales = "free_x") +
  xlab("Predictor Variable") +
  ylab("Prob. of Density") +
  ggtitle(paste0("Response curves for ", sp, " models")) +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_color_manual(values = algo_colors)

# Arrange and save
gg4 <- grid.arrange(gg4)

output_path <- file.path("./Plots/ES_Plots/RURcount_response_curves.png")
ggsave(output_path, plot = gg4, width = 8, height = 6, dpi = 300)

print("Finished: Graphing single model scores")
```

#Build ensemble models
```{r}
EnsembleModelsOut <- BIOMOD_EnsembleModeling(bm.mod= myBiomodModelOut,
                                             models.chosen = 'all', #define models kept, can exclude models here
                                             em.by='all', #chr, way kept models are combined to build ensemble models, see below 
                                             em.algo = c('EMwmean'), #weighted mean by the evaluation metric 
                                             metric.select = c('Rsquared'), #metric select and threshold determines which single models to keep 
                                             metric.select.thresh = c(0.3),
                                             metric.select.dataset = c('evaluation'),
                                             metric.eval= c('Rsquared'),
                                             EMci.alpha = 0.05, #num, significance level to estimate confidence interval
                                             EMwmean.decay ="proportional", # value defining relative importance of weights, discrimate good v bad models
                                             var.import=100) #vect, number of permutations to determine variable importance


print("Finished: Ensemble models")
    
```

#Ensemble model evaluation metrics 
```{r}
eval <- get_evaluations(EnsembleModelsOut)

summary(eval)

write.csv(eval, "./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_ensemble_model_scores.csv", row.names = FALSE)

var <- get_variables_importance(EnsembleModelsOut)

summary(var)

write.csv(var, "./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_ensemble_variable_importance.csv", row.names = FALSE)


print("Finished: Plotting saving ensemble model scores")
```

#Plot ensemble evaluation 
```{r}
eval_scores <- eval
eval_predictor <- var

#Plot relative predictor importance 
predictor_summary <- eval_predictor %>%
  group_by(expl.var) %>%
  summarise(
    mean_var_imp = mean(var.imp),
    sd_var_imp = sd(var.imp),
    .groups = "drop"
  ) %>%
  mutate(
    var.imp.percent = mean_var_imp * 100,
    sd.percent = sd_var_imp * 100,
    min = var.imp.percent - sd.percent,
    max = var.imp.percent + sd.percent
  )

plot <- ggplot() +
  geom_col(
    data = predictor_summary,
    mapping = aes(
      x = var.imp.percent,
      y = reorder(expl.var, var.imp.percent)
    ),
    position = position_dodge2(padding = 0),
    width = 0.7,
    preserve = "total",
    fill = "#e6360F"
  ) +
  geom_errorbar(
    data = predictor_summary,
    mapping = aes(
      y = expl.var,
      xmin = min,
      xmax = max
    ),
    position = position_dodge2(preserve = "total"),
    width = 0.5,
    size = 0.7
  ) +
  theme_cowplot() +
  guides(x.sec = "axis", y.sec = "axis") +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 9)
  ) +
  xlab("Predictor Importance (%)") +
  ylab("Predictor Variable")

plot


output_path <- file.path("./Plots/PI_Plots/RURcount_ensemble_predictor_importance.png")
ggsave(output_path, plot = plot, width = 8, height = 6, dpi = 300)

#ES strip 
#response curves 
# Get response data
plot_output <- bm_PlotResponseCurves(
  bm.out = EnsembleModelsOut,
  models.chosen = "all",
  fixed.var = 'mean',
  do.bivariate = FALSE,
  do.plot = FALSE,
  do.progress = FALSE
)

ggdat <- plot_output$tab %>%
  dplyr::mutate(expl.var.factor = as.factor(expl.name))

write.csv(ggdat, "./Generated_Data/ES_datasets/RUR_ES_Count.csv", row.names = FALSE)

print("Finished: Plotting ensemble models")
```

#Projections 
```{r}

gc()

#only select models that end up exceeding the 0.3 R-squared threshold 
EnsembleModelsOut <- get(load("./models_biomod/count/RUR/RUR.RUR.ensemble.models.out"))
myBiomodModelOut <- get(load("./models_biomod/count/RUR/RUR.RUR.models.out"))

models <- get_built_models(myBiomodModelOut, algo= all_models)

#this works below, from Jenn's code 
sp_proj <- BIOMOD_Projection(
         bm.mod =  myBiomodModelOut,
         new.env = proj.var ,
         proj.name = 'RUR_current_count',
         models.chosen =models,
         digits = 2, #number of digits for predicted values
         build.clamping.mask = TRUE, 
         output_format = ".tif",
         do.stack = TRUE,
         nb.cpu = 10
       )

gc()

png(filename = "./Plots/Maps/Abundance_Maps/RUR_count_projection.png", width = 1200, height = 1000)
plot(sp_proj)
dev.off()

gc()

#run projections on ensemble models, using current climatology
ens_proj <- BIOMOD_EnsembleForecasting(
         bm.em = EnsembleModelsOut,
         #new.env = proj.var,
         bm.proj = sp_proj,
         proj.name = 'RUR_current_count_ensembles',
         models.chosen = 'all',
         build.clamping.mask = TRUE, 
         output.format=".tif",
         do.stack=TRUE,
         #on_0_to_1000 = FALSE 
         )

png(filename = "./Plots/Maps/Abundance_Maps/RUR_count_projection_ensemble.png", width = 1200, height = 1000)
plot(ens_proj)
dev.off()


```

#Quantifying uncertainty among different algorithm projections 
```{r}
myBiomodModelOut <- get(load('./models_biomod/count/RUR/RUR.RUR.models.out'))

models.kept <- BIOMOD_LoadModels(myBiomodModelOut)

# Manually exclude models associated with RUN1, RUN4, RUN6, and allRun
models.chosen <- models.kept[!grepl("RUN8|RUN5|RUN6|allRun", models.kept)]

EnsembleModelsOut <- BIOMOD_EnsembleModeling(
  bm.mod = myBiomodModelOut,
  models.chosen = models.chosen,
  em.by = 'PA+run',
  em.algo = c('EMwmean'),
  metric.select = c('Rsquared'),
  metric.select.thresh = c(0.3),
  metric.eval = c('Rsquared'),
  metric.select.dataset = c('evaluation'),
  EMci.alpha = 0.05,
  EMwmean.decay = "proportional",
  var.import = 100
)


#models <- get_built_models(myBiomodModelOut, algo= c("GBM", "XGBOOST",  "GLM", "RF"))
#this works below, from Jenn's code 
ens_proj <- BIOMOD_EnsembleForecasting(
         bm.em = EnsembleModelsOut,
         new.env = proj.var,
         #bm.proj = sp_proj,
         proj.name = 'RUR_current_count_ensembles_algos',
         models.chosen = 'all',
         build.clamping.mask = TRUE, 
         output.format=".tif",
         do.stack=TRUE,
         #on_0_to_1000 = FALSE 
         )

Prediction <- terra::rast("./models_biomod/count/RUR/proj_RUR_current_count_ensembles_algos/proj_RUR_current_count_ensembles_algos_RUR_ensemble.tif")


Model_Uncertainty <- terra::stdev(Prediction)

terra::writeRaster(Model_Uncertainty, "./processed_data/Uncertainty/RUR_Method_Uncertainty_Count.tif", overwrite = TRUE)

RC_Map <- function(RC_Layer, Legend_Label, layername, BC) {
  # Compute min and max values for the chosen layer
  vmin <- round(min(RC_Layer[[layername]], na.rm = TRUE), 2)
  vmax <- round(max(RC_Layer[[layername]], na.rm = TRUE), 2)

  # Calculate break points for the legend
  breaks <- round(seq(vmin, vmax, length.out = 5), 2)
  labels <- as.character(breaks)

  # Build the ggplot
  RC_Plot <- ggplot(BC) +
    geom_sf(data = BC, fill = "#d9d9d9", color = NA) +
    geom_raster(data = RC_Layer, aes(x = x, y = y, fill = .data[[layername]])) +
    coord_sf(crs = st_crs(3005)) +
    xlim(c(515000, 1255000)) +
    ylim(c(340000, 1100000)) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true",
                           pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                           style = north_arrow_fancy_orienteering) +
    theme_cowplot() +
    scale_fill_continuous(
      type = "viridis",
      na.value = NA,
      name = Legend_Label,
      limits = c(vmin - 0.01, vmax + 0.01),
      breaks = breaks,
      labels = labels
    ) +
    xlab(element_blank()) + ylab(element_blank()) +
    theme(
      legend.position = c(1, 1),
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 10),
      legend.text.align = 0,
      legend.title.align = 0,
      legend.justification = c("right", "top"),
      legend.box.just = "right",
      panel.background = element_rect(fill = "white")
    )

  return(RC_Plot)
}

# Convert SpatRaster to data frame
RC_Layer_df <- as.data.frame(Model_Uncertainty, xy = TRUE)

# Convert SpatVector to sf
Land <- terra::vect("./raw_data/BC_Buffer/Land.shp")
Land <- sf::st_as_sf(Land)

# Create the plot
RUR_Method_Uncertainty_Plot <- RC_Map(RC_Layer_df, "Model\nUncertainty", "std", Land)
RUR_Method_Uncertainty_Plot

ggsave2(filename = "./Plots/Maps/Uncertainty_Plots/RUR_Model_Uncertainty_Map_Count.jpeg", plot = RUR_Method_Uncertainty_Plot, scale = 1, height = 140, width = 115, units = "mm")

```

#Evaluating NULL models
```{r}
gc()

#Create a dataframe of the environment to pull null model training dataset from 
Training_Data <- RUR_training_df  #needs to be a dataframe with xy, resp, env with only TRAINING data 
Testing_Data <- RUR_evaluation_df #needs to be a dataframe with xy, resp, env with only TESTING data 
Survey_Enviro_df <- env_df #includes all environmental data 
Predictor_num <- ncol(env_df)

#Create null models using the same parameters and randomly drawn presence points for the enviro_buffer dataframe (same background points)

#Set up a dataframe to hold all of the evaluation metrics for the null models
Null_metrics <- data.frame(
  full.name = character(),
  merged.by.PA = character(),
  merged.by.run = character(),
  merged.by.algo = character(),
  algo = character(),
  metric.eval = character(),
  cutoff = logical(),
  sensitivity= logical(),
  calibration = numeric(),
  validation = numeric(),
  evaluation = numeric()
)


#Run 1000 null models
for (j in 1:10){
  
  #Perform a random pull from the environmental dataframe for each presence point in the training dataset
  for (i in 1:nrow(Training_Data)) {
    set.seed(313*j + round(46*i))
    randnum <- sample(seq_len(nrow(Survey_Enviro_df)), size = 1)
    Training_Data[i, 4:(3+ Predictor_num)] <- Survey_Enviro_df[randnum,]
    
    train_resp_df <- Training_Data$mean
    train_xy <- Training_Data %>% dplyr::select(x, y)
    train_env <- Training_Data %>% dplyr::select(TAlk_10m_mean, salt_10m_min, temp_s_max, temp_s_min, temp_10m_mean, temp_10m_min, do_10m_mean, do_10m_min, Current, REI, Rock, Slope, Depth)
    
  }
  
  sp <- c('RUR')
  
  #define models to run 
  all_models <- c("GLM","GBM","RF", "XGBOOST" ) 
  
  #fit my *real* biomod2 models to null TRAINING data, made by for loop above
  myBiomodData <- BIOMOD_FormatingData(resp.var = train_resp_df, #as SpatVector without associated data, for a single species
                                       expl.var = train_env, #SpatRaster containing explanatory variables in layers
                                       resp.name = sp, #character corresponding to species name
                                       resp.xy = train_xy, #if resp.var is vector, 2 column matrix associated xy coordinates
                                       dir.name = c('./models_biomod/null_count'), 
                                       data.type = c("abundance"), #can also be 'ordinal'
                                       eval.expl.var = eval_env, #SpatRaster without associated data, independent data to evaluate
                                       eval.resp.var = eval_resp_df, #SpatVector without associated data, independent data to evaluate 
                                       eval.resp.xy = eval_xy, #if resp.vect is vector, 2 column matrix of xy coordinates
                                       na.rm = TRUE, 
                                       filter.raster = TRUE, #filter points in same cell
                                       seed.val = 1234)
  
  
  cv <- bm_CrossValidation(bm.format = myBiomodData,
                           strategy = "kfold", #options are random, k-fold, strat, block, env, or user-defined
                           nb.rep = 1, #integer, number of repetitions of CV points to be drawn 
                           k = 10, #number of partitions                                         
                           do.full.models = TRUE #calibrated and validated over the whole dataset (and pseudo-absence datasets) or not
  ) 
  
  
  
  myBiomodModelOut <- BIOMOD_Modeling(bm.format = myBiomodData,
                                      models = all_models,
                                      CV.strategy = "user.defined",
                                      CV.user.table = cv,
                                      OPT.strategy = "user-defined",
                                      OPT.user = opt.t,
                                      metric.eval = c("MAE", "Rsquared"),
                                      var.import = 100,
                                      seed.val = 42,
                                      nb.cpu = 10)
  
  EnsembleModelsOut <- BIOMOD_EnsembleModeling(bm.mod= myBiomodModelOut,
                                               models.chosen = 'all', #define models kept, can exclude models here
                                               em.by='all', #chr, way kept models are combined to build ensemble models, see below 
                                               em.algo = c('EMwmean'), #weighted mean by the evaluation metric 
                                               metric.select = c('Rsquared'), #metric select and threshold determines which single models to keep 
                                               #metric.select.thresh = c(0.3),
                                               metric.eval= c('Rsquared'),
                                               metric.select.dataset = c('evaluation'),
                                               EMci.alpha = 0.05, #num, significance level to estimate confidence interval
                                               EMwmean.decay ="proportional", # value defining relative importance of weights, discrimate good v bad models
                                               var.import=100) #vect, number of permutations to determine variable importance
  
  #get model evaluation metrics 
  eval_ensemble <- get_evaluations(EnsembleModelsOut)
  eval_ensemble <- as.data.frame(eval_ensemble)
  
  #add eval_ensemble to empty df of null metrics above
  Null_metrics <- rbind(Null_metrics, eval_ensemble)
  
}


#save csv of Null Metrics 
write.csv(Null_metrics, "./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_null_model_eval.csv", row.names = FALSE)

```

#Graphing comparison of real and null models 
```{r}
#read in files 
models <- read.csv("./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_ensemble_model_scores.csv")

null <- read.csv("./Generated_Data/Evaluation_datasets/Count/RUR/RURcount_null_model_eval.csv")

#Format 
#Make a difference between testing and evaluation R²
models$difference <- abs(models$evaluation - models$calibration)
null$difference <- abs(null$evaluation - null$calibration)

#rename columns
models <- models %>%
  rename("R² Training" = calibration,
         "R² Testing" = evaluation, 
         "R² Difference" = difference)


null <- null %>%
  rename("R² Training" = calibration,
         "R² Testing" = evaluation, 
         "R² Difference" = difference)



#PLot
#Create long-form dataframes for the model and null model metrics for plotting
  Model_metrics_long <- models %>% pivot_longer(cols = c("R² Training", "R² Testing", "R² Difference"), names_to = "Metric", 
                        values_to = "Value") %>% select('Metric', 'Value')
  
Model_metrics_long$Metric <- factor(Model_metrics_long$Metric, levels = c("R² Training", "R² Testing", "R² Difference"))
  
Null_metrics_long <- null %>%
  pivot_longer(cols = c("R² Training", "R² Testing", "R² Difference"),
               names_to = "Metric", values_to = "Value") %>% select(Metric, Value)


Null_metrics_long$Metric <- factor(Null_metrics_long$Metric, levels = c("R² Training", "R² Testing", "R² Difference"))
  
  #Plot the model and null model evaluation metrics
  RUR_Eval_Plot <- ggplot() + geom_boxplot(data = Null_metrics_long, aes(x = Metric, y = Value, fill = Metric)) + 
    scale_fill_manual(values = c("#613f74", "#91B3B6", "#F4E76E", "#EDAF97"), labels = c("R² Training", "R² Testing", "R² Difference")) + 
    stat_summary(data = Model_metrics_long, aes(x = Metric, y = Value), fun.y = mean, geom = "point", colour = "#e6360F", size = 3) + 
    xlab("Evaluation Metric") + ylab("Value") + theme_classic() + 
    guides(x.sec = "axis", y.sec = "axis")+ theme(axis.line.y.left = element_line(size = 0.75, color = "black"), axis.line.y.right = element_line(size = 0.75, color = "black"), axis.line.x.top = element_line(size = 0.75, color = "black"), axis.line.x.bottom = element_line(size = 0.75, color = "black"),axis.ticks.x.top = element_blank(), axis.ticks.x.bottom = element_blank(), axis.ticks.y.right = element_blank(), axis.text.x.top = element_blank(), axis.text.y.right = element_blank()) + 
    theme(legend.position = "none", legend.justification = "center") + 
    theme(axis.text = element_text(size=8), axis.title = element_text(size = 12))
  
  
  RUR_Eval_Plot
  

  ggsave2(filename = "./Plots/Eval_Plots/Count/RUR_Eval_Plot.jpeg", plot = RUR_Eval_Plot, scale = 1, height = 75, width = 100, units = "mm")
```

#Plot final ensemble response curves 
```{r}
#count response curve data 
count_RUR <- read.csv("./Generated_Data/ES_datasets/RUR_ES_Count.csv")


sp <- c("RUR")
ggdat <- count_RUR
ggdat$id <- as.factor(ggdat$id)

#Rename 
ggdat <- ggdat %>%
  mutate(
    expl.var.factor = recode(expl.var.factor,
                             "Current" = "Current (m/s)",
                             "Slope" = "Slope (deg)",
                             "Rock" = "Percent Rock",
                             "TAlk_10m_mean" = "Talkbmax (mEq/m³)",
                             "salt_10m_min" = "Sbmin (PSU)",
                             "temp_s_max" = "Tsmax (°C)",
                             "temp_s_min" = "Tsmin (°C)",
                             "temp_10m_mean" = "Tbmean (°C)",
                             "temp_10m_min" = "Tbmin (°C)",
                             "do_10m_mean" = "Oxybmean (ppm)",
                             "do_10m_min" = "Oxybmin (ppm)",
                             "Depth" = "Depth (m)"
    )
  )


ggdat <- ggdat %>%
  mutate(
    expl.name = recode(expl.name,
                       "Current" = "Current (m/s)",
                       "Slope" = "Slope (deg)",
                       "Rock" = "Percent Rock",
                       "TAlk_10m_mean" = "Talkbmax (mEq/m³)",
                       "salt_10m_min" = "Sbmin (PSu)",
                       "temp_s_max" = "Tsmax (°C)",
                       "temp_s_min" = "Tsmin (°C)",
                       "temp_10m_mean" = "Tbmean (°C)",
                       "temp_10m_min" = "Tbmin (°C)",
                       "do_10m_mean" = "Oxybmean (ppm)",
                       "do_10m_min" = "Oxybmin (ppm)",
                       "Depth" = "Depth (m)"
    )
  )



#Separate the two groups of predictors
subset_0_10 <- ggdat %>%
  filter(!expl.var.factor %in% c("Oxybmean (ppm)", "Oxybmin (ppm)"))

subset_0_100 <- ggdat %>%
  filter(expl.var.factor %in% c("Oxybmean (ppm)", "Oxybmin (ppm)"))


#Plot for most variables (y = 0–10)
p1 <- ggplot(subset_0_10, aes(x = expl.val, y = pred.val)) +
  geom_smooth(aes(group = expl.var.factor), fill = "grey30", color = "#e6360F", se = TRUE, method = "loess") +
  facet_wrap(~ expl.var.factor, scales = "free_x", strip.position = "bottom") +  # strip labels below
  coord_cartesian(ylim = c(0, 10)) +
  xlab("") +
  ylab("Prob. of Density") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_blank(),   # remove box background
    strip.placement = "outside"            # place strip outside plot panel
  )

# Plot for do_10m predictors (y = -200 to 200)
p2 <- ggplot(subset_0_100, aes(x = expl.val, y = pred.val)) +
  geom_smooth(aes(group = expl.var.factor), fill = "grey30", color = "#e6360F", se = TRUE, method = "loess") +
  facet_wrap(~ expl.var.factor, scales = "free_x", strip.position = "bottom") +  # strip labels below
  coord_cartesian(ylim = c(-200, 200)) +
  xlab("Predictor Variable") +
  ylab("Prob. of Density") +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_blank(),  # remove box background
    strip.placement = "outside"           # place strip outside plot panel
  )
  

# Combine the two
p3 <- p1 / p2  # patchwork stacks them vertically

p3
#Save 
ggsave("./Plots/ES_Plots/RUR_ES_Count.png", plot = p3, width = 6, height = 10, dpi = 300)
```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
