---
title: "C_GUR_Ordinal_Code"
author: "Tessa Rehill"
date: "2025-05-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install packages 
```{r}
library(dplyr)
library(tidyr)
library(tidyverse)
library(terra)
library(sf)
library(ggspatial)
library(cowplot)
library(biomod2)
library(ggtext)
library(usdm)
library(gridExtra)
library(ggsci)
library(blockCV)
library(parallel)
library(doParallel)
```

#Call necessary functions 

```{r}
source("Count_Mapping_Functions.R")
```

#Creating RD datasets (for GUR)
```{r}
#Import the mask which will be important for masking any unmasked data at this point
BC_Mask <- terra::vect("./raw_data/BC_Buffer/BC_Buffer.shp") %>% terra::project("epsg:3005")

#Import the shapefile we will be using to represent land/the coastline and project to the correct crs/mask to NC
Coastline <- vect("./raw_data/BC_Buffer/GSHHS_f_L1.shp") %>% terra::project("epsg:3005")
Land <- Coastline %>% terra::mask(BC_Mask)

Bathymetry <- terra::rast("./processed_data/ENV_layers/DEM_100.tif")

#Load in the prepped RD data 
Urchin_Transects_Filtered <- sf::st_read("./processed_data/Urchin_RDensity/Urchin_RDensity_Transects_Filtered.shp")
```

#Create green urchin shapefiles 
```{r}

selected_species <- "GUR"

filtered_data <- Urchin_Transects_Filtered %>% filter(species == selected_species)

GUR_Year_list <- unique((filtered_data %>% dplyr::filter(species == selected_species))$year)

#For each year in the GUR year list generate RD points, a RD raster, and a map of where RDs were observed
for (i in 1:length(GUR_Year_list)){
  Shapefile_Output <- Shapefile_Creator(filtered_data %>% dplyr::filter(year == GUR_Year_list[i]), Bathymetry, selected_species)
  terra::writeVector(Shapefile_Output, paste0("./processed_data/Urchin_RDensity/GUR_RDensity_Points_", GUR_Year_list[i],".shp"), overwrite = TRUE)
}


```

## Create the final RD dataset for GUR
```{r}
#Generate a list of years for which occurrence data is available 
GUR_Year_list <- unique((Urchin_Transects_Filtered %>% dplyr::filter(species == "GUR"))$year)

#For each year in the GUR year list generate Count points, a relative density raster, and a map of where relative density were observed
for (i in 1:length(GUR_Year_list)){
  GUR_Output <- Point_Creator(Urchin_Transects_Filtered %>% dplyr::filter(year == GUR_Year_list[i]), "GUR", Bathymetry)
  GUR_Count_Points <- GUR_Output[[2]]
  GUR_Raster <- GUR_Output[[1]]
  terra::writeVector(GUR_Count_Points, paste0("./processed_data/Urchin_RDensity/GUR_RD_Points_", GUR_Year_list[i],".shp"), overwrite = TRUE)
  terra::writeRaster(GUR_Raster, paste0("./processed_data/Urchin_RDensity/GUR_RD_Raster_", GUR_Year_list[i], ".tif"), overwrite = TRUE)
  GUR_Pres_Map <- Pres_Map(Land, GUR_Count_Points %>% terra::project(crs(Land)))
  ggsave(paste0("./Plots/Maps/Species_RDensity_Maps/GUR_RD_Map_", GUR_Year_list[i], ".jpeg"), GUR_Pres_Map, units = "mm", width = 120, height = 140)
}

#Do the same for the entire time period 
GUR_Output <- Point_Creator(Urchin_Transects_Filtered, "GUR", Bathymetry)
GUR_Count_Points <- GUR_Output[[2]]
GUR_Raster <- GUR_Output[[1]]
terra::writeVector(GUR_Count_Points, "./processed_data/Urchin_RDensity/GUR_RD_Points.shp", overwrite = TRUE)
terra::writeRaster(GUR_Raster, "./processed_data/Urchin_RDensity/GUR_RD_Raster.tif", overwrite = TRUE)

GUR_Count_Points <- terra::vect("./processed_data/Urchin_RDensity/GUR_RD_Points.shp")
GUR_Pres_Map <- Pres_Map(Land, GUR_Count_Points %>% terra::project(crs(Land)))

ggsave("./Plots/Maps/Species_RDensity_Maps/GUR_Count_Map.jpeg", GUR_Pres_Map, units = "mm", width = 120, height = 140)
```

#BIOMOD PIPELINE HERE
```{r}

Land <- terra::vect("./raw_data/BC_Buffer/Land.shp")

Urchin_Transects_Filtered <- sf::st_read("./processed_data/Urchin_RDensity/Urchin_RDensity_Transects_Filtered.shp")
colnames(Urchin_Transects_Filtered)[4] <- "Year"
colnames(Urchin_Transects_Filtered)[11] <- "Species"


GUR_Transects <- terra::subset(Urchin_Transects_Filtered, Species == "GUR")
GUR_Year_list <- unique(GUR_Transects$Year)

#Set up dataframes to hold 
resp_df <- data.frame(mean = numeric())
env_df <- data.frame(
  ID = numeric(),
  TAlk_10m_mean = numeric(),
  salt_10m_mean = numeric(),
  salt_10m_min = numeric(),
  temp_s_mean = numeric(),
  temp_s_max = numeric(),
  temp_s_min = numeric(),
  temp_10m_mean = numeric(),
  temp_10m_max = numeric(),
  temp_10m_min = numeric(),
  do_10m_mean = numeric(),
  do_10m_min = numeric(),
  Current = numeric(),
  REI = numeric(),
  Rock = numeric(),
  Slope = numeric(),
  Depth = numeric()
)


#Create an empty data frame to store x,y
xy <- data.frame(
  x = numeric(),
  y = numeric()
)


for (i in 1:length(GUR_Year_list)){
  #Load the environmental layers for year i
  env.var <- terra::rast(paste0("./processed_data/ENV_layers/BCCM_Rasters/Predictor_hindcast_climatology_", GUR_Year_list[i], ".tif"))
  
  #Load shapefile for year i 
  resp.var <- terra::vect(paste0("./processed_data/Urchin_RDensity/GUR_RDensity_Points_", GUR_Year_list[i], ".shp"))
  
  #Convert ordinal category to a factor (for summarization in nect step)
  resp.var$density <- factor(resp.var$density, levels = c("Abundant", "Many", "Few", "None"))
  
  #Remove rows where 'density' is 'Null' or NA
  resp.var <- resp.var[!is.na(resp.var$density) & resp.var$density != "Null", ]
  
  # Convert the 'density' factor to numeric (you can assign values based on levels order)
  resp.var$density <- as.numeric(factor(resp.var$density, levels=c("None", "Few", "Many", "Abundant")))
  
  #Rasterize shapefile, using env raster, take average density
  resp_raster <- terra::rasterize(resp.var, env.var, field = "density", fun = "mean") #takes most common density category per cell, creates raster
  
  #Resample shapefile, to env raster
  resp_raster <- terra::resample(resp_raster, env.var) #align resolutions and extents to environmental raster
  
  #Convert raster to point (using as.points) 
  points <- terra::as.points(resp_raster) #now each point has a location and averaged density value 
  
  
  points_df <- as.data.frame(points)
  
  # Convert density to an ordered factor with custom labels
  points_df$mean <- factor(points_df$mean,
                                    levels = c(1,2,3,4),
                                    labels = c("None","Few","Many", "Abundant"),
                                    ordered = TRUE)
  
  #Extract coordinates of shapefile (using geom)
  coordinates <- terra::geom(points, df=TRUE) %>% dplyr::select(x, y)
  
  #Extract environmental variables for these coordinates 
  env_values <- terra::extract(env.var, points)
  
  #Append points_df to resp_df
  resp_df <- rbind(resp_df, points_df)
  
  #Append env_values to env_df
  env_df <- rbind(env_df, env_values)
  
  #Append coordinate to xy dataframe using rbind
  xy <- rbind(xy, coordinates)
}

summary(resp_df)
summary(env_df)

#create full training dataframe 
print("Finished: Create dataframes BY year")

env_df <- env_df[, !(colnames(env_df) %in% c("ID", "temp_s_mean", "salt_10m_mean", "temp_10m_max"))]

#Read in projection layers, and remove the same variables

proj.var <- rast("./processed_data/ENV_layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif") 

proj.var <- proj.var[[!names(proj.var) %in% c("temp_s_mean", "salt_10m_mean", "temp_10m_max")]] #remove same variables

names(proj.var)

print("Finished: Running collinearity") 

#Create evaluation dataset using blockCV

#Create final training dataset
GUR_training_data <- cbind(xy, resp_df, env_df) %>% na.omit() #includes all data, 2105  observations (rows)  


#Set the size of spatial blocks based on previously completed autocorrelation analysis
Block_Size <-90000  #CODEEDIT to adjust block size for the species if needed 
#Set the k-value for k-fold cross-validation: number of spatial folds
k <- 10

#Load in hindcast climatology #CODEEDIT select for only the variables I want included in my models
Hindcast_Climatology <- terra::rast("./processed_data/ENV_layers/BCCM_Rasters/Predictor_Hindcast_Climatologies_2012-2023.tif")

Hindcast_Climatology <- terra::subset(Hindcast_Climatology, !(names(Hindcast_Climatology) %in% c("temp_s_mean", "salt_10m_mean", "temp_10m_max")))


#Turn the final training dataset into an sf object
GUR_training_data_sf <- sf::st_as_sf(GUR_training_data, coords = c("x", "y"), crs = 3005)

#Set seed
set.seed(157)
#Spatially block based on the optimal size suggested above
GUR_training_blocks <- blockCV::cv_spatial(GUR_training_data_sf, column = "mean", hexagon = FALSE, k = k+1, size = Block_Size, selection = "random", iteration = 200)

GUR_training_data$fold <- GUR_training_blocks$folds_ids
GUR_training_data_sf$fold <- GUR_training_blocks$folds_ids

#Check to see how many  points ended up in each block
Point_summary <- terra::as.data.frame(GUR_training_data) %>%
  dplyr::group_by(fold) %>%
  dplyr::summarise(Point_num = dplyr::n())

write.csv(Point_summary, "./Generated_Data/Interim_datasets/GURord_point_summary.csv")

#Check to see how many of each ordinal category is in each fold 
Point_summary <-GUR_training_data %>%
  group_by(fold, mean) %>%
  summarise(count = n()) %>%
  arrange(fold, desc(mean))

# Summarize the counts
plot_data <- GUR_training_data %>%
  group_by(fold, mean) %>%
  summarise(count = n(), .groups = "drop")

# Plot
plot <- ggplot(plot_data, aes(x = factor(fold), y = count, fill = mean)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Fold", y = "Count", fill = "Density Category",
       title = "Density Category Counts per Fold") +
  scale_fill_manual(values = c("None" = "#999999", "Few" = "#FFD700", "Many" = "#FFA500", "Abundant" = "#E31A1C")) +
  theme_minimal()

plot

write.csv(Point_summary, "./Generated_Data/Interim_datasets/GURord_point_summary.csv")
ggsave("./Plots/GUR_point_summary.png", width = 8, height = 6, dpi = 300)


#Adjust folds if needed by reshuffling presence points from folds with a number of presence points over some cutoff based on the total number of points in the dataset
#GUR_training_data_sf <- CV_redistribute_ordinal(GUR_training_data_sf, k+1, 80) #adjust this cutoff to be reasonable based on the number of presence points and folds you are working with (the idea is to better balance presences within folds, but not to make things completely equal)

GUR_training_data_vec <- terra::vect(GUR_training_data_sf)
GUR_training_data_df <- terra::as.data.frame(GUR_training_data)

#Save the dataset for GUR
write.csv(GUR_training_data_df, "./Generated_Data/Blocked_datasets/GURord_blocks.csv")
terra::writeVector(GUR_training_data_vec, "./processed_data/Blocked_Spatial_Data/GURord_blocks.shp", overwrite = TRUE)


print("Finished: Evaluation Dataset Set Up")

#Create final training and evaluation dataset

#Create a final training df from the first k folds
GUR_training_df <- GUR_training_data_df %>% dplyr::filter(fold != (k+1))

train_resp_df <- GUR_training_df$mean
train_xy <- GUR_training_df %>% dplyr::select(x, y)
train_env <- GUR_training_df %>% dplyr::select(TAlk_10m_mean, salt_10m_min, temp_s_max, temp_s_min, temp_10m_mean, temp_10m_min, do_10m_mean, do_10m_min, Current, REI, Rock, Slope, Depth)

#Create a final evaluation df from the K+1 fold
GUR_evaluation_df <- GUR_training_data_df %>% dplyr::filter(fold == (k+1))

eval_resp_df <- GUR_evaluation_df$mean
eval_xy <- GUR_evaluation_df %>% dplyr::select(x, y)
eval_env <- GUR_evaluation_df %>% dplyr::select(TAlk_10m_mean, salt_10m_min, temp_s_max, temp_s_min, temp_10m_mean, temp_10m_min, do_10m_mean, do_10m_min, Current, REI, Rock, Slope, Depth)

#Format data

#define species to model (ch)
sp <- c('GUR')

#define species to model (ch)
all_models <- c("GAM" ) #no GBM for ordinal data, no MAXENT for any abundance

#format
myBiomodData <- BIOMOD_FormatingData(resp.var = train_resp_df, #as SpatVector without associated data, for a single species
                                     expl.var = train_env, #SpatRaster containing explanatory variables in layers
                                     resp.name = sp, #character corresponding to species name
                                     resp.xy = train_xy, #if resp.var is vector, 2 column matrix associated xy coordinates
                                     dir.name = c('./models_biomod/ordinal'), 
                                     data.type = c("ordinal"), 
                                     eval.expl.var = eval_env, #SpatRaster without associated data, independent data to evaluate
                                     eval.resp.var = eval_resp_df, #SpatVector without associated data, independent data to evaluate 
                                     eval.resp.xy = eval_xy, #if resp.vect is vector, 2 column matrix of xy coordinates
                                     na.rm = TRUE, 
                                     filter.raster = TRUE, #filter points in same cell
                                     seed.val = 1234
) 


print("Finished: BIOMOD formatting data function")

#Creating CV table ###################################################################
#folds <- sort(unique(GUR_training_df$fold))  # get sorted unique fold numbers

# Create a new dataframe with one column per fold
#fold_matrix <- sapply(folds, function(f) GUR_training_df$fold == f)

# Name the columns as Fold_1, Fold_2, etc. (optional)
#colnames(fold_matrix) <- paste0("_allData_RUN", 1:10)

# Convert to a data frame
#fold_df <- as.data.frame(fold_matrix)


#cv <- bm_CrossValidation(bm.format = myBiomodData,
#                         strategy = "user.defined", #options are random, k-fold, strat, block, env, or user-defined
#                         user.table = fold_df,
#                         do.full.models = TRUE #calibrated and validated over the whole dataset (and pseudo-absence datasets) or not
#) 

cv <- bm_CrossValidation(bm.format = myBiomodData,
                         strategy = "kfold", #options are random, k-fold, strat, block, env, or user-defined
                         nb.rep = 1, #integer, number of repetitions of CV points to be drawn 
                         k = 10, #number of partitions                                         
                         do.full.models = TRUE #calibrated and validated over the whole dataset (and pseudo-absence datasets) or not
) 

print("Finished: CV fold")

#Setting model options #########################################################

#Tuning models and setting model options


opt.d <- bm_ModelingOptions(data.type = 'ordinal', #character corresponding to data type to be used (see below)
                            models = all_models,
                            strategy = 'bigboss',
                            bm.format = myBiomodData)


opt.e <- bm_ModelingOptions(data.type = 'ordinal', #character corresponding to data type to be used (see below)
                            models = all_models,
                            strategy = 'default',
                            bm.format = myBiomodData)


opt.t <- opt.d
#
#library(doParallel)
#stopImplicitCluster() # make sure no cluster is registered

#Tuning CTA
 
 #tuned.cta <- bm_Tuning(
     #model = "CTA",
     #tuning.fun = "rpart",
     #do.formula = TRUE,
     #bm.options = opt.d@options$CTA.ordinal.rpart.rpart,
     #bm.format = myBiomodData,
     #metric.eval = "Accuracy",
     #params.train = list(CTA.maxdepth = c(5, 10, 20)) # no CTA. prefix
# )



##


#Tuning RF

 
#tuned.rf <- bm_Tuning(
  #model = 'RF',
  #tuning.fun = 'rf',
  #do.formula = TRUE, #not sure if this should be true?, when run true = throws error, also TRUE = slower
  #bm.options = opt.d@options$RF.ordinal.randomForest.randomForest, #should be an object returned bybm_ModellingOptions,
  #bm.format = myBiomodData, 
  #calib.lines = cv,
  #metric.eval = c("Accuracy"), 
#params.train = list(RF.mtry=c(2, 3, 4, 5))
#)


#print("Finished: RF tuning")

#Tuning XGBOOST 

 
#tuned.xgboost <- bm_Tuning(
  #model = 'XGBOOST',
  #tuning.fun = 'xgbTree',
  #do.formula = TRUE, #not sure if this should be true?, when run true = throws error, also TRUE = slower
  #bm.options = opt.d@options$XGBOOST.ordinal.xgboost.xgboost, #should be an object returned bybm_ModellingOptions,
  #bm.format = myBiomodData, 
  #calib.lines = cv,
  #metric.eval = c("Accuracy"), #Recall, F1, and F1
  #params.train = list(XGBOOST.nrounds = c(3,4,5), XGBOOST.max_depth =c(2,3,4), XGBOOST.eta =c(0.01,0.01, 1), XGBOOST.gamma=c(0,0.1)))

#increase gamma if you're seeing overfitting
#can also change colsample_bytree - fraction of features used per tree, lower = better for overfitting
#subsample - fraction of rows used per tree, prevents overfitting

print("Finished: XGBOOST tuning")


#tuning GAM model 
#tuned.gam <- bm_Tuning(model = 'GAM',
#                       tuning.fun = 'gam', ## see in ModelsTable
#                       do.formula = TRUE,
#                       bm.options = opt.d@options$GAM.ordinal.mgcv.bam,
#                       bm.format = myBiomodData,
#                       metric.eval=c("Accuracy"),
                       #calib.lines = cv,
#                       params.train = list(GAM.select =c('TRUE'), GAM.method = c('GCV.Cp'), GAM.knots=c(3)))

#print("Finished: GAM tuning")


#Setting tuning options
#opt.t <- bm_ModelingOptions(data.type = 'ordinal',
                            #models = all_models,
                            #strategy = "user.defined",
                            #user.val = list(
                                            #GBM.nonbinary.gbm.gbm = tuned.gbm, 
                                            #RF.ordinal.randomForest.randomForest = tuned.rf,
                                            #MARS.nonbinary.earth.earth=tuned.mars,
                                            #CTA.ordinal.rpart.rpart=tuned.cta,
                                            #XGBOOST.ordinal.xgboost.xgboost = tuned.xgboost), 
                                            #GAM.ordinal.mgcv.bam = tuned.game),
                            
                            #user.base = "bigboss",
                            #bm.format = myBiomodData,
                            #calib.lines = cv)

#names(opt.t@options)


print("Finished: Modeling Options")


print("Finished:Tuning")

#model species 

myBiomodModelOut <- BIOMOD_Modeling(bm.format = myBiomodData,
                                    models = all_models,
                                    CV.strategy = "user.defined",
                                    CV.user.table = cv,
                                    OPT.strategy = "user-defined",
                                    OPT.user = opt.d,
                                    metric.eval = c("F1", "Accuracy"),
                                    var.import = 3,
                                    seed.val = 42,
                                    nb.cpu = 1)

#for ordinal data this includes: Accuracy, Recall, F1 and F1 
print("Finished: Modeling (out)")


#Single model evaluation scores 

eval <- get_evaluations(myBiomodModelOut)

summary(eval)

write.csv(eval, "./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_single_model_scores.csv", row.names = FALSE)

var <- get_variables_importance(myBiomodModelOut)

summary(var)

write.csv(var, "./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_single_variable_importance.csv", row.names = FALSE)

print("Finished: Saving single model metrics")

plot_output <-  bm_PlotEvalMean(bm.out = myBiomodModelOut) 
summary(plot_output)

ggdat <- plot_output$tab
summary(ggdat)

#Graph single model evaluation scores

#model scores plot 1
plot_output <-  bm_PlotEvalMean(bm.out = myBiomodModelOut) 
summary(plot_output)

ggdat <- plot_output$tab
summary(ggdat)


names(ggdat) <- c('algo', 'Accuracy_mean','F1_mean', 'Accuracy_sd', "F1_sd") ###################NEED TO MAKE SURE THIS IS DOING WHAT ITS SUPPOSED TO 
ggdat$algo <- as.factor(ggdat$algo)

gg1 <-ggplot(ggdat, aes(x = Accuracy_mean, y = F1_mean, color = algo)) + 
  geom_point() + 
  geom_errorbarh(aes(xmax = Accuracy_mean + Accuracy_sd, xmin = Accuracy_mean - Accuracy_sd),height=0)+ 
  geom_linerange(aes(ymax= F1_mean + F1_sd, ymin=F1_mean - F1_sd)) +
  theme_bw()+ theme(legend.title = element_blank(), 
                    legend.key = element_rect(fill = "white"), 
                    legend.position = "bottom",
                    axis.text.x = element_text(angle = 45, hjust = 1))+ 
  theme_classic() + xlab("Accuracy")+ ylab("F1") +
  ggtitle(paste0("Comparison of model calibration metrics for ", sp," models" ))+
  scale_color_npg() 


gg1

# Save the plot
output_path <- file.path("./Plots/Eval_Plots/Ordinal/GURord_mean_model_scores.png")
ggsave(output_path, plot = gg1, width = 8, height = 6, dpi = 300)


#model scores plot 2
plot_output <- bm_PlotEvalBoxplot(bm.out = myBiomodModelOut, group.by = c('algo', 'algo'))
ggdat <- plot_output$tab
accuracy_data <- ggdat %>% filter(metric.eval == "Accuracy")
F1_data <- ggdat %>% filter(metric.eval == "F1")

# Create Accuracy calibration plot
gg_accuracy_cal <- ggplot(accuracy_data, aes(y = calibration, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("Accuracy (Calibration)") +
  xlab("Algorithm") +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("Accuracy Calibration Metrics for ", sp, " Models"))

# Create F1 calibration plot with y-axis from 0 to 1
gg_F1_cal <- ggplot(F1_data, aes(y = calibration, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("F1 (Calibration)") +
  xlab("Algorithm") +
  ylim(0, 1) +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("F1 Calibration Metrics for ", sp, " Models"))

# Create accuracy validation plot
gg_accuracy_val <- ggplot(accuracy_data, aes(y = validation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("Accuracy (Validation)") +
  xlab("Algorithm") +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("Accuracy Validation Metrics for ", sp, " Models"))

# Create validation validation plot with y-axis from 0 to 1
gg_F1_val <- ggplot(F1_data, aes(y = validation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "bottom"
  ) +
  ylab("F1 (Evaluation)") +
  xlab("Algorithm") +
  ylim(0, 1) +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("F1 Validation Metrics for ", sp, " Models"))

# Create Accuracy calibration plot
gg_accuracy_eval <- ggplot(accuracy_data, aes(y = evaluation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("Accuracy (Evaluation)") +
  xlab("Algorithm") +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("Accuracy Evaluation Metrics for ", sp, " Models"))

# Create R-squared calibration plot with y-axis from 0 to 1
gg_F1_eval <- ggplot(F1_data, aes(y = evaluation, x = algo, color = algo, fill = algo)) +
  geom_boxplot() +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "none"
  ) +
  ylab("F1 (Evaluation)") +
  xlab("Algorithm") +
  ylim(0, 1) +
  scale_color_npg() +
  scale_fill_npg() +
  ggtitle(paste0("F1 Evaluation Metrics for ", sp, " Models"))


# Combine all four plots into one figure
gg2 <- grid.arrange(gg_accuracy_cal, gg_F1_cal, gg_accuracy_val, gg_F1_val, gg_accuracy_eval, gg_F1_eval, ncol = 2, nrow = 3)

gg2

# Save the combined plot
output_path <- file.path("./Plots/Eval_Plots/Ordinal/GURord_combined_boxplot_model_scores.png")
ggsave(output_path, plot = gg2, width = 12, height = 10, dpi = 300)

#variable importance 

plot_output <- bm_PlotVarImpBoxplot(bm.out = myBiomodModelOut, group.by = c('expl.var', 'algo', 'algo')) 
ggdat <- plot_output$tab
ggdat <- ggdat %>%
  dplyr::mutate(expl.var.factor=as.factor(expl.var))%>%
  group_by(expl.var.factor,algo)%>%
  dplyr::summarize(mean = mean(var.imp),SD=sd(x=var.imp,na.rm=TRUE),N=sum(var.imp),SE= SD/sqrt(N))

gg3 <-ggplot(ggdat, aes(y = mean, x = reorder(expl.var.factor, -mean), color = algo, fill= algo)) + 
  geom_bar(position = "dodge",stat="identity") +
  geom_errorbar(aes(ymax=mean + SE, ymin = mean - SE),position="dodge",stat="identity",color="grey10")+
  ylim(-0.05,1)+
  theme_classic() + 
  theme(legend.title = element_blank(), 
        legend.key = element_rect(fill = "white"), 
        axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "bottom") + 
  ylab("Variable Importance")+ 
  xlab("Calibration")+ scale_color_npg()+ scale_fill_npg()+
  ggtitle(paste0("Variable importance for ", sp," models" ))
gg3 <- grid.arrange(gg3)

gg3

# Save the plot
output_path <- file.path("./Plots/Eval_Plots/Ordinal/GURord_single_model_variable_importance.png")
ggsave(output_path, plot = gg3, width = 8, height = 6, dpi = 300)

plot_output <- bm_PlotResponseCurves(
  bm.out = myBiomodModelOut,
  models.chosen = "all",
  fixed.var = 'mean',
  do.bivariate = FALSE,
  do.plot = FALSE,
  do.progress = FALSE
)

ggdat <- plot_output$tab %>%
  dplyr::mutate(expl.var.factor = as.factor(expl.name))

#write.csv(ggdat, "./Generated_Data/ES_datasets/GUR_ES_Ord_single.csv", row.names = FALSE)

# Assign algorithm labels based on prediction name
ggdat$algo <- ifelse(grepl("_GLM", ggdat$pred.name, ignore.case = TRUE), "GLM",
                     #ifelse(grepl("_GBM", ggdat$pred.name, ignore.case = TRUE), "BRT",
                            ifelse(grepl("_RF", ggdat$pred.name, ignore.case = TRUE), "RF",
                                   ifelse(grepl("_CTA", ggdat$pred.name, ignore.case = TRUE), "CTA",
                                          ifelse(grepl("_MARS", ggdat$pred.name, ignore.case = TRUE), "MARS",
                                                 ifelse(grepl("_XGBOOST", ggdat$pred.name, ignore.case = TRUE), "XGBOOST",
                                                        ifelse(grepl("_GAM", ggdat$pred.name, ignore.case = TRUE), "GAM", "NA"))))))


ggdat$algo <- as.factor(ggdat$algo)
ggdat$id <- as.factor(ggdat$id)

# Set custom colors
algo_colors <- c(
  "BRT" = "#FDE725FF",
  "GAM" = "#EAD7D1",
  "GLM" = "#238A8DFF",
  "MARS"= "red",
  "RF"= "blue",
  "CTA"="purple",
  "XGBOOST"="green",
  "Ensemble" = "#301934"
)

# Plot
gg4 <- ggplot(ggdat, aes(x = expl.val, y = pred.val, color = algo)) +
  geom_smooth(aes(group = algo), fill = "grey30", se = TRUE, method = "loess", alpha = 0.01) +  # By algorithm
  geom_smooth(color = algo_colors["Ensemble"], fill = "grey30", se = TRUE, method = "loess") + # Ensemble line
  facet_wrap(~ expl.var.factor, scales = "free_x") +
  xlab("Predictor Variable") +
  ylab("Prob. of Density") +
  ggtitle(paste0("Response curves for ", sp, " models")) +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_color_manual(values = algo_colors)

# Arrange and save
gg4 <- grid.arrange(gg4)

output_path <- file.path("./Plots/Eval_Plots/Ordinal/GURord_response_curves.png")
ggsave(output_path, plot = gg4, width = 8, height = 6, dpi = 300)

print("Finished: Graphing single model scores")

#Ensemble modelling ####################################################################
 
EnsembleModelsOut <- BIOMOD_EnsembleModeling(bm.mod= myBiomodModelOut,
                                             models.chosen = 'all', #define models kept, can exclude models here
                                             em.by='all', #chr, way kept models are combined to build ensemble models, see below 
                                             em.algo = c('EMwmean'), #weighted mean by the evaluation metric 
                                             metric.select = c('F1'), #metric select and threshold determines which single models to keep 
                                             metric.select.thresh = c(0.3),
                                             metric.select.dataset = c('evaluation'),
                                             metric.eval= c('F1'),
                                             EMci.alpha = 0.05, #num, significance level to estimate confidence interval
                                             EMwmean.decay ="proportional", # value defining relative importance of weights, discrimate good v bad models
                                             var.import=100) #vect, number of permutations to determine variable importance



print("Finished: Ensemble models")

#Save evaluation metrics
eval <- get_evaluations(EnsembleModelsOut)
#threshold 2 = 0.209
#treshold .23 = 0.209


summary(eval)

write.csv(eval, "./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_ensemble_model_scores.csv", row.names = FALSE)

var <- get_variables_importance(EnsembleModelsOut)

summary(var)

write.csv(var, "./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_ensemble_variable_importance.csv", row.names = FALSE)


print("Finished: Plotting saving ensemble model scores")

#Plot ensemble evaluation ################################################################


eval_scores <- eval
eval_predictor <- var

#Plot distributionf model evaluation scores by run 
#reshape, filter, and relabel
eval_long <- eval_scores %>%
  pivot_longer(cols = c(calibration, evaluation), 
               names_to = "type", 
               values_to = "value") %>%
  filter(metric.eval %in% c("F1"), !is.na(value)) %>%
  mutate(type = recode(type, 
                       calibration = "training", 
                       evaluation = "testing"),
         type = factor(type, levels = c("training", "testing")))  # set order

#plot
plot <- ggplot(eval_long, aes(x = type, y = value)) +
  geom_boxplot(fill = 'green') +
  facet_wrap(~metric.eval, scales = "free_y") +
  labs(x = NULL, y = "Value") +  # remove x-axis label
  theme_classic() +
  guides(x.sec = "axis", y.sec = "axis") +
  theme(
    axis.line.x.top = element_blank(),
    axis.line.y.right = element_blank(),
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    
    axis.ticks.x.top = element_blank(),
    axis.ticks.x.bottom = element_blank(),
    axis.ticks.y.right = element_blank(),
    
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    
    legend.position = "none",
    legend.justification = "center",
    
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 12),
    
    strip.background = element_blank(),
    strip.text = element_text(size = 10, face = "plain")
  )

plot 

#save the plot
output_path <- file.path("./Plots/Eval_Plots/Ordinal/GURord_ensemble_eval_scores.png")
ggsave(output_path, plot = plot, width = 8, height = 6, dpi = 300)

#Plot relative predictor importance 
predictor_summary <- eval_predictor %>%
  group_by(expl.var) %>%
  summarise(
    mean_var_imp = mean(var.imp),
    sd_var_imp = sd(var.imp),
    .groups = "drop"
  ) %>%
  mutate(
    var.imp.percent = mean_var_imp * 100,
    sd.percent = sd_var_imp * 100,
    min = var.imp.percent - sd.percent,
    max = var.imp.percent + sd.percent
  )

plot <- ggplot() +
  geom_col(
    data = predictor_summary,
    mapping = aes(
      x = var.imp.percent,
      y = reorder(expl.var, var.imp.percent)
    ),
    position = position_dodge2(padding = 0),
    width = 0.7,
    preserve = "total",
    fill = "green"
  ) +
  geom_errorbar(
    data = predictor_summary,
    mapping = aes(
      y = expl.var,
      xmin = min,
      xmax = max
    ),
    position = position_dodge2(preserve = "total"),
    width = 0.5,
    size = 0.7
  ) +
  theme_cowplot() +
  guides(x.sec = "axis", y.sec = "axis") +
  theme(
    axis.line.y.left = element_line(size = 0.75, color = "black"),
    axis.line.y.right = element_line(size = 0.75, color = "black"),
    axis.line.x.top = element_line(size = 0.75, color = "black"),
    axis.line.x.bottom = element_line(size = 0.75, color = "black"),
    axis.ticks.x.top = element_blank(),
    axis.ticks.y.right = element_blank(),
    axis.text.x.top = element_blank(),
    axis.text.y.right = element_blank(),
    axis.title = element_text(size = 16),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 9)
  ) +
  xlab("Predictor Importance (%)") +
  ylab("Predictor Variable")

plot


output_path <- file.path("./Plots/PI_Plots/GURord_ensemble_predictor_importance.png")
ggsave(output_path, plot = plot, width = 8, height = 6, dpi = 300)

#ES strip 
#response curves 
# Get response data
plot_output <- bm_PlotResponseCurves(
  bm.out = EnsembleModelsOut,
  models.chosen = "all",
  fixed.var = 'mean',
  do.bivariate = FALSE,
  do.plot = FALSE,
  do.progress = FALSE
)

ggdat <- plot_output$tab %>%
  dplyr::mutate(expl.var.factor = as.factor(expl.name))

write.csv(ggdat, "./Generated_Data/ES_datasets/GUR_ES_Ordinal.csv", row.names = FALSE)

# Assign algorithm labels based on prediction name
ggdat$algo <- ifelse(grepl("_GLM", ggdat$pred.name, ignore.case = TRUE), "GLM",
                     ifelse(grepl("_GBM", ggdat$pred.name, ignore.case = TRUE), "BRT",
                            ifelse(grepl("_RF", ggdat$pred.name, ignore.case = TRUE), "RF",
                                   ifelse(grepl("_CTA", ggdat$pred.name, ignore.case = TRUE), "CTA",
                                          ifelse(grepl("_MARS", ggdat$pred.name, ignore.case = TRUE), "MARS",
                                                 ifelse(grepl("_XGBOOST", ggdat$pred.name, ignore.case = TRUE), "XGBOOST",
                                                        ifelse(grepl("_GAM", ggdat$pred.name, ignore.case = TRUE), "GAM", "NA")))))))


ggdat$algo <- as.factor(ggdat$algo)
ggdat$id <- as.factor(ggdat$id)

# Set custom colors
algo_colors <- c(
  "BRT" = "#FDE725FF",
  "GAM" = "#EAD7D1",
  "GLM" = "#238A8DFF",
  "MARS"= "red",
  "RF"= "blue",
  "CTA"="purple",
  "XGBOOST"="green",
  "Ensemble" = "#301934"
)

# Plot
gg4 <- ggplot(ggdat, aes(x = expl.val, y = pred.val, color = algo)) +
  geom_smooth(aes(group = algo), fill = "grey30", se = TRUE, method = "loess", alpha = 0.01) +  # By algorithm
  geom_smooth(color = algo_colors["Ensemble"], fill = "grey30", se = TRUE, method = "loess") + # Ensemble line
  facet_wrap(~ expl.var.factor, scales = "free_x") +
  xlab("Predictor Variable") +
  ylab("Prob. of Density") +
  ggtitle(paste0("Response curves for ", sp, " models")) +
  theme_classic() +
  theme(
    legend.title = element_blank(),
    legend.key = element_rect(fill = "white"),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_color_manual(values = algo_colors)

# Arrange and save
gg4 <- grid.arrange(gg4)

output_path <- file.path("./Plots/ES_Plots/GUR_response_curves_ensemble.png")
ggsave(output_path, plot = gg4, width = 8, height = 6, dpi = 300)


print("Finished: Plotting ensemble models")
```

#Projections 
```{r}
gc()

ens_proj <- BIOMOD_EnsembleForecasting(
  bm.em = EnsembleModelsOut,
  new.env = proj.var,
  #bm.proj = sp_proj,
  proj.name = 'GUR_current_ord_ensembles',
  models.chosen = 'all',
  build.clamping.mask = TRUE, 
  output.format=".tif",
  do.stack=TRUE,
  #on_0_to_1000 = FALSE 
)

#this works below, from Jenn's code 
#sp_proj <- BIOMOD_Projection(
         #bm.mod =  myBiomodModelOut,
         #new.env = proj.var ,
         #proj.name = 'GUR_current_ord',
         #models.chosen =models,
         #digits = 2, #number of digits for predicted values
         #build.clamping.mask = TRUE, 
         #output_format = ".tif",
         #do.stack = TRUE,
         #nb.cpu = 10
       #)

gc()

#this does NOT run anymore on BIOMOD

#run projections on ensemble models, using current climatology
#ens_proj <- BIOMOD_EnsembleForecasting(
#         bm.em = EnsembleModelsOut,
#         new.env = proj.var,
         #bm.proj= sp_proj,
#         proj.name = 'test',
         #models.chosen = 'all',
         #build.clamping.mask = TRUE, 
##         output.format=".tif",
         #do.stack=TRUE,
         #on_0_to_1000 = FALSE 

```


#Null models
```{r}
gc()

#Create a dataframe of the environment to pull null model training dataset from 
Training_Data <- GUR_training_df  #needs to be a dataframe with xy, resp, env with only TRAINING data 
Testing_Data <- GUR_evaluation_df #needs to be a dataframe with xy, resp, env with only TESTING data 
Survey_Enviro_df <- env_df #includes all environmental data 
Predictor_num <- ncol(env_df)

#Create null models using the same parameters and randomly drawn presence points for the enviro_buffer dataframe (same background points)

#Set up a dataframe to hold all of the evaluation metrics for the null models
Null_metrics <- data.frame(
  full.name = character(),
  merged.by.PA = character(),
  merged.by.run = character(),
  merged.by.algo = character(),
  algo = character(),
  metric.eval = character(),
  cutoff = logical(),
  sensitivity= logical(),
  calibration = numeric(),
  validation = numeric(),
  evaluation = numeric()
)


#Run 1000 null models
for (j in 1:10){
  
  #Perform a random pull from the environmental dataframe for each presence point in the training dataset
  for (i in 1:nrow(Training_Data)) {
    set.seed(313*j + round(46*i))
    randnum <- sample(seq_len(nrow(Survey_Enviro_df)), size = 1)
    Training_Data[i, 4:(3+ Predictor_num)] <- Survey_Enviro_df[randnum,]
    
    train_resp_df <- Training_Data$mean
    train_xy <- Training_Data %>% dplyr::select(x, y)
    train_env <- Training_Data %>% dplyr::select(TAlk_10m_mean, salt_10m_min, temp_s_max, temp_s_min, temp_10m_mean, temp_10m_min, do_10m_mean, do_10m_min, Current, REI, Rock, Slope, Depth)
    
  }
  
  sp <- c('GUR')
  
  #define models to run 
all_models <- c("GAM" ) 
  
  #fit my *real* biomod2 models to null TRAINING data, made by for loop above
myBiomodData <- BIOMOD_FormatingData(resp.var = train_resp_df, #as SpatVector without associated data, for a single species
                                     expl.var = train_env, #SpatRaster containing explanatory variables in layers
                                     resp.name = sp, #character corresponding to species name
                                     resp.xy = train_xy, #if resp.var is vector, 2 column matrix associated xy coordinates
                                     dir.name = c('./models_biomod/null_ordinal/GUR'), 
                                     data.type = c("ordinal"), 
                                     eval.expl.var = eval_env, #SpatRaster without associated data, independent data to evaluate
                                     eval.resp.var = eval_resp_df, #SpatVector without associated data, independent data to evaluate 
                                     eval.resp.xy = eval_xy, #if resp.vect is vector, 2 column matrix of xy coordinates
                                     na.rm = TRUE, 
                                     filter.raster = TRUE, #filter points in same cell
                                     seed.val = 1234
) 

  
cv <- bm_CrossValidation(bm.format = myBiomodData,
                         strategy = "kfold", #options are random, k-fold, strat, block, env, or user-defined
                         nb.rep = 1, #integer, number of repetitions of CV points to be drawn 
                         k = 10, #number of partitions                                         
                         do.full.models = TRUE #calibrated and validated over the whole dataset (and pseudo-absence datasets) or not
) 
  
  

myBiomodModelOut <- BIOMOD_Modeling(bm.format = myBiomodData,
                                    models = all_models,
                                    CV.strategy = "user.defined",
                                    CV.user.table = cv,
                                    OPT.strategy = "user-defined",
                                    OPT.user = opt.e,
                                    metric.eval = c("F1", "Accuracy"),
                                    var.import = 3,
                                    seed.val = 42,
                                    nb.cpu = 14)

EnsembleModelsOut <- BIOMOD_EnsembleModeling(bm.mod= myBiomodModelOut,
                                             models.chosen = 'all', #define models kept, can exclude models here
                                             em.by='all', #chr, way kept models are combined to build ensemble models, see below 
                                             em.algo = c('EMwmean'), #weighted mean by the evaluation metric 
                                             metric.select = c('F1'), #metric select and threshold determines which single models to keep 
                                             #metric.select.thresh = c(0.45),
                                             metric.select.dataset = c('evaluation'),
                                             metric.eval= c('F1'),
                                             EMci.alpha = 0.05, #num, significance level to estimate confidence interval
                                             EMwmean.decay ="proportional", # value defining relative importance of weights, discrimate good v bad models
                                             var.import=100) #vect, number of permutations to determine variable importance

  #get model evaluation metrics 
  eval_ensemble <- get_evaluations(EnsembleModelsOut)
  eval_ensemble <- as.data.frame(eval_ensemble)
  
  #add eval_ensemble to empty df of null metrics above
  Null_metrics <- rbind(Null_metrics, eval_ensemble)
  
}


#save csv of Null Metrics 
write.csv(Null_metrics, "./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_null_model_eval.csv", row.names = FALSE)


```

#Plotting dataset
```{r}

Land <- terra::vect("./raw_data/BC_Buffer/Land.shp") 
 # Convert land to sf
  land_sf <- st_as_sf(Land)
  
  map_df <- cbind(resp_df, xy) %>% na.omit() 
  
  summary(map_df)
  
GUR_RA_Map <- ggplot() +
  geom_sf(data = land_sf, fill = "grey80", color = NA) +
  geom_point(
    data = map_df,
    aes(x = x, y = y, fill = mean, color = mean),
    size = 2,
    shape = 21,
    alpha = 0.2,
    stroke = 0.4
  ) +
  scale_fill_manual(
    name = "Relative Density",
    values = c("None" = NA, "Few" = "#cfeec9", "Many" = "#55C667FF", "Abundant" = "#1b5817"),
    na.translate = FALSE
  ) +
  scale_color_manual(
    name = "Relative Density",
    values = c("None" = "grey50", "Few" = "#cfeec9", "Many" = "#55C667FF", "Abundant" = "#1b5817"),
    na.translate = FALSE
  ) +
  annotation_scale(location = "bl", width_hint = 0.5) +
  annotation_north_arrow(
    location = "bl", which_north = "true",
    pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
    style = north_arrow_fancy_orienteering
  ) +
  coord_sf(crs = st_crs(3005), xlim = c(515000, 1255000), ylim = c(340000, 1100000)) +
  theme_cowplot() +
  xlab(NULL) + ylab(NULL) +
  theme(
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    axis.text.x = element_text(angle = 60, hjust = 1, size = 16),
    axis.text.y = element_text(size = 16),
    axis.line = element_blank(),
    panel.background = element_rect(fill = "white"),
    panel.border = element_rect(color = "black", fill = NA, size = 0.5),
    legend.position = 'top',
    legend.background = element_rect(fill = "white", color = NA)
  ) +
  guides(fill = guide_legend(override.aes = list(size = 4, alpha = 1)))


ggsave(
  filename = "./Plots/Maps/Species_RDensity_Maps/GUR_RelativeDataset_Map.png",   # or another path/filename
  plot = GUR_RA_Map,
  width = 10,        
  height = 8,       
  dpi = 300
)

```

#Creating uncertainty maps 
```{r}
myBiomodModelOut <- get(load('./models_biomod/ordinal/GUR/GUR.1756451565.models.out'))

#GUR.1756451565.models.out
Uncertainty_EnsembleModelsOut <- BIOMOD_EnsembleModeling(bm.mod= myBiomodModelOut,
                                             models.chosen = 'all', #define models kept, can exclude models here
                                             em.by='PA+algo', #chr, way kept models are combined to build ensemble models, see below 
                                             em.algo = c('EMwmean'), #weighted mean by the evaluation metric 
                                             #metric.select = c('F1'), #metric select and threshold determines which single models to keep 
                                             #metric.select.thresh = c(0.25),
                                             metric.select.dataset = c('evaluation'),
                                             metric.eval= c('F1'),
                                             EMci.alpha = 0.05, #num, significance level to estimate confidence interval
                                             EMwmean.decay ="proportional", # value defining relative importance of weights, discrimate good v bad models
                                             #var.import=100
                                             ) #vect, number of permutations to determine variable importance


print("Finished: Ensemble models")

#sp_proj <- BIOMOD_Projection(
#         bm.mod =  myBiomodModelOut,
#         new.env = proj.var ,
#         proj.name = 'GUR_current_ordinal_algos',
#         models.chosen =models.kept,
#         digits = 2, #number of digits for predicted values
#         build.clamping.mask = TRUE, 
#         output_format = ".tif",
#         do.stack = TRUE,
#         nb.cpu = 10
#       )

gc()

ens_proj <- BIOMOD_EnsembleForecasting(
  bm.em = Uncertainty_EnsembleModelsOut,
  new.env = proj.var,
 # bm.proj = sp_proj,
  proj.name = 'GUR_current_ordinal_ensembles_algos',
  models.chosen = 'all',
  build.clamping.mask = TRUE, 
  output.format=".tif",
  do.stack=TRUE,
  #on_0_to_1000 = FALSE 
)

Prediction <- terra::rast("./models_biomod/ordinal/GUR/proj_GUR_current_ordinal_ensembles_algos/proj_GUR_current_ordinal_ensembles_algos_GUR_ensemble.tif")

Prediction <- terra::rast("./models_biomod/ordinal/GUR/proj_GUR_current_ord/proj_GUR_current_ord_GUR.tif")

Model_Uncertainty <- terra::stdev(Prediction)


terra::writeRaster(Model_Uncertainty, "./processed_data/Uncertainty/GUR_Method_Uncertainty_RA.tif", overwrite = TRUE)

RC_Map <- function(RC_Layer, Legend_Label, layername, BC) {
  # Compute min and max values for the chosen layer
  vmin <- round(min(RC_Layer[[layername]], na.rm = TRUE), 2)
  vmax <- round(max(RC_Layer[[layername]], na.rm = TRUE), 2)
  
  # Calculate break points for the legend
  breaks <- round(seq(vmin, vmax, length.out = 5), 2)
  labels <- as.character(breaks)
  
  # Build the ggplot
  RC_Plot <- ggplot(BC) +
    geom_sf( data = BC, fill = "#d9d9d9", color = NA) +
    geom_raster(data = RC_Layer, aes(x = x, y = y, fill = .data[[layername]])) +
    coord_sf(crs = st_crs(3005)) +
    xlim(c(515000, 1255000)) +
    ylim(c(340000, 1100000)) +
    annotation_scale(location = "bl", width_hint = 0.5) +
    annotation_north_arrow(location = "bl", which_north = "true",
                           pad_x = unit(0.01, "in"), pad_y = unit(0.25, "in"),
                           style = north_arrow_fancy_orienteering) +
    theme_cowplot() +
    scale_fill_continuous(
      type = "viridis",
      na.value = NA,
      name = Legend_Label,
      limits = c(vmin - 0.01, vmax + 0.01),
      breaks = breaks,
      labels = labels
    ) +
    xlab(element_blank()) + ylab(element_blank()) +
    theme(
      legend.position = c(1, 1),
      legend.title = element_text(size = 11),
      legend.text = element_text(size = 10),
      legend.text.align = 0,
      legend.title.align = 0,
      legend.justification = c("right", "top"),
      legend.box.just = "right",
      panel.background = element_rect(fill = "white")
    )
  
  return(RC_Plot)
}

# Convert SpatRaster to data frame
RC_Layer_df <- as.data.frame(Model_Uncertainty, xy = TRUE)

# Convert SpatVector to sf
Land <- terra::vect("./raw_data/BC_Buffer/Land.shp")
Land <- sf::st_as_sf(Land)

# Create the plot
GUR_Method_Uncertainty_Plot <- RC_Map(RC_Layer_df, "Model\nUncertainty", "std", Land)

GUR_Method_Uncertainty_Plot

ggsave2(filename = "./Plots/Maps/Uncertainty_Plots/GUR_Model_Uncertainty_Map_RA.jpeg", plot = GUR_Method_Uncertainty_Plot, scale = 1, height = 140, width = 115, units = "mm")
```

#Response curves
```{r}

ord_GUR <- read.csv("./Generated_Data/ES_datasets/GUR_ES_Ordinal.csv")


sp <- c("GUR")
ggdat <- ord_GUR
ggdat$id <- as.factor(ggdat$id)

#Rename 
ggdat <- ggdat %>%
  mutate(
    expl.var.factor = recode(expl.var.factor,
                             "Current" = "Current (m/s)",
                             "Slope" = "Slope (deg)",
                             "Rock" = "Percent Rock",
                             "TAlk_10m_mean" = "Talkbmax (mEq/m³)",
                             "salt_10m_min" = "Sbmin (PSU)",
                             "temp_s_max" = "Tsmax (°C)",
                             "temp_s_min" = "Tsmin (°C)",
                             "temp_10m_mean" = "Tbmean (°C)",
                             "temp_10m_min" = "Tbmin (°C)",
                             "do_10m_mean" = "Oxybmean (ppm)",
                             "do_10m_min" = "Oxybmin (ppm)",
                             "Depth" = "Depth (m)"
    )
  )


ggdat <- ggdat %>%
  mutate(
    expl.name = recode(expl.name,
                       "Current" = "Current (m/s)",
                       "Slope" = "Slope (deg)",
                       "Rock" = "Percent Rock",
                       "TAlk_10m_mean" = "Talkbmax (mEq/m³)",
                       "salt_10m_min" = "Sbmin (PSU)",
                       "temp_s_max" = "Tsmax (°C)",
                       "temp_s_min" = "Tsmin (°C)",
                       "temp_10m_mean" = "Tbmean (°C)",
                       "temp_10m_min" = "Tbmin (°C)",
                       "do_10m_mean" = "Oxybmean (ppm)",
                       "do_10m_min" = "Oxybmin (ppm)",
                       "Depth" = "Depth (m)"
    )
  )



#Plot for most variables (y = 0–10)
p1 <- ggplot(ggdat, aes(x = expl.val, y = pred.val)) +
  geom_smooth(
    aes(group = expl.var.factor),
    method = "loess",
    se = TRUE,
    fill = "grey30",
    color = "#55C667FF"
  ) +
  facet_wrap(
    ~ expl.var.factor,
    scales = "free_x",
    strip.position = "bottom"
  ) +
  coord_cartesian(ylim = c(0, 2)) +  # Set y-axis range to 0–4 without trimming data
  labs(
    x = NULL,
    y = "Prob. of Relative Density"
  ) +
  theme_classic() +
  theme(
    legend.position = "none",
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_blank(),
    strip.placement = "outside"
  )

p1

#Save 
ggsave("./Plots/ES_Plots/GUR_ES_Ordinal.png", plot = p1, width = 10, height = 30, dpi = 300)


```
#Comparing null vs. model 
```{r}
#read in files 
models <- read.csv("./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_ensemble_model_scores.csv")

null <- read.csv("./Generated_Data/Evaluation_datasets/Ordinal/GUR/GURord_null_model_eval.csv")

#Format 
#Make a difference between testing and evaluation F1
models$difference <- abs(models$evaluation - models$calibration)
null$difference <- abs(null$evaluation - null$calibration)

#rename columns
models <- models %>%
  rename("F1 Training" = calibration,
         "F1 Testing" = evaluation, 
         "F1 Difference" = difference)


null <- null %>%
  rename("F1 Training" = calibration,
         "F1 Testing" = evaluation, 
         "F1 Difference" = difference)



#PLot
#Create long-form dataframes for the model and null model metrics for plotting
  Model_metrics_long <- models %>% pivot_longer(cols = c("F1 Training", "F1 Testing", "F1 Difference"), names_to = "Metric", 
                        values_to = "Value") %>% dplyr::select('Metric', 'Value')
  
Model_metrics_long$Metric <- factor(Model_metrics_long$Metric, levels = c("F1 Training", "F1 Testing", "F1 Difference"))
  
Null_metrics_long <- null %>%
  pivot_longer(cols = c("F1 Training", "F1 Testing", "F1 Difference"),
               names_to = "Metric", values_to = "Value") %>% dplyr::select(Metric, Value)


Null_metrics_long$Metric <- factor(Null_metrics_long$Metric, levels = c("F1 Training", "F1 Testing", "F1 Difference"))
  
  #Plot the model and null model evaluation metrics
  GUR_Eval_Plot <- ggplot() + geom_boxplot(data = Null_metrics_long, aes(x = Metric, y = Value, fill = Metric)) + 
    scale_fill_manual(values = c("#613f74", "#91B3B6", "#F4E76E", "#EDAF97"), labels = c("F1 Training", "F1 Testing", "F1 Difference")) + 
    stat_summary(data = Model_metrics_long, aes(x = Metric, y = Value), fun.y = mean, geom = "point", colour = "#e6360F", size = 3) + 
    xlab("Evaluation Metric") + ylab("Value") + theme_classic() + 
    guides(x.sec = "axis", y.sec = "axis")+ theme(axis.line.y.left = element_line(size = 0.75, color = "black"), axis.line.y.right = element_line(size = 0.75, color = "black"), axis.line.x.top = element_line(size = 0.75, color = "black"), axis.line.x.bottom = element_line(size = 0.75, color = "black"),axis.ticks.x.top = element_blank(), axis.ticks.x.bottom = element_blank(), axis.ticks.y.right = element_blank(), axis.text.x.top = element_blank(), axis.text.y.right = element_blank()) + 
    theme(legend.position = "none", legend.justification = "center") + 
    theme(axis.text = element_text(size=8), axis.title = element_text(size = 12))
  
  
  GUR_Eval_Plot
  

  ggsave2(filename = "./Plots/Eval_Plots/Ordinal/GUR_Eval_Plot.jpeg", plot = GUR_Eval_Plot, scale = 1, height = 75, width = 100, units = "mm")

```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
